### 2017-12-24 ManuscriptGeophysics2018.tex
Add manuscript

% \documentclass[manuscript]{geophysics}
\documentclass[paper,twocolumn,twoside]{geophysics}
% \documentclass[paper]{geophysics}
% \documentclass[paper, revised]{geophysics}
% \documentclass[manuscript,revised]{geophysics}
%%fakesection ===    PACKAGES & DEFINITIONS    ===
% \usepackage{showframe}

%% ~ ADDITIONAL PACKAGES TO GEOPHYSICS.CLS
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\usepackage{color}
\usepackage{tabularx}
\usepackage{colortbl}
\usepackage{booktabs}
\usepackage[USenglish]{babel}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{amssymb, amsmath, amsfonts}
\usepackage{upquote}
\usepackage[pdftex, final]{hyperref}
\hypersetup{allcolors=blue, allbordercolors={0 0 .5}, colorlinks=true}

\usepackage{listings}

% Figure directory
\renewcommand{\figdir}{./figures}
\usepackage{xspace}
\newcommand{\empymod}{\texttt{empymod}\xspace}
\usepackage[strings]{underscore}

\definecolor{MyGray}{gray}{0.85}
\newcolumntype{Y}{>{\columncolor{MyGray}\raggedleft\arraybackslash}r}
\newcolumntype{W}{>{\raggedleft\arraybackslash}r}
\newcolumntype{A}{>{\columncolor{MyGray}\raggedleft\arraybackslash}X}
\newcolumntype{B}{>{\raggedleft\arraybackslash}X}

\hyphenation{iso-tro-pic pe-ne-tra-ting}

\newcommand{\rmk}[1]{\textbf{\color{red}{#1}}}
\newcommand{\mr}[1]{\mathrm{#1}}
\frenchspacing

\begin{document}

\title{A digital filter designing tool for the Hankel and Fourier transforms in
potential-, diffusive-, and wavefield modeling}

\renewcommand{\thefootnote}{\fnsymbol{footnote}}

\ms{GEO-2018-????}

\address{
\footnotemark[1]Instituto Mexicano del Petróleo,
Eje Central Lázaro Cárdenas Norte 152,
Col. San Bartolo Atepehuacan C.P. 07730,
Ciudad de México, México,
E-mail: \href{mailto:dieter@werthmuller.org}{Dieter@Werthmuller.org};
\footnotemark[2]Lamont-Doherty Earth Observatory,
305C Oceanography,
61 Route 9W, PO Box 1000,
Palisades NY 10964-8000 US,
E-mail: \href{mailto:KKey@ldeo.columbia.edu}{KKey@ldeo.columbia.edu};
\footnotemark[3]TU Delft,
Building 23,
Stevinweg 1 / PO-box 5048,
2628 CN Delft,
E-mail: \href{mailto:E.C.Slob@tudelft.nl}{E.C.Slob@tudelft.nl}.}

\author{%
Dieter Werthmüller\footnotemark[1], %
Kerry Key\footnotemark[2], and %
Evert C.\ Slob\footnotemark[3]%
}

\footer{}
\lefthead{Werthmüller et al.}
\righthead{Digital filter designing tool}

\maketitle

%%fakesection ===    ABSTRACT    ===
\begin{abstract} % 1-2 sentence(s) each
%
% (1) PRINCIPAL OBJECTIVES AND SCOPE OF THE WORK
  The open-source code \texttt{fdesign} makes it possible to design digital
  linear filters for the Hankel and Fourier transforms used in potential-,
  diffusive-, and wavefield modeling. Digital filters can be derived for any
  electromagnetic method, such as methods in the diffusive limits (DC, CSEM) as
  well as methods using higher frequency content (GPR, acoustic and elastic
  wavefields).
%
% (2) METHODOLOGY
  The direct matrix inversion method is used for the derivation of the filter
  values, and a brute-force inversion is carried out over the defined spacing
  and shifting values of the filter basis. Included or user-provided
  theoretical transform pairs are used for the inversion. Alternatively, one
  can provide 1D subsurface models that will be computed with a precise
  quadrature method using the EM modeler \texttt{empymod} to generate
  numerical transform pairs.
%
% (3) RESULTS
  The comparison of the presented 201\,pt filter with previously presented
  filters shows that it performs better for some standard CSEM cases. The
  derivation of longer 2001\,pt filter for a GPR example with 250\,MHz centre
  frequency proves that the filter method works also for wave phenomena, not
  only for diffusive EM fields.
%
% (4) CONCLUSIONS
  The presented algorithm provides a tool to create problem specific digital
  filters. Such purpose-built filters can be made shorter and can speed up
  consecutive  potential-, diffusive-, and wavefield inversions.
%
\end{abstract}

\section{Introduction}

Deva Prasad Ghosh proposed in his Ph.D. thesis \citep{PhD.70.Ghosh} a linear
filter method for the Hankel transform which revolutionized the computational
modeling of electromagnetic responses in the field of geophysical exploration.
If you use a code that calculates electromagnetic responses in the
wavenumber-frequency domain and transforms them to the space-frequency domain
chances are very high that it uses the \emph{digital linear filter} method
(DLF). \cite{GP.71.Ghosh} states that the idea is based on suggestions made
four decades earlier by \cite{PHY.33.Slichter} and \cite{GEO.40.Pekeris}, in
that ``\emph{the kernel function is dependent only on the layer parameters, and
  an expression relating it to the field measurements can be obtained by
mathematical processes.}'' However, until the introduction of DLF by Ghosh
these suggestions found no widespread use, probably not last because of the
missing computer power to calculate the filter coefficients. He further states
that credit goes to his Ph.D. supervisor Otto \cite{BK.68.Koefoed,
GP.70.Koefoed}, who retook the task of direct interpretational methods with the
introduction of raised kernel functions. DLF is, as such, an improvement of
that approach, providing a faster and simpler method.

% \subsection{Review}

The introduction of linear filters to electromagnetic geophysics initiated a
wealth of investigation, publications, and development of computer programs
that extended and improved the method. A few dozen articles have been published
in the 70s and 80s by Ghosh, Koefoed, Das, Anderson, and many others, the vast
majority of it in the journal \emph{Geophysical Prospecting}.
%  (see Figure~\ref{fig:authors})
Another common name of DLF is \emph{fast Hankel transform} (FHT), popular
because of the similarity of the name to the well known \emph{fast Fourier
transform} (FFT). The name was introduced, as far as we can see, by
\cite{GP.79.Johansen} in their article of the same name. However, the name FHT
can be misleading as it has \emph{Hankel} in the name, but the DLF can also be
applied to the Fourier transform (and other linear transforms).
% \begin{figure}[tbp]
%   \centering
%   \fbox{\includegraphics[width=10cm]{figures/author-figure}}
%     \caption{List by year of publications included in the review (year and
%       first three letters of principal author).}
%     \label{fig:authors}
% \end{figure}
Publications fall broadly into one or several of three categories: (1) new
methods (application of DLF to new measurement techniques); (2) filter
improvements (new filters or providing new or improved methods for the
determination of filter coefficients); and (3) computational tools. The
following is a brief review of the most relevant publications regarding DLF,
without claiming completeness.

\subsubsection{(1) New methods}

Ghosh used the method originally for the computation of type curves for
Schlumberger and Wenner resistivity soundings. He published the main results of
his thesis the following year in two publications: in \cite{GP.71.Ghosh} he
derives a resistivity model from given Schlumberger or Wenner sounding curves;
and in \cite{GP.71.Ghoshb} he provides filters for the inverse operation,
deriving resistivity sounding curves from a given resistivity model.  The
method was next applied to electromagnetic soundings with horizontal and
perpendicular coils \citep{GP.72.Koefoed}, to vertical coplanar coil systems
\citep{GP.73.Verma}, to dipoles and other two electrode systems
\citep{GP.74.Das, GP.74.Dasb, GP.80.Das, GEO.94.Sorensen}, and to vertical
dikes, hence vertical instead of horizontal layers \citep{GEO.75.Niwas}. Whilst
the first filters were very specific to one type of curves and one type of
transformation, various publications used the method to get one type curve from
another type curve \citep{GP.77.Kumar, GP.78.Kumar} or generalized the method
to be applicable to a wider set of problems \citep{EXG.80.Davis, GXP.81.Das,
GEO.84.Das, GP.84.ONeill}. Eventually, it passed from pure layered modeling to
primary-secondary formulations in 3D problems, where DLF is used to compute the
spatial Fourier-Hankel transforms in a horizontally layered background medium
and to compute transient responses from frequency domain computations
\citep{GJI.81.Das, GJI.82.Das, GEO.84.Anderson, GEO.86.Newman,
MGS.17.Kruglyakov}. Various authors delved into the theory of the method,
analyzing the oscillating behaviour of the filters and trying to estimate the
error of DLF \citep{GP.72.Koefoedb, GP.76.Koefoedb, GP.79.Johansen,
GP.90.Christensen}.

\subsubsection{(2) Filter improvements}

\cite{PhD.70.Ghosh} derived the filter coefficients in the spectral domain by
dividing the output spectrum by the input spectrum followed by an inverse
Fourier transform. Improvements to the determination of filter coefficients
were provided by \cite{EXG.75.ONeill, GEO.77.Nyman, GEO.82.Das}, or
specifically for the Fourier transform by \cite{GP.86.Nissen}. A direct
integration method was used by \cite{GP.76.Bichara, GP.78.Bernabini}.
\cite{GP.79.Koefoed} proposed a Wiener-Hopf least-squares method, which was
further improved by many authors \citep{GP.82.Guptasarma, GEO.1982.Murakami,
GP.97.Guptasarma}. \cite{GP.07.Kong} proposes a direct matrix inversion method
to solve the convolution equation, which requires only the input and output
sample values. To evaluate the filters he defines the criteria of a good filter
as one that recovers small or weak wavefields. This method was also used by
\cite{GEO.09.Key, GEO.12.Key}. Most authors publish filters for the Hankel
transform with $J_0$ and $J_1$ Bessel functions (or $J_{-1/2}$, $J_{1/2}$ if
applied to the Fourier sine/cosine transform), as all higher Bessel functions
can be rewritten to only use these two. \cite{GP.94.Mohsen} is one of the rare
cases which provides $J_2$ filter weights.

\subsubsection{(3) Codes}

The best known codes are likely the freely available ones by Anderson.
\cite{USGS.73.Anderson} extends the method to transient responses, applying DLF
not only to the Hankel transform, but also to the Fourier transform. A
transient signal can therefore be obtained be applying twice a digital filter
to the wavenumber-frequency domain calculation. In \cite{USGS.75.Anderson,
GEO.79.Anderson} he presents improved filters for both Fourier and Hankel
transforms, introducing measures to significantly speed-up the calculation,
such as the lagged convolution or using the same abscissae for $J_0$ and $J_1$.
\cite{TMS.82.Anderson} and the included 801\,pt filter became sort of the
industry standard, to which subsequent filters were compared. In
\cite{GEO.89.Anderson} he presents a hybrid solution, permitting to use either
DLF or quadrature and as such permits to compare the two. Other examples
include the codes by \cite{GP.75.Johansen}, an interactive system for
interpretation of resistivity soundings, and a tool to calculate filter
coefficients by \cite{GP.90.Christensen}. The latter is available upon request
and was used, for instance, in all the open-source modeling and inversion
routines of CSIRO in the Amira Project 223 \citep{ASEG.07.Raiche}. \newline

% \subsection{Outline}

All mentioned publications have in common that they were derived for direct
current methods (DC) or low frequency methods, such as time-domain shallow EM
methods (TEM), or controlled-source electromagnetics (CSEM), but not for high
frequency methods such as ground-penetrating radar (GPR). Generally it was even
thought that the filter method works only in the diffusive limit
\citep[e.g.,][]{GEO.15.Hunziker}. Also, there is no open-source filter
designing tool readily available, they are available as a described method
in articles or as code upon request by the author, if at all. The presented
algorithm \texttt{fdesign} tries to fill this gap by providing the tools to
design general or purpose-built filters using the direct matrix inversion
method. After a brief overview of the methodology, the theory, and the code we
show examples of its usage, presenting some new filters for both, CSEM and GPR
data. The algorithm and many more examples of its usage can be found on
\href{https://github.com/empymod/article-fdesign}{github.com/empymod/article-fdesign}.
The examples can be used as templates to design new filters.

\section{Methodology}

The algorithm \texttt{fdesign} is a filter designing tool using the direct
matrix inversion method as described by \cite{GP.07.Kong} and based on scripts
by \cite{GEO.12.Key}. The tool is an add-on to the electromagnetic modeler
\texttt{empymod} \citep{GEO.17.Werthmuller}, written in Python, and hosted on
GitHub, which should foster interaction and enable anyone to toy around,
improve, and extend it. It can be used to derive digital linear filters for the
Hankel and the Fourier transforms for potential-, diffusive-, and wavefields
(hence from DC to GPR). Theoretically, it can be used to derive linear filters
for any linear transform, as long as you feed it with a theoretical, or
accurately computed, transform pair of the transform. It permits to derive
filters with the help of theoretical transform pairs or, alternatively, with
the electromagnetic modeler \texttt{empymod}, using a quadrature method to
derive accurate curves used for the inversion.

The main points of the method and some differences between it and the methods
it is based upon are:
\begin{itemize}
  \item The algorithm computes many different filters for various spacing and
    shift values (brute-force).
  \item \cite{GP.07.Kong, GEO.12.Key} optimize a filter for the $J_0$ filter,
    and then use the obtained shift and spacing value to calculate the $J_1$
    values. The presented algorithm can optimize either $J_0$, $J_1$, or both
    in the optimization process. Five transform-pairs for each $J_0$ and $J_1$,
    and three transform pairs for each sine and cosine of often used functions
    are included in the routine, but any other transform pair can be provided
    as input.
  \item More complex or more specific models can be provided instead of
    theoretical transform pairs by using the modeler \texttt{empymod}.
  \item \cite{GP.07.Kong} defines a \emph{good} filter as one that recovers
    \emph{weak} diffusive EM fields. In the presented algorithm you can define
    a relative error level which defines up to what error an obtained result is
    good or not. The presented algorithm can also minimize the amplitude, but
    it has additionally another mode, where you maximise the abscissae $r$ of
    the right-hand-side of the transform pair instead of minimising the
    amplitude. The two modes yield the same result in the general, simple case
    of fast decaying transform pairs. However, if you use complex models for
    the design, and specifically high frequencies, then maximising $r$ will
    yield much better and more consistent results.
  \item The algorithm allows to run under-, equal-, and over-determined
    systems.
  \item the real or the imaginary part can be used for the inversion of complex
    transform pairs.
\end{itemize}

The quality of a filter depends heavily on the model chosen for comparison. An
obtained filter might be very good for one model, but not that good for another
one, which is no different for this designing tool. There is no way to estimate
the error of a result obtained with a certain filter if you apply it to any
other model, unless you calculate this other model with another method for
comparison. All results presented here should therefore be taken with a certain
care.

\subsection{Theory}

Most of the articles mentioned in the review have detailed derivations of the
digital filter method. In this article we focus on the algorithm, and summarize
the theory only very briefly by following \cite{GEO.12.Key} (equations
(2)-(6)). In electromagnetics we often have to evaluate integrals of the form
%
\begin{equation}
  F(r) = \int^\infty_0 f(l)K(l r)\mr{d}l \ ,
  \label{eq:HankelInt}
\end{equation}
%
where $l$ and $r$ denote left-hand-side and right-hand-side evaluation values,
and $K$ is the kernel function. In the specific case of the Hankel transform
$l$ corresponds to wavenumber, $r$ to offset, and $K$ to Bessel functions; in
the case of the Fourier transform $l$ corresponds to frequency, $r$ to time,
and $K$ to sine or cosine functions. In both cases it is an infinite integral
which numerical integration is very time-consuming because of the slow decay
of the kernel function and its oscillatory behaviour.

By substituting $r = e^x$ and $l = e^{-y}$ we get
\begin{equation}
  e^x F(e^x) = \int^\infty_{-\infty} f(e^{-y})K(e^{x-y})e^{x-y}\mr{d}y\ .
  \label{eq:filtint}
\end{equation}
This can be re-written as a convolution integral and be approximated for an
$N$-point filter by
\begin{equation}
  F(r) \approx \sum^N_{n=1} \frac{f(b_n/r) h_n}{r}\ ,
  \label{eq:filtapprox}
\end{equation}
where $h$ is the digital linear filter, and the logarithmically spaced filter
abscissae is a function of the spacing $\Delta$ and the shift $\delta$,
\begin{equation}
  b_n = \exp\left\{\Delta(-N/2+n) + \delta\right\} \ .
  \label{eq:base}
\end{equation}
From equation~\ref{eq:filtapprox} it can be seen that the filter method
requires $N$ evaluations at each $r$. So to calculate for instance the
frequency domain result for 100 offsets with a 201\,pt filter requires 20'100
evaluations in the wavenumber domain. This is why the DLF often uses
interpolation to minimise the required evaluations, either in the right-hand
side or in the left-hand side.

\subsection{Pseudo-code}
The main input variables are the filter length ($N$), the spacing ($\Delta$)
and shift ($\delta$) ranges over which to loop, and the transform pairs for
the inversion ($f_\mr{I}$) and the check of quality ($f_\mr{C}$). If
$f_\mr{C}$ is not provided, then $f_\mr{I}$ is used for both. There are
additional, optional input parameters, for instance to adjust how the rhs
abscissae $r = f(b, N)$ are calculated, where $b$ is the filter base.

The basic steps are as follows:
\begin{enumerate}
  \item Evaluate rhs of check-function $f_\mr{C}$:\newline
        $d_\mr{R} = f_\mr{C}.\mr{rhs}(r)$
  \item Loop over each value $\Delta_i$, $\delta_j$ (brute force):
    \begin{enumerate}
      \item Calculate filter base:\newline
        $b_n = \exp\left\{\Delta_i(-N/2+n) + \delta_j\right\}$
      \item Get required rhs ($r$) and lhs ($l$) evaluation points:\newline
        $r = f(b, N)$\newline
        $l = b/r$
      \item Invert for filter coefficients:\newline
        $J_{ij} = \mr{solve}(f_\mr{I}.\mr{lhs}(l), f_\mr{I}.\mr{rhs}(r))$
    \item Calculate numerically rhs of check-function $f_\mr{C}$ with current
      filter $J_{ij}$:\newline
      $d_\mr{F} = f_\mr{C}.\mr{lhs}(l)\cdot J_{ij}\ /\ r$
      \item Store minimum recovered amplitude or maximum $r$ where relative
        error is less than the provided error:\newline
        $\chi_{ij} = g\left[
        \mr{argmin}\left(|(d_\mr{F}-d_\mr{R})/d_\mr{R}| >
        \mr{error}\right) -1\right]\ $, where $g$ is either $d_\mr{R}$
        or $1/r$.
    \end{enumerate}
  \item Return filter coefficients which yield minimum amplitude or maximum $r$
    (a local minimization can be run to polish the brute-force result):\newline
    return $J\left[\mr{argmin}(\chi_{ij})\right] $\newline
\end{enumerate}

\subsection{Minimization criteria}
Figure~\ref{fig:InvCrit} shows the differences between the different
minimization approaches; in (a) for a conventional, fast decaying transform
pair; and in (b) for a more complex, high frequency layered model. The circles
show the minimum amplitude used in previous approaches. This criteria is not
ideal, as it is subject to some random fluctuations and also depends on the
choice of $r$. The squares show the minimum amplitude given a certain
acceptable error, and the diamonds show the maximum $r$ given a certain
acceptable error (the inversion is a minimization process, it therefore
minimizes $1/r$, not $r$). In simple cases the minimum amplitude given an
acceptable error and the maximum $r$ given an acceptable error will yield the
same result. However, in complex cases the maximum $r$ is more consistent and
therefore a much better criteria.
%
\plot*{InvCrit}{width=.9\textwidth}{(a) A regular rhs curve of a transform pair
  with a purely decaying function. Minimum amplitude or maximum $r$ yield the
  same result in this scenario. Using a relative error criteria is more stable
  than just the absolute minimum amplitude. (b) A 1D model for $f=80\,$MHz.
  Here, the maximum $r$ provides a better criteria for the inversion.}
%

It is important to note that independent of the inversion criteria the obtained
absolute value of it is only good to compare inversion results for the same
model. If you calculate filters with a different transform pair you might get
very different values, but this difference says nothing about the quality of
it. For instance, a certain transform pair might yield a lower minimum
amplitude than the other transform pair, but the resulting filter is worse if
both filters are compared to the same transform pair. This is due to the
characteristic of each transform pair.

\section{Numerical examples}

The numerical examples are focused on the Hankel transform, although
\texttt{fdesign} can be used to design digital linear filters for both Hankel
and Fourier transforms. As there is no difference in the procedure of the two,
this should be sufficient to demonstrate the algorithm.

\subsection{Design}
Figure \ref{fig:201} shows the solution spaces of four consecutive inversion
runs, where each run is a focus on a subsection of the previous run, indicated
by the red square.
%
\plot*{201}{width=\textwidth}{%
  Solution spaces of four consecutive inversion runs for a 201\,pt filter. Each
  consecutive run zooms into a portion of the previous solution space,
  indicated by the red square. The more in detail we obtain the solution space,
  the more random appears to be the distribution.}%
%
The filter length in this case is $N=201$, and the theoretical transform pairs
is (as given, e.g., in \cite{USGS.75.Anderson})
%
\begin{align}
  \int^\infty_0\,l \exp\left(-al^2\right) J_0(rl)\,dl &=
  \frac{\exp\left(\frac{-r^2}{4a}\right)}{2a}\ , \\
  \int^\infty_0\,l^2 \exp\left(-al^2\right) J_1(rl)\,dl &=
  \frac{r}{4a^2} \exp\left(-\frac{r^2}{4a}\right)\ ,
  \label{eq:j01}
\end{align}
%
where $a$ was set to 5. In the algorithm you can provide one pair for the inversion, and a different pair to get the minimum amplitude or the maximum $r$. In
this case the same transform pair was used for both the inversion and the check
of quality. The acceptable relative error was set to 1\,\%.
The rhs evaluation parameter for the inversion was set to $r_\mr{def} = (1, 1,
2)$, which means that an over-determined system was evaluated with $2N$
equations, where $r$ was logarithmically spaced from $\log_{10}(1/\max(b)) - 1$
to $\log_{10}(1/\min(b)) + 1$, where $b$ are the filter abscissae as given in
equation~\ref{eq:base}.

From the low resolution overview runs (a) and (b) it looks like this is a
standard, straight-forward minimization problem. However, from the more
detailed results in (c) and (d) it becomes obvious that it is a minimization
problem that has to be solved stochastically, as there are solutions with 2
orders of magnitude difference apparently randomly next to each other.

Figure \ref{fig:201b} shows in (a) the filter values for $J_0$ and $J_1$ of the
best 201 point filter obtained this way, and in (b) its right-hand-side
solution. The black dots indicate negative values, from which it can seen that
adjacent values are often alternating between positive and negative
contributions.
%
\plot*{201b}{width=.9\textwidth}{(a) Filter values of the best obtained 201\,pt
filter with the corresponding check of quality in (b). Black points indicate
negative values, which shows that adjacent values have often opposite signs.}
%

Designing filters is to a large extent trial and error. All input variables
influence the outcome, and often you will come across a good filter by shear
luck of choosing the right starting parameters. Each of the input parameters
has an effect to the outcome. It depends a lot on the transform pairs
$f_\mr{I;C}$, and functions that decay rapidly are generally better, as noted
by earlier authors \citep[e.g.][]{USGS.75.Anderson}. It also depends on the
filter length, and obviously heavily on the spacing and shift values, as this
is what we invert for. Another important point is how you define the right hand
evaluation points of the inversion ($r_\mr{def}$). Evaluating corresponding
transform pairs separately or jointly also leads to different filter
coefficients ($J_0$, $J_1$, or $J_0$ \& $J_1$, or equally sine, cosine, or sine
\& cosine); if the real or the imaginary part is used when complex transform
pairs or \texttt{empymod} is used; and if it is inverted for the minimum
amplitude or for the maximum $r$.

\subsection{CSEM}
In this section we compare the 201\,pt filter derived in the previous section
to CSEM models used in \cite{GP.07.Kong}, \cite{GEO.12.Key}, and a land case.

Figures \ref{fig:kongf5} compares the derived 201\,pt filter with the two
half-spaces case used by \cite{GP.07.Kong} in his figure 5. The model consists
of a water layer with $\rho_\mr{w} = 0.3125\,\Omega\,$m of infinite thickness
above a subsurface half-space with $\rho = 1\,\Omega\,$m. The signal of an
$x$-directed electric source 50\,m above the interface is measured at an
$x$-directed electric receiver at the interface, frequency is $f=1\,$Hz.
%
\plot*{kongf5}{width=.9\textwidth}{(a) Results of different filters and QWE for
  the model of figure 5 from \cite{GP.07.Kong} with the relative errors shown
  in (b), using the QWE result. The relative error is meaningless from about
  15.5\,km onwards, as QWE failed as well for these very low amplitudes.}
%
In (a) it can be seen that the new 201\,pt filter is able to recover smaller
amplitudes than the 241\,pt filter from \cite{GP.07.Kong}, the 201\,pt filter
from \cite{GEO.12.Key}, and the 801\,pt filter from \cite{TMS.82.Anderson}
(Wer201, Kong241, Key201, and And801, respectively). It behaves equally well as
the quadrature with extrapolation (QWE), for which we used a 51\,pt quadrature
with relative and absolute tolerance of 1e-12 and 1e-30, respectively. The
relative error is shown in (b), where the QWE result was taken as \emph{truth}.
For offsets greater than roughly 15.5\,km the relative error becomes
meaningless, as the QWE fails itself; this part is greyed out in the figure.

Figure \ref{fig:keyf5} (a) is the canonical CSEM model as used in
\cite{GEO.12.Key} in his figure 5: Water depth of 2\,km with
$\rho_\mr{w}=0.303\,\Omega\,$m, a background resistivity of
$\rho_\mr{b}=1\,\Omega\,$m, in which a target is embedded of
$\rho_\mr{t}=100\,\Omega\,$m at 1000\,m below the seafloor, 100\,m thick.
Source depth is 1990\,m, receivers are on the seafloor.
%
\plot*{keyf5}{width=.9\textwidth}{(a) The canonical marine CSEM model of
  \cite{GEO.12.Key} and (b) a land model with source and receiver at the
  surface. The new filter has generally relative errors which are several
  magnitudes lower than the other filters.}
%
Figure \ref{fig:keyf5} (b) is a land case, with a background resistivity of
$\rho_\mr{b}=10\,\Omega\,$m, in which a target is embedded of
$\rho_\mr{t}=500\,\Omega\,$m at 1000\,m below the surface, 100\,m thick. Source
depth is 0.5\,m, receiver depth is 0.8\,m. In both cases the new filter
\emph{Wer201} has generally a relative error which is magnitudes lower than the
other filters.

It is very important to note again that other scenarios might yield very
different error plots. Although the new 201\,pt filter proves to be very
accurate for these three models, it might not be the best filter in other
cases.

\subsection{Ground-penetrating radar}

\rmk{TODO: Add second GPR result; layers swapped}

Figure~\ref{fig:GPR} shows the GPR example as calculated in
\cite{GEO.15.Hunziker, GEO.17.Werthmuller}; the model parameters are given in
subplot (a). The filter used for this example for the Hankel transform is a
2001\,pt filter, derived with the fullspace solution with $f=500\,$MHz for the
inversion and the check of quality. Vertical source and receiver separation is
1\,m, resistivity $\rho=200\,\Omega\,$m, $\varepsilon_\mr{r}$=10,
$\mu_\mr{r}=1$. For the Fourier transform a 4096\,pt FFT was used with
regularly spaced frequencies from 0.5\,MHz to 850\,MHz and then zero-padded up
to 2048\,MHz for both the calculation with the adaptive quadrature (QUA) and
with the DLF. The frequency result is convolved with a Ricker wavelet with a
center frequency of 250\,MHz, and a gain function ($1 + |t^3|, t$ in ns) is
applied. The calculation with the digital filter took under 9 minutes and is
therefore roughly 80 times faster then the adaptive quadrature calculation
which took roughly 11 hours and 27 minutes. Calculating the same model with QWE
took 7 hours and 20 minutes. However, QWE uses in roughly 1/3 of the
calculation internally the adaptive quadrature in this example, see
\cite{GEO.17.Werthmuller}. (Note that DLF was run in parallel using 4 threads
at once, taking effectively only 2 minutes and 10 seconds to calculate.
%EMmod \citep{GEO.15.Hunziker} took for the same model about 18 hours 55
%minutes, using a 61\,pt Gauss-Kronrod integration routine.
The lagged convolution version of DLF and the splined version of QWE were used
in this comparison.)
%
\plot*{GPR}{width=\textwidth}{GPR example for the model given in (a) using
  \texttt{empymod} with (b) the adaptive quadrature (QAGSE from the Fortran
  QUADPACK library) or with (c) a 2001\,pt digital linear filter (b). The
  quadrature took about 11\,h 27\,min to calculate, whereas the filter took
  2\,min 10\,s.}
%

Figure~\ref{fig:ftGPR} shows in (a)-(c) the real part of the frequency domain
results and in (d)-(f) the time domain results for offsets of 0.2\,m, 2.0\,m,
and 3.0\,m.
%
\plot*{ftGPR}{width=\textwidth}{The real-part frequency-domain and the
time-domain responses for offsets of (a) 0.2\,m, (b) 2.0\,m, and (c) 3.0\,m.}
%
These examples show clearly that the filter method can indeed be applied to
high frequency EM modeling and therefore wave phenomena. We are convinced that
with further tests and analysis much better filters could be achieved, and
various concepts could be checked. One approach is to derive a filter for each
frequency band, say one for 1\,MHz--10\,MHz, one for 10\,MHz--100\,MHz, and one
for 100\,MHz--1\,GHz. Another approach could be to derive distinct filters for
$J_0$ and $J_1$, with different spacing and shift values. The first idea would
triple the calculation cost, the second idea double them; however, they both
would still be very fast compared to standard quadrature methods.

\section{Conclusions}

The presented, free and open-source algorithm \texttt{fdesign} can be used to
design digital linear filters for the Hankel and Fourier transforms (and more
generally for any linear transform) using either analytical transform pairs or
1D subsurface models together with the EM-modeler \texttt{empymod}.
The code is available from GitHub in the add-on repository of \texttt{empymod},
called \texttt{empyscripts}, from version 0.1.2 onwards.

The presented 201\,pt filter achieves more precise results in the three
presented CSEM cases than other filters, and is included in \texttt{empymod}
from version 1.4.5 onwards. However, as with any digital filter, the quality
depends heavily on the model, and this new filter might or might not behave
that well for other models.

The shown GPR result shows that the digital linear filter method can also be
used for wave phenomena, not only for the diffusive approximation limit of low
frequency EM modeling.

We hope the presented algorithm is useful for at least 3 scenarios:
\begin{enumerate}
  \item Provide a fast method to design problem-specific filters. For bigger
    inversion projects (such as, for instance, massive stochastic CSEM
    inversions) a purpose designed, short filter might save much more time than
    it costs to design it. This could even be integrated into inversion codes,
    as an optional pre-inversion step.
  \item Extend the filter method to new areas, namely higher frequencies.
  \item Raise a new interest for digital linear filters in geophysics, by
    making it very easy for anyone to play around, create their own filters,
    and get a better understanding of it. We are sure there are many great
    filters to be discovered.
\end{enumerate}

\section{Acknowledgment}
D.W. would like to thank his colleague Oleg Gurevich Titov, who introduced him
to linear digital filters in the first place.
% This project would not have been successful without the fruitful discussions
% with Evert Slob and Kerry Key. Many thanks as well to Kerry Key for providing
% me with his Matlab scripts, which served as a starting point and initiated
% this project.

% REFERENCES
\bibliographystyle{seg}
\bibliography{\string~/.dotfiles/LaTeX/dtrRef}

\end{document}

### 2017-12-24 ManuscriptGeophysics2018.tex
Draft V2 of manuscript

@@
   filters shows that it performs better for some standard CSEM cases. The
-  derivation of longer 2001\,pt filter for a GPR example with 250\,MHz centre
+  derivation of a longer 2001\,pt filter for a GPR example with 250\,MHz centre
   frequency proves that the filter method works also for wave phenomena, not
@@
 further improved by many authors \citep{GP.82.Guptasarma, GEO.1982.Murakami,
-GP.97.Guptasarma}. \cite{GP.07.Kong} proposes a direct matrix inversion method
+GP.97.Guptasarma}. \cite{GP.07.Kong} proposed a direct matrix inversion method
 to solve the convolution equation, which requires only the input and output
 sample values. To evaluate the filters he defines the criteria of a good filter
-as one that recovers small or weak wavefields. This method was also used by
-\cite{GEO.09.Key, GEO.12.Key}. Most authors publish filters for the Hankel
-transform with $J_0$ and $J_1$ Bessel functions (or $J_{-1/2}$, $J_{1/2}$ if
-applied to the Fourier sine/cosine transform), as all higher Bessel functions
-can be rewritten to only use these two. \cite{GP.94.Mohsen} is one of the rare
-cases which provides $J_2$ filter weights.
+as one that recovers small or weak diffusive EM fields. This method was also
+used by \cite{GEO.09.Key, GEO.12.Key}. Most authors publish filters for the
+Hankel transform with $J_0$ and $J_1$ Bessel functions (or $J_{-1/2}$,
+$J_{1/2}$ if applied to the Fourier sine/cosine transform), as all higher
+Bessel functions can be rewritten to only use these two. \cite{GP.94.Mohsen} is
+one of the rare cases which provides $J_2$ filter weights.
 
@@
 methods (TEM), or controlled-source electromagnetics (CSEM), but not for high
-frequency methods such as ground-penetrating radar (GPR). Generally it was even
-thought that the filter method works only in the diffusive limit
+frequency methods such as ground-pe\-ne\-tra\-ting radar (GPR). Generally it
+was even thought that the filter method works only in the diffusive limit
 \citep[e.g.,][]{GEO.15.Hunziker}. Also, there is no open-source filter
@@
   \item The algorithm computes many different filters for various spacing and
-    shift values (brute-force).
+    shift values (brute-force), with an optional minimization routine for
+    the best outcome of the brute-force step.
   \item \cite{GP.07.Kong, GEO.12.Key} optimize a filter for the $J_0$ filter,
@@
     systems.
-  \item the real or the imaginary part can be used for the inversion of complex
+  \item The real or the imaginary part can be used for the inversion of complex
     transform pairs.
@@
 digital filter method. In this article we focus on the algorithm, and summarize
-the theory only very briefly by following \cite{GEO.12.Key} (equations
-(2)-(6)). In electromagnetics we often have to evaluate integrals of the form
+the theory only very briefly by following \cite{GEO.12.Key}. In
+electromagnetics we often have to evaluate integrals of the form
 %
@@
 From equation~\ref{eq:filtapprox} it can be seen that the filter method
-requires $N$ evaluations at each $r$. So to calculate for instance the
-frequency domain result for 100 offsets with a 201\,pt filter requires 20'100
-evaluations in the wavenumber domain. This is why the DLF often uses
-interpolation to minimise the required evaluations, either in the right-hand
-side or in the left-hand side.
+requires $N$ evaluations at each $r$. To calculate for instance the frequency
+domain result for 100 offsets with a 201\,pt filter requires 20'100 evaluations
+in the wavenumber domain. This is why the DLF often uses interpolation to
+minimise the required evaluations, either in the right-hand side (often called
+\emph{lagged convolution DLF}) or in the left-hand side (\emph{splined DLF}).
 
@@
 The filter length in this case is $N=201$, and the theoretical transform pairs
-is (as given, e.g., in \cite{USGS.75.Anderson})
+are (as given, e.g., in \cite{USGS.75.Anderson})
 %
@@
 filter length, and obviously heavily on the spacing and shift values, as this
-is what we invert for. Another important point is how you define the right hand
+is what we invert for. Another important point is how you define the right-hand
 evaluation points of the inversion ($r_\mr{def}$). Evaluating corresponding
@@
 coefficients ($J_0$, $J_1$, or $J_0$ \& $J_1$, or equally sine, cosine, or sine
-\& cosine); if the real or the imaginary part is used when complex transform
-pairs or \texttt{empymod} is used; and if it is inverted for the minimum
-amplitude or for the maximum $r$.
+\& cosine); also if the real or the imaginary part is used when complex
+transform pairs or \texttt{empymod} is used; and if it is inverted for the
+minimum amplitude or for the maximum $r$.
 
@@
 
-Figures \ref{fig:kongf5} compares the derived 201\,pt filter with the two
+Figure~\ref{fig:kongf5} compares the derived 201\,pt filter with the two
 half-spaces case used by \cite{GP.07.Kong} in his figure 5. The model consists
@@
   surface. The new filter has generally relative errors which are several
-  magnitudes lower than the other filters.}
+  orders of magnitude lower than the other filters.}
 %
-Figure \ref{fig:keyf5} (b) is a land case, with a background resistivity of
-$\rho_\mr{b}=10\,\Omega\,$m, in which a target is embedded of
+Figure \ref{fig:keyf5} (b) is a land case, with a background resistivity
+of $\rho_\mr{b}=10\,\Omega\,$m, in which a target is embedded of
 $\rho_\mr{t}=500\,\Omega\,$m at 1000\,m below the surface, 100\,m thick. Source
 depth is 0.5\,m, receiver depth is 0.8\,m. In both cases the new filter
-\emph{Wer201} has generally a relative error which is magnitudes lower than the
-other filters.
+\emph{Wer201} has generally a relative error which is orders of magnitude
+lower than the other filters. The relative error of the real part is given
+in black, whereas the relative error of the imaginary part is given in grey.
+It is interesting to note that the real and imaginary parts have very similar
+errors in \emph{Wer201}, \emph{Key201}, and \emph{And801}, but that the
+imaginary part of \emph{Kong241} seems to behave considerably better for most
+part than its real counterpart. To compare the real and the imaginary parts is
+insofar interesting as the digital filters are purely real valued.
 
@@
 \subsection{Ground-penetrating radar}
-
-\rmk{TODO: Add second GPR result; layers swapped}
 
@@
 to 2048\,MHz for both the calculation with the adaptive quadrature (QUA) and
-with the DLF. The frequency result is convolved with a Ricker wavelet with a
+with the DLF. The frequency result is multiplied with a Ricker wavelet with a
 center frequency of 250\,MHz, and a gain function ($1 + |t^3|, t$ in ns) is
 applied. The calculation with the digital filter took under 9 minutes and is
-therefore roughly 80 times faster then the adaptive quadrature calculation
+therefore roughly 80 times faster than the adaptive quadrature calculation
 which took roughly 11 hours and 27 minutes. Calculating the same model with QWE
@@
   quadrature took about 11\,h 27\,min to calculate, whereas the filter took
-  2\,min 10\,s.}
+  2\,min 10\,s in parallel on 4 nodes, hence under 9\,min in total.}
 %
@@
 \plot*{ftGPR}{width=\textwidth}{The real-part frequency-domain and the
-time-domain responses for offsets of (a) 0.2\,m, (b) 2.0\,m, and (c) 3.0\,m.}
+time-domain responses for offsets of (a/d) 0.2\,m, (b/e) 2.0\,m, and (c/f)
+3.0\,m.}
 %
+
+The 2001\,pt filter was derived with the fullspace solution of a medium which
+is very similar to the layer in which source and receiver reside in our GPR
+example. To show that this filter can also be applied to different layer
+parameters we run a test where we swapped layers one and two of the model
+in Figure \ref{fig:GPR} (a). The resulting real and imaginary parts for the
+offsets 0.2\,m, 2.0\,m, and 3.0\,m are given in Figure~\ref{fig:ftGPR2nd}.
+
+%
+\plot*{ftGPR2nd}{width=\textwidth}{The real and imaginary parts of the
+  frequency-domain response for the model given in Figure~\ref{fig:GPR} (a)
+  with swapped first and second layers for offsets of (a) 0.2\,m, (b) 2.0\,m,
+  and (c) 3.0\,m.}
+%
+
 These examples show clearly that the filter method can indeed be applied to
@@
 triple the calculation cost, the second idea double them; however, they both
-would still be very fast compared to standard quadrature methods.
+would still be very fast compared to standard quadrature methods. One could
+also apply the digital filter method to acoustic and elastic wavefields as
+well.
 
@@
 
-The presented, free and open-source algorithm \texttt{fdesign} can be used to
-design digital linear filters for the Hankel and Fourier transforms (and more
+The presented, free and open-source code \texttt{fdesign} can be used to design
+digital linear filters for the Hankel and Fourier transforms (and more
 generally for any linear transform) using either analytical transform pairs or
-1D subsurface models together with the EM-modeler \texttt{empymod}.
-The code is available from GitHub in the add-on repository of \texttt{empymod},
-called \texttt{empyscripts}, from version 0.1.2 onwards.
+1D subsurface models together with the EM-modeler \texttt{empymod}. The code is
+available from GitHub in the add-on repository of \texttt{empymod}, called
+\texttt{empyscripts}, from version 0.1.2 onwards.
 
@@
 
-We hope the presented algorithm is useful for at least 3 scenarios:
+We except the presented algorithm is useful for at least 3 scenarios:
 \begin{enumerate}
@@
     as an optional pre-inversion step.
-  \item Extend the filter method to new areas, namely higher frequencies.
+  \item Extend the filter method to new areas, namely higher frequencies and
+    acoustic and elastic wavefields.
   \item Raise a new interest for digital linear filters in geophysics, by
@@
 D.W. would like to thank his colleague Oleg Gurevich Titov, who introduced him
-to linear digital filters in the first place.
+to linear digital filters in the first place, and Jürg Hunziker for valuable
+feedback which improved this manuscript considerably.
 % This project would not have been successful without the fruitful discussions
@@
 % REFERENCES
-\bibliographystyle{seg}
-\bibliography{\string~/.dotfiles/LaTeX/dtrRef}
+\begin{thebibliography}{}
+\itemsep0pt
+
+\bibitem[Anderson, 1973]{USGS.73.Anderson}
+Anderson, W.~L.,  1973, Fortran {IV} programs for the determination of the
+  transient tangential electric field and vertical magnetic field about a
+  vertical magnetic dipole for an m-layer stratified earth by numerical
+  integration and digital linear filtering: USGS, {\bf PB221240}.
+
+\bibitem[Anderson, 1975]{USGS.75.Anderson}
+--------, 1975, Improved digital filters for evaluating {F}ourier and {H}ankel
+  transform integrals: USGS, {\bf PB242800}.
+\newblock
+  (\href{https://pubs.er.usgs.gov/publication/70045426}{https://pubs.er.usgs.gov/publication/70045426}).
+
+\bibitem[Anderson, 1979]{GEO.79.Anderson}
+--------, 1979, Numerical integration of related {H}ankel transforms of orders
+  0 and 1 by adaptive digital filtering: Geophysics, {\bf 44}, 1287--1305.
+\newblock (\href{http://doi.org/10.1190/1.1441007}{doi: 10.1190/1.1441007}).
+
+\bibitem[Anderson, 1982]{TMS.82.Anderson}
+--------, 1982, Fast {H}ankel transforms using related and lagged convolutions:
+  ACM Trans. Math. Softw., {\bf 8}, 344--368.
+\newblock (\href{http://doi.acm.org/10.1145/356012.356014}{doi:
+  10.1145/356012.356014}).
+
+\bibitem[Anderson, 1984]{GEO.84.Anderson}
+--------, 1984, {On: “Numerical integration of related Hankel transforms by
+  quadrature and continued fraction expansion” by \cite{GEO.83.Chave}}:
+  Geophysics, {\bf 49}, 1811--1812.
+\newblock (\href{http://doi.org/10.1190/1.1441595}{doi: 10.1190/1.1441595}).
+
+\bibitem[Anderson, 1989]{GEO.89.Anderson}
+--------, 1989, A hybrid fast {H}ankel transform algorithm for electromagnetic
+  modeling: Geophysics, {\bf 54}, 263--266.
+\newblock (\href{http://doi.org/10.1190/1.1442650}{doi: 10.1190/1.1442650}).
+
+\bibitem[Bernabini and Cardarelli, 1978]{GP.78.Bernabini}
+Bernabini, M., and E. Cardarelli,  1978, The use of filtered {B}essel functions
+  in direct interpretation of geoelectrical soundings: Geophysical Prospecting,
+  {\bf 26}, 841--852.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1978.tb01636.x}{doi:
+  10.1111/j.1365-2478.1978.tb01636.x}).
+
+\bibitem[Bichara and Lakshmanan, 1976]{GP.76.Bichara}
+Bichara, M., and J. Lakshmanan,  1976, Fast automatic processing of resistivity
+  soundings*: Geophysical Prospecting, {\bf 24}, 354--370.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1976.tb00932.x}{doi:
+  10.1111/j.1365-2478.1976.tb00932.x}).
+
+\bibitem[Chave, 1983]{GEO.83.Chave}
+Chave, A.~D.,  1983, Numerical integration of related {H}ankel transforms by
+  quadrature and continued fraction expansion: Geophysics, {\bf 48},
+  1671--1686.
+\newblock (\href{http://doi.org/10.1190/1.1441448}{doi: 10.1190/1.1441448}).
+
+\bibitem[Christensen, 1990]{GP.90.Christensen}
+Christensen, N.~B.,  1990, Optimized fast {H}ankel transform filters:
+  Geophysical Prospecting, {\bf 38}, 545--568.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1990.tb01861.x}{doi:
+  10.1111/j.1365-2478.1990.tb01861.x}).
+
+\bibitem[Das, 1982]{GEO.82.Das}
+Das, U.~C.,  1982, Designing digital linear filters for computing resistivity
+  and electromagnetic sounding curves: Geophysics, {\bf 47}, 1456--1459.
+\newblock (\href{https://doi.org/10.1190/1.1441295}{doi: 10.1190/1.1441295}).
+
+\bibitem[Das, 1984]{GEO.84.Das}
+--------, 1984, A single digital linear filter for computations in electrical
+  methods--a unifying approach: Geophysics, {\bf 49}, 1115--1118.
+\newblock (\href{https://doi.org/10.1190/1.1441726}{doi: 10.1190/1.1441726}).
+
+\bibitem[Das and Ghosh, 1974]{GP.74.Das}
+Das, U.~C., and D.~P. Ghosh,  1974, The determination of filter coefficients
+  for the computation of standard curves for dipole resistivity sounding over
+  layered earth by linear digital filtering: Geophysical Prospecting, {\bf 22},
+  765--780.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1974.tb00117.x}{doi:
+  10.1111/j.1365-2478.1974.tb00117.x}).
+
+\bibitem[Das et~al., 1974]{GP.74.Dasb}
+Das, U.~C., D.~P. Ghosh, and D.~T. Biewinga,  1974, Transformation of dipole
+  resistivity sounding measurements over layered earth by linear digital
+  filtering: Geophysical Prospecting, {\bf 22}, 476--489.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1974.tb00100.x}{doi:
+  10.1111/j.1365-2478.1974.tb00100.x}).
+
+\bibitem[Das and Verma, 1980]{GP.80.Das}
+Das, U.~C., and S.~K. Verma,  1980, Digital linear filter for computing type
+  curves for the two-electrode system of resistivity sounding: Geophysical
+  Prospecting, {\bf 28}, 610--619.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1980.tb01246.x}{doi:
+  10.1111/j.1365-2478.1980.tb01246.x}).
+
+\bibitem[Das and Verma, 1981a]{GJI.81.Das}
+--------, 1981a, Numerical considerations on computing the {EM} response of
+  three-dimensional inhomogeneities in a layered earth: Geophysical Journal
+  International, {\bf 66}, 733--740.
+\newblock (\href{http://doi.org/10.1111/j.1365-246X.1981.tb04897.x}{doi:
+  10.1111/j.1365-246X.1981.tb04897.x}).
+
+\bibitem[Das and Verma, 1981b]{GXP.81.Das}
+--------, 1981b, The versatility of digital linear filters used in computing
+  resistivity and em sounding curves: Geoexploration, {\bf 18}, 297--310.
+\newblock (\href{http://doi.org/10.1016/0016-7142(81)90059-4}{doi:
+  10.1016/0016-7142(81)90059-4}).
+
+\bibitem[Das and Verma, 1982]{GJI.82.Das}
+--------, 1982, Electromagnetic response of an arbitrarily shaped
+  three-dimensional conductor in a layered earth — numerical results:
+  Geophysical Journal International, {\bf 69}, 55--66.
+\newblock (\href{http://doi.org/10.1111/j.1365-246X.1982.tb04935.x}{doi:
+  10.1111/j.1365-246X.1982.tb04935.x}).
+
+\bibitem[Davis et~al., 1980]{EXG.80.Davis}
+Davis, P.~A., S.~A. Greenhalgh, and N.~P. Merrick,  1980, Resistivity sounding
+  computations with any array using a single digital filter: Exploration
+  Geophysics, {\bf 11}, 54--62.
+\newblock (\href{http://library.seg.org/doi/abs/10.1071/EG980054}{doi:
+  10.1071/EG980054}).
+
+\bibitem[Ghosh, 1970]{PhD.70.Ghosh}
+Ghosh, D.~P.,  1970, The application of linear filter theory to the direct
+  interpretation of geoelectrical resistivity measurements: {Ph.D. Thesis}, {TU
+  Delft}.
+\newblock
+  (\href{http://resolver.tudelft.nl/uuid:88a568bb-ebee-4d7b-92df-6639b42da2b2}{uuid:
+  88a568bb-ebee-4d7b-92df-6639b42da2b2}).
+
+\bibitem[Ghosh, 1971a]{GP.71.Ghosh}
+--------, 1971a, The application of linear filter theory to the direct
+  interpretation of geoelectrical resistivity sounding measurements:
+  Geophysical Prospecting, {\bf 19}, 192--217.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1971.tb00593.x}{doi:
+  10.1111/j.1365-2478.1971.tb00593.x}).
+
+\bibitem[Ghosh, 1971b]{GP.71.Ghoshb}
+--------, 1971b, Inverse filter coefficients for the computation of apparent
+  resistivity standard curves for a horizontally stratified earth: Geophysical
+  Prospecting, {\bf 19}, 769--775.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1971.tb00915.x}{doi:
+  10.1111/j.1365-2478.1971.tb00915.x}).
+
+\bibitem[Guptasarma, 1982]{GP.82.Guptasarma}
+Guptasarma, D.,  1982, Optimization of short digital linear filters for
+  increased accuracy: Geophysical Prospecting, {\bf 30}, 501--514.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1982.tb01320.x}{doi:
+  10.1111/j.1365-2478.1982.tb01320.x}).
+
+\bibitem[Guptasarma and Singh, 1997]{GP.97.Guptasarma}
+Guptasarma, D., and B. Singh,  1997, New digital linear filters for {H}ankel
+  {J}0 and {J}1 transforms: Geophysical Prospecting, {\bf 45}, 745--762.
+\newblock (\href{http://doi.org/10.1046/j.1365-2478.1997.500292.x}{doi:
+  10.1046/10.1046/j.1365-2478.1997.500292.x}).
+
+\bibitem[Hunziker et~al., 2015]{GEO.15.Hunziker}
+Hunziker, J., J. Thorbecke, and E. Slob,  2015, The electromagnetic response in
+  a layered vertical transverse isotropic medium: {A} new look at an old
+  problem: Geophysics, {\bf 80}, no. 1, F1--F18.
+\newblock (\href{http://doi.org/10.1190/geo2013-0411.1}{doi:
+  10.1190/geo2013-0411.1}).
+
+\bibitem[Johansen, 1975]{GP.75.Johansen}
+Johansen, H.~K.,  1975, An interactive computer/graphic-display-terminal system
+  for interpretation of resistivity soundings: Geophysical Prospecting, {\bf
+  23}, 449--458.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1975.tb01541.x}{doi:
+  10.1111/j.1365-2478.1975.tb01541.x}).
+
+\bibitem[Johansen and Sørensen, 1979]{GP.79.Johansen}
+Johansen, H.~K., and K. Sørensen,  1979, Fast {H}ankel transforms: Geophysical
+  Prospecting, {\bf 27}, 876--901.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1979.tb01005.x}{doi:
+  10.1111/j.1365-2478.1979.tb01005.x}).
+
+\bibitem[Key, 2009]{GEO.09.Key}
+Key, K.,  2009, {1D} inversion of multicomponent, multifrequency marine {CSEM}
+  data: {M}ethodology and synthetic studies for resolving thin resistive
+  layers: Geophysics, {\bf 74}, no. 2, F9--F20.
+\newblock (\href{http://doi.org/10.1190/1.3058434}{doi: 10.1190/1.3058434}).
+
+\bibitem[Key, 2012]{GEO.12.Key}
+--------, 2012, Is the fast {H}ankel transform faster than quadrature?:
+  Geophysics, {\bf 77}, no. 3, F21--F30.
+\newblock (\href{http://doi.org/10.1190/GEO2011-0237.1}{doi:
+  10.1190/GEO2011-0237.1}).
+
+\bibitem[Koefoed, 1968]{BK.68.Koefoed}
+Koefoed, O.,  1968, The application of the kernel function in interpreting
+  geoelectrical resistivity measurements: Gebrüder Borntraeger.
+\newblock ({I}SBN: 9783443130022).
+
+\bibitem[Koefoed, 1970]{GP.70.Koefoed}
+--------, 1970, A fast method for determining the layer distribution from the
+  raised kernel function in geoelectrical sounding: Geophysical Prospecting,
+  {\bf 18}, 564--570.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1970.tb02129.x}{doi:
+  10.1111/j.1365-2478.1970.tb02129.x}).
+
+\bibitem[Koefoed, 1972]{GP.72.Koefoedb}
+--------, 1972, A note on the linear filter method of interpreting resistivity
+  sounding data: Geophysical Prospecting, {\bf 20}, 403--405.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1972.tb00643.x}{doi:
+  10.1111/j.1365-2478.1972.tb00643.x}).
+
+\bibitem[Koefoed, 1976]{GP.76.Koefoedb}
+--------, 1976, Error propagation and uncertainty in the interpretation of
+  resistivity sounding data: Geophysical Prospecting, {\bf 24}, 31--48.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1976.tb00383.x}{doi:
+  10.1111/j.1365-2478.1976.tb00383.x}).
+
+\bibitem[Koefoed and Dirks, 1979]{GP.79.Koefoed}
+Koefoed, O., and F.~J.~H. Dirks,  1979, Determination of resistivity sounding
+  filters by the {W}iener-{H}opf least-squares method: Geophysical Prospecting,
+  {\bf 27}, 245--250.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1979.tb00968.x}{doi:
+  10.1111/j.1365-2478.1979.tb00968.x}).
+
+\bibitem[Koefoed et~al., 1972]{GP.72.Koefoed}
+Koefoed, O., D.~P. Ghosh, and G.~J. Polman,  1972, Computation of type curves
+  for electromagnetic depth sounding with a horizontal transmitting coil by
+  means of a digital linear filter: Geophysical Prospecting, {\bf 20},
+  406--420.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1972.tb00644.x}{doi:
+  10.1111/j.1365-2478.1972.tb00644.x}).
+
+\bibitem[Kong, 2007]{GP.07.Kong}
+Kong, F.~N.,  2007, Hankel transform filters for dipole antenna radiation in a
+  conductive medium: Geophysical Prospecting, {\bf 55}, 83--89.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.2006.00585.x}{doi:
+  10.1111/j.1365-2478.2006.00585.x}).
+
+\bibitem[Kruglyakov and Bloshanskaya, 2017]{MGS.17.Kruglyakov}
+Kruglyakov, M., and L. Bloshanskaya,  2017, High-performance parallel solver
+  for integral equations of electromagnetics based on {G}alerkin method:
+  Mathematical Geosciences, {\bf 49}, 751--776.
+\newblock (\href{https://doi.org/10.1007/s11004-017-9677-y}{doi:
+  10.1007/s11004-017-9677-y}).
+
+\bibitem[Kumar and Das, 1977]{GP.77.Kumar}
+Kumar, R., and U.~C. Das,  1977, Transformation of dipole to {S}chlumberger
+  sounding curves by means of digital linear filters: Geophysical Prospecting,
+  {\bf 25}, 780--789.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1977.tb01204.x}{doi:
+  10.1111/j.1365-2478.1977.tb01204.x}).
+
+\bibitem[Kumar and Das, 1978]{GP.78.Kumar}
+--------, 1978, Transformation of {S}chlumberger apparent resistivity to dipole
+  apparent resistivity over layered earth by the application of digital linear
+  filters: Geophysical Prospecting, {\bf 26}, 352--358.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1978.tb01598.x}{doi:
+  10.1111/j.1365-2478.1978.tb01598.x}).
+
+\bibitem[Mohsen and Hashish, 1994]{GP.94.Mohsen}
+Mohsen, A.~A., and E.~A. Hashish,  1994, The fast {H}ankel transform:
+  Geophysical Prospecting, {\bf 42}, 131--139.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1994.tb00202.x}{doi:
+  10.1111/j.1365-2478.1994.tb00202.x}).
+
+\bibitem[Murakami and Uchida, 1982]{GEO.1982.Murakami}
+Murakami, Y., and T. Uchida,  1982, Accuracy of the linear filter coefficients
+  determined by the iteration of the least-squares method: Geophysics, {\bf
+  47}, 244--256.
+\newblock (\href{https://doi.org/10.1190/1.1441331}{doi: 10.1190/1.1441331}).
+
+\bibitem[Newman et~al., 1986]{GEO.86.Newman}
+Newman, G.~A., G.~W. Hohmann, and W.~L. Anderson,  1986, Transient
+  electromagnetic response of a three-dimensional body in a layered earth:
+  Geophysics, {\bf 51}, 1608--1627.
+\newblock (\href{http://doi.org/10.1190/1.1442212}{doi: 10.1190/1.1442212}).
+
+\bibitem[Nissen and Enmark, 1986]{GP.86.Nissen}
+Nissen, J., and T. Enmark,  1986, An optimized digital filter for the {F}ourier
+  transform: Geophysical Prospecting, {\bf 34}, 897--903.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1986.tb00500.x}{doi:
+  10.1111/j.1365-2478.1986.tb00500.x}).
+
+\bibitem[Niwas, 1975]{GEO.75.Niwas}
+Niwas, S.,  1975, Direct interpretation of geoelectric measurements by the use
+  of linear filter theory: Geophysics, {\bf 40}, 121--122.
+\newblock (\href{https://doi.org/10.1190/1.1440510}{doi: 10.1190/1.1440510}).
+
+\bibitem[Nyman and Landisman, 1977]{GEO.77.Nyman}
+Nyman, D.~C., and M. Landisman,  1977, {VES} dipole-dipole filter coefficients:
+  Geophysics, {\bf 42}, 1037--1044.
+\newblock (\href{https://doi.org/10.1190/1.1440763}{doi: 10.1190/1.1440763}).
+
+\bibitem[O'Neill, 1975]{EXG.75.ONeill}
+O'Neill, D.,  1975, Improved linear filter coefficients for application in
+  apparent resistivity computations: Exploration Geophysics, {\bf 6}, 104--109.
+\newblock (\href{http://library.seg.org/doi/abs/10.1071/EG975104}{doi:
+  10.1071/EG975104}).
+
+\bibitem[O'Neill and Merrick, 1984]{GP.84.ONeill}
+O'Neill, D.~J., and N.~P. Merrick,  1984, A digital linear filter for
+  resistivity sounding with a generalized electrode array: Geophysical
+  Prospecting, {\bf 32}, 105--123.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1984.tb00720.x}{doi:
+  10.1111/j.1365-2478.1984.tb00720.x}).
+
+\bibitem[Pekeris, 1940]{GEO.40.Pekeris}
+Pekeris, C.~L.,  1940, Direct method of interpretation in resistivity
+  prospecting: Geophysics, {\bf 5}, 31--42.
+\newblock (\href{https://doi.org/10.1190/1.1441791}{doi: 10.1190/1.1441791}).
+
+\bibitem[Raiche et~al., 2007]{ASEG.07.Raiche}
+Raiche, A., F. Sugeng, and G. Wilson,  2007, Practical {3D} {EM} inversion -
+  the {P223F} software suite: ASEG Technical Program Expanded Abstracts,  1--4.
+\newblock (\href{http://doi.org/10.1071/ASEG2007ab114}{doi:
+  10.1071/ASEG2007ab114}).
+
+\bibitem[Slichter, 1933]{PHY.33.Slichter}
+Slichter, L.~B.,  1933, The interpretation of the resistivity prospecting
+  method for horizontal structures: Physics, {\bf 4}, 307--322.
+\newblock (\href{http://doi.org/10.1063/1.1745198}{doi : 10.1063/1.1745198}).
+
+\bibitem[Sørensen and Christensen, 1994]{GEO.94.Sorensen}
+Sørensen, K.~I., and N.~B. Christensen,  1994, The fields from a finite
+  electrical dipole—a new computational approach: Geophysics, {\bf 59},
+  864--880.
+\newblock (\href{https://doi.org/10.1190/1.1443646}{10.1190/1.1443646}).
+
+\bibitem[Verma and Koefoed, 1973]{GP.73.Verma}
+Verma, R.~K., and O. Koefoed,  1973, A note on the linear filter method of
+  computing electromagnetic sounding curves: Geophysical Prospecting, {\bf 21},
+  70--76.
+\newblock (\href{http://doi.org/10.1111/j.1365-2478.1973.tb00015.x}{doi:
+  10.1111/j.1365-2478.1973.tb00015.x}).
+
+\bibitem[Werthmüller, 2017]{GEO.17.Werthmuller}
+Werthmüller, D.,  2017, An open-source full {3D} electromagnetic modeler for
+  {1D} {VTI} media in {P}ython: empymod: Geophysics, {\bf 82}, no. 6,
+  WB9--WB19.
+\newblock (\href{http://doi.org/10.1190/geo2016-0626.1}{doi:
+  10.1190/geo2016-0626.1}).
+
+\end{thebibliography}
 

### 2017-12-29 ManuscriptGeophysics2018.tex
Well then
Make intro (1/2/3)-titles clearer

@@
 
-% \subsection{Review}
-
 The introduction of linear filters to electromagnetic geophysics initiated a
@@
 in the 70s and 80s by Ghosh, Koefoed, Das, Anderson, and many others, the vast
-majority of it in the journal \emph{Geophysical Prospecting}.
-%  (see Figure~\ref{fig:authors})
-Another common name of DLF is \emph{fast Hankel transform} (FHT), popular
-because of the similarity of the name to the well known \emph{fast Fourier
-transform} (FFT). The name was introduced, as far as we can see, by
-\cite{GP.79.Johansen} in their article of the same name. However, the name FHT
-can be misleading as it has \emph{Hankel} in the name, but the DLF can also be
-applied to the Fourier transform (and other linear transforms).
-% \begin{figure}[tbp]
-%   \centering
-%   \fbox{\includegraphics[width=10cm]{figures/author-figure}}
-%     \caption{List by year of publications included in the review (year and
-%       first three letters of principal author).}
-%     \label{fig:authors}
-% \end{figure}
-Publications fall broadly into one or several of three categories: (1) new
-methods (application of DLF to new measurement techniques); (2) filter
-improvements (new filters or providing new or improved methods for the
-determination of filter coefficients); and (3) computational tools. The
-following is a brief review of the most relevant publications regarding DLF,
-without claiming completeness.
+majority of it in the journal \emph{Geophysical Prospecting}. Another common
+name of DLF is \emph{fast Hankel transform} (FHT), popular because of the
+similarity of the name to the well known \emph{fast Fourier transform} (FFT).
+The name was introduced, as far as we can see, by \cite{GP.79.Johansen} in
+their article of the same name. However, the name FHT can be misleading as it
+has \emph{Hankel} in the name, but the DLF can also be applied to the Fourier
+transform (and other linear transforms). Publications fall broadly into one or
+several of three categories: (1) new applications (extend the usage of DLF to
+other measurement techniques); (2) filter improvements (new filters or
+providing new or improved methods for the determination of filter
+coefficients); and (3) computational tools. The following is a brief review of
+the most relevant publications regarding DLF, without claiming completeness.
 
-\subsubsection{(1) New methods}
+\subsubsection{(1) New applications}
 
@@
 
-\subsubsection{(3) Codes}
+\subsubsection{(3) Computational tools}
 
@@
 
-% \subsection{Outline}
-
 All mentioned publications have in common that they were derived for direct
@@
 design general or purpose-built filters using the direct matrix inversion
-method. After a brief overview of the methodology, the theory, and the code we
-show examples of its usage, presenting some new filters for both, CSEM and GPR
-data. The algorithm and many more examples of its usage can be found on
+method. The current article falls therefore into all three mentioned categories
+by applying DLF also to wave phenomena and providing new filters with the
+accompanying code. After a brief overview of the methodology, the theory, and
+the code we show examples of its usage, presenting some new filters for both,
+CSEM and GPR data. The algorithm and many more examples of its usage can be
+found on
 \href{https://github.com/empymod/article-fdesign}{github.com/empymod/article-fdesign}.
@@
 \end{enumerate}
-
-\section{Acknowledgment}
-D.W. would like to thank his colleague Oleg Gurevich Titov, who introduced him
-to linear digital filters in the first place, and Jürg Hunziker for valuable
-feedback which improved this manuscript considerably.
-% This project would not have been successful without the fruitful discussions
-% with Evert Slob and Kerry Key. Many thanks as well to Kerry Key for providing
-% me with his Matlab scripts, which served as a starting point and initiated
-% this project.
+~
 

### 2018-01-16 ManuscriptGeophysics2018.tex
Start acknowledgment
Clean preamble
Adjust empyscript version
Include feeback KK

@@
 % \documentclass[manuscript,revised]{geophysics}
+
 %%fakesection ===    PACKAGES & DEFINITIONS    ===
@@
 %% ~ ADDITIONAL PACKAGES TO GEOPHYSICS.CLS
-\DeclareGraphicsExtensions{.pdf,.png,.jpg}
-\usepackage{color}
-\usepackage{tabularx}
-\usepackage{colortbl}
-\usepackage{booktabs}
 \usepackage[USenglish]{babel}
@@
 \usepackage{upquote}
+\usepackage[strings]{underscore}
 \usepackage[pdftex, final]{hyperref}
@@
 
-\usepackage{listings}
-
-% Figure directory
+% Figures
+\DeclareGraphicsExtensions{.pdf,.png,.jpg}
 \renewcommand{\figdir}{./figures}
+
+% Own commands
 \usepackage{xspace}
 \newcommand{\empymod}{\texttt{empymod}\xspace}
-\usepackage[strings]{underscore}
-
-\definecolor{MyGray}{gray}{0.85}
-\newcolumntype{Y}{>{\columncolor{MyGray}\raggedleft\arraybackslash}r}
-\newcolumntype{W}{>{\raggedleft\arraybackslash}r}
-\newcolumntype{A}{>{\columncolor{MyGray}\raggedleft\arraybackslash}X}
-\newcolumntype{B}{>{\raggedleft\arraybackslash}X}
-
-\hyphenation{iso-tro-pic pe-ne-tra-ting}
-
-\newcommand{\rmk}[1]{\textbf{\color{red}{#1}}}
 \newcommand{\mr}[1]{\mathrm{#1}}
-\frenchspacing
 
@@
 
-\title{A digital filter designing tool for the Hankel and Fourier transforms in
-potential-, diffusive-, and wavefield modeling}
+\title{A tool for designing digital filters for the Hankel and Fourier transforms in
+potential, diffusive, and wavefield modeling}
 
@@
 E-mail: \href{mailto:dieter@werthmuller.org}{Dieter@Werthmuller.org};
-\footnotemark[2]Lamont-Doherty Earth Observatory,
+\footnotemark[2]Lamont-Doherty Earth Observatory, Columbia University,
 305C Oceanography,
@@
 
-Deva Prasad Ghosh proposed in his Ph.D. thesis \citep{PhD.70.Ghosh} a linear
-filter method for the Hankel transform which revolutionized the computational
-modeling of electromagnetic responses in the field of geophysical exploration.
-If you use a code that calculates electromagnetic responses in the
-wavenumber-frequency domain and transforms them to the space-frequency domain
-chances are very high that it uses the \emph{digital linear filter} method
-(DLF). \cite{GP.71.Ghosh} states that the idea is based on suggestions made
-four decades earlier by \cite{PHY.33.Slichter} and \cite{GEO.40.Pekeris}, in
-that ``\emph{the kernel function is dependent only on the layer parameters, and
-  an expression relating it to the field measurements can be obtained by
-mathematical processes.}'' However, until the introduction of DLF by Ghosh
-these suggestions found no widespread use, probably not last because of the
-missing computer power to calculate the filter coefficients. He further states
-that credit goes to his Ph.D. supervisor Otto \cite{BK.68.Koefoed,
-GP.70.Koefoed}, who retook the task of direct interpretational methods with the
-introduction of raised kernel functions. DLF is, as such, an improvement of
-that approach, providing a faster and simpler method.
-
-The introduction of linear filters to electromagnetic geophysics initiated a
-wealth of investigation, publications, and development of computer programs
-that extended and improved the method. A few dozen articles have been published
-in the 70s and 80s by Ghosh, Koefoed, Das, Anderson, and many others, the vast
-majority of it in the journal \emph{Geophysical Prospecting}. Another common
-name of DLF is \emph{fast Hankel transform} (FHT), popular because of the
-similarity of the name to the well known \emph{fast Fourier transform} (FFT).
-The name was introduced, as far as we can see, by \cite{GP.79.Johansen} in
-their article of the same name. However, the name FHT can be misleading as it
-has \emph{Hankel} in the name, but the DLF can also be applied to the Fourier
-transform (and other linear transforms). Publications fall broadly into one or
-several of three categories: (1) new applications (extend the usage of DLF to
-other measurement techniques); (2) filter improvements (new filters or
-providing new or improved methods for the determination of filter
-coefficients); and (3) computational tools. The following is a brief review of
-the most relevant publications regarding DLF, without claiming completeness.
+In his Ph.D. thesis, \cite{PhD.70.Ghosh} proposed a linear filter method for
+the numerical evaluation of Hankel transforms that subsequently revolutionized
+the computation of electromagnetic (EM) responses in the field of geophysical
+exploration. If you use a code that calculates EM responses in the
+wavenumber-frequency domain and transforms them to the space-frequency domain,
+chances are high that it uses the \emph{digital linear filter} (DLF) method.
+Most practical 1D EM modeling codes rely on the DLF method for rapid
+computations; these codes are not only used for standalone simulations of EM
+fields in layered 1D media, but they are also commonly embedded within 3D
+modeling codes for computing the primary fields in scattered-field formulations
+or for the background fields required by integral equation methods. Thus, the
+DLF method is an important component of many commonly used modeling codes for
+EM geophysical data.
+
+\cite{GP.71.Ghosh} states that the DLF idea is based on suggestions made four
+decades earlier by \cite{PHY.33.Slichter} and \cite{GEO.40.Pekeris}, in that
+``\emph{the kernel function is dependent only on the layer parameters, and an
+expression relating it to the field measurements can be obtained by
+mathematical processes.}'' However, until the introduction of DLF, these
+suggestions found no widespread use, likely because of the missing computer
+power to calculate the filter coefficients. He further states that credit goes
+to his Ph.D. supervisor Otto \cite{BK.68.Koefoed, GP.70.Koefoed}, who retook
+the task of direct interpretational methods with the introduction of raised
+kernel functions. DLF is, as such, an improvement of that approach, providing a
+faster and simpler method.
+
+The DLF technique is sometimes referred to as the \emph{fast Hankel transform}
+(FHT), popular because of the similarity of the name to the well known
+\emph{fast Fourier transform} (FFT), although the algorithms for these
+techniques are completely different. The FHT name was likely inspired by the
+title of a paper by \cite{GP.79.Johansen}. However, the name FHT can be
+misleading as it has \emph{Hankel} in the name, whereas the DLF approach can
+more generally be applied to other linear transforms, for example Fourier sine
+and cosine transforms.
+
+The introduction of linear filters to EM geophysics, in parallel with rising
+computational power, initiated a wealth of investigations, leading to the
+development of computer programs that extended and improved the DLF method, and
+to numerous publications. These publications fall broadly into one or several
+of three categories: (1) new applications that extend the usage of DLF to other
+EM measurement techniques; (2) filter improvements that provide either new
+filters or improved methods for the determination of filter coefficients; and
+(3) computational tools that compute EM responses using DLF techniques. Here,
+we briefly review the most relevant publications, without claiming
+completeness.
 
@@
 Ghosh used the method originally for the computation of type curves for
-Schlumberger and Wenner resistivity soundings. He published the main results of
-his thesis the following year in two publications: in \cite{GP.71.Ghosh} he
-derives a resistivity model from given Schlumberger or Wenner sounding curves;
-and in \cite{GP.71.Ghoshb} he provides filters for the inverse operation,
-deriving resistivity sounding curves from a given resistivity model.  The
-method was next applied to electromagnetic soundings with horizontal and
-perpendicular coils \citep{GP.72.Koefoed}, to vertical coplanar coil systems
-\citep{GP.73.Verma}, to dipoles and other two electrode systems
-\citep{GP.74.Das, GP.74.Dasb, GP.80.Das, GEO.94.Sorensen}, and to vertical
-dikes, hence vertical instead of horizontal layers \citep{GEO.75.Niwas}. Whilst
-the first filters were very specific to one type of curves and one type of
-transformation, various publications used the method to get one type curve from
-another type curve \citep{GP.77.Kumar, GP.78.Kumar} or generalized the method
-to be applicable to a wider set of problems \citep{EXG.80.Davis, GXP.81.Das,
-GEO.84.Das, GP.84.ONeill}. Eventually, it passed from pure layered modeling to
-primary-secondary formulations in 3D problems, where DLF is used to compute the
-spatial Fourier-Hankel transforms in a horizontally layered background medium
-and to compute transient responses from frequency domain computations
+Schlumberger and Wenner resistivity soundings: \cite{GP.71.Ghosh} derives a
+resistivity model from given Schlumberger or Wenner sounding curves; and
+\citet{GP.71.Ghoshb} provides filters for the inverse operation, deriving
+resistivity sounding curves from a given resistivity model. The method was next
+applied to electromagnetic soundings with horizontal and perpendicular coils
+\citep{GP.72.Koefoed}, to vertical coplanar coil systems \citep{GP.73.Verma},
+to dipoles and other two electrode systems \citep{GP.74.Das, GP.74.Dasb,
+GP.80.Das, GEO.94.Sorensen}, and to vertical dikes, hence vertical instead of
+horizontal layers \citep{GEO.75.Niwas}. The first filters were specific to a
+particular resistivity sounding type and its transform; later publications used
+the method to get one type curve from another type curve \citep{GP.77.Kumar,
+GP.78.Kumar}, or generalized the method to be applicable to a wider set of
+problems \citep{EXG.80.Davis, GXP.81.Das, GEO.84.Das, GP.84.ONeill}.
+Eventually, it passed from pure layered modeling to primary-secondary field
+formulations for 3D problems, where DLF is used to compute the spatial
+Fourier-Hankel transforms in a horizontally layered background medium and to
+compute transient responses from frequency domain computations
 \citep{GJI.81.Das, GJI.82.Das, GEO.84.Anderson, GEO.86.Newman,
-MGS.17.Kruglyakov}. Various authors delved into the theory of the method,
+MGS.17.Kruglyakov}. Other publications delved into the theory of the method,
 analyzing the oscillating behaviour of the filters and trying to estimate the
@@
 specifically for the Fourier transform by \cite{GP.86.Nissen}. A direct
-integration method was used by \cite{GP.76.Bichara, GP.78.Bernabini}.
+integration method was used by \cite{GP.76.Bichara} and \cite{GP.78.Bernabini}.
 \cite{GP.79.Koefoed} proposed a Wiener-Hopf least-squares method, which was
@@
 to solve the convolution equation, which requires only the input and output
-sample values. To evaluate the filters he defines the criteria of a good filter
+sample values. To evaluate the filter's effectiveness, he defines a good filter
 as one that recovers small or weak diffusive EM fields. This method was also
-used by \cite{GEO.09.Key, GEO.12.Key}. Most authors publish filters for the
-Hankel transform with $J_0$ and $J_1$ Bessel functions (or $J_{-1/2}$,
-$J_{1/2}$ if applied to the Fourier sine/cosine transform), as all higher
-Bessel functions can be rewritten to only use these two. \cite{GP.94.Mohsen} is
-one of the rare cases which provides $J_2$ filter weights.
+used by \cite{GEO.09.Key, GEO.12.Key} to create a suite of filters accurate for
+marine EM data. Most works have published filters for the Hankel transform with
+$J_0$ and $J_1$ Bessel functions (or $J_{-1/2}$, $J_{1/2}$ if applied to the
+Fourier sine/cosine transform), as all higher Bessel functions can be
+rewritten, via recurrence relations, using only use these two.
+\cite{GP.94.Mohsen} is one of the rare cases which provides $J_2$ filter
+weights.
 
@@
 
-The best known codes are likely the freely available ones by Anderson.
+The most well-known codes are likely Anderson's freely available ones.
 \cite{USGS.73.Anderson} extends the method to transient responses, applying DLF
@@
 transient signal can therefore be obtained be applying twice a digital filter
-to the wavenumber-frequency domain calculation. In \cite{USGS.75.Anderson,
-GEO.79.Anderson} he presents improved filters for both Fourier and Hankel
+to the wavenumber-frequency domain calculation. \cite{USGS.75.Anderson,
+GEO.79.Anderson} presents improved filters for both Fourier and Hankel
 transforms, introducing measures to significantly speed-up the calculation,
 such as the lagged convolution or using the same abscissae for $J_0$ and $J_1$.
-\cite{TMS.82.Anderson} and the included 801\,pt filter became sort of the
-industry standard, to which subsequent filters were compared. In
-\cite{GEO.89.Anderson} he presents a hybrid solution, permitting to use either
-DLF or quadrature and as such permits to compare the two. Other examples
-include the codes by \cite{GP.75.Johansen}, an interactive system for
-interpretation of resistivity soundings, and a tool to calculate filter
-coefficients by \cite{GP.90.Christensen}. The latter is available upon request
-and was used, for instance, in all the open-source modeling and inversion
-routines of CSIRO in the Amira Project 223 \citep{ASEG.07.Raiche}. \newline
+\cite{TMS.82.Anderson} and the included 801\,pt filter became the de facto
+industry standard, to which subsequent filters were compared.
+\cite{GEO.89.Anderson} presents a hybrid solution that uses either the DLF or a
+much slower but highly accurate adaptive-quadrature approach, which can also be
+used to measure the accuracy of the DLF. \cite{GEO.12.Key} presented codes for
+comparing the DLF approach with a speed optimized quadrature method call
+quadrature-with-extrapolation (QWE); despite its optimizations, QWE is still
+not as fast as DLF, but it remains valuable as an independent tool for testing
+the accuracy of particular filters for the DLF method. Other examples include
+the codes by \cite{GP.75.Johansen}, an interactive system for interpretation of
+resistivity soundings, and a tool to calculate filter coefficients by
+\cite{GP.90.Christensen}. The latter is available upon request and was used,
+for instance, in all the open-source modeling and inversion routines of CSIRO
+in the Amira Project 223 \citep{ASEG.07.Raiche}.
 
@@
 frequency methods such as ground-pe\-ne\-tra\-ting radar (GPR). Generally it
-was even thought that the filter method works only in the diffusive limit
-\citep[e.g.,][]{GEO.15.Hunziker}. Also, there is no open-source filter
-designing tool readily available, they are available as a described method
-in articles or as code upon request by the author, if at all. The presented
-algorithm \texttt{fdesign} tries to fill this gap by providing the tools to
-design general or purpose-built filters using the direct matrix inversion
-method. The current article falls therefore into all three mentioned categories
-by applying DLF also to wave phenomena and providing new filters with the
-accompanying code. After a brief overview of the methodology, the theory, and
-the code we show examples of its usage, presenting some new filters for both,
-CSEM and GPR data. The algorithm and many more examples of its usage can be
-found on
+was even thought that the filter method works only in the diffusive limit where
+the quasistatic approximation is valid \citep[e.g.,][]{GEO.15.Hunziker}.
+Nevertheless, \cite{GEO.17.Werthmuller} tested DLF for modeling 250 MHz
+center-frequency GPR data using a 401 point filter that was designed for
+diffusive EM, and found that it was orders of magnitude faster than quadrature
+approaches. However, good quality results were only obtained for the first
+arrivals at short time-offsets and the later arrivals had poor quality. Yet,
+this promising initial test suggested it might be possible to specifically
+design a special filter for GPR frequencies where the EM fields propagate as
+waves rather than by diffusion.
+
+While there are several versions of digital filters that are freely available,
+there are no open-source filter designing tools that readily available; they
+are either only described methods in articles or the codes are only available
+upon request by the author. Here, we address this gap by presenting the code
+\texttt{fdesign}, which is an open-source tool for designing general or
+purpose-built filters using the direct matrix inversion method. The code is
+completely open-source and can be run using freely available programming
+environments, thus there are few barriers to obtaining and running the code.
+
+In the following sections we give a brief overview of the methodology, the
+theory, and the code, and then show examples of its usage, as well as
+presenting new filters for both CSEM and GPR data. The current article falls
+therefore into all three mentioned categories by the new application of DLF to
+wave phenomena and by providing new filters with the accompanying code. The
+algorithm and many more examples of its usage can be found on
 \href{https://github.com/empymod/article-fdesign}{github.com/empymod/article-fdesign}.
@@
 matrix inversion method as described by \cite{GP.07.Kong} and based on scripts
-by \cite{GEO.12.Key}. The tool is an add-on to the electromagnetic modeler
+developed by \cite{GEO.12.Key}. The tool is an add-on to the EM modeler
 \texttt{empymod} \citep{GEO.17.Werthmuller}, written in Python, and hosted on
@@
 improve, and extend it. It can be used to derive digital linear filters for the
-Hankel and the Fourier transforms for potential-, diffusive-, and wavefields
-(hence from DC to GPR). Theoretically, it can be used to derive linear filters
-for any linear transform, as long as you feed it with a theoretical, or
-accurately computed, transform pair of the transform. It permits to derive
-filters with the help of theoretical transform pairs or, alternatively, with
-the electromagnetic modeler \texttt{empymod}, using a quadrature method to
-derive accurate curves used for the inversion.
-
-The main points of the method and some differences between it and the methods
-it is based upon are:
+Hankel and the Fourier transforms for potential, diffusive, and wavefield
+modeling (hence from DC to GPR).
+
+Theoretically, it can be used to derive linear filters for any linear
+transform, as long as it is supplied with the input and outputs of a transform
+pair. Previous works have relied on using theoretical transforms pairs commonly
+found in tables of integral transforms. Here we follow a similar approach, but
+we also propose a new method where filters could be derived from accurately
+computed numerical transform pairs, such as those computed using quadrature
+approaches. For example, the EM modeler \texttt{empymod} could be used to
+compute transform pairs for a particular class of model and data parameters
+(for example, a specific EM method and model of interest) using a slow but
+highly accurate quadrature method. A good filter derived from the resulting
+transform pairs could then be used for faster modeling of similar model and
+data parameters.
+
+The main features of \texttt{fdesign} and some differences between it and the
+methods it is based upon are:
 \begin{itemize}
@@
     the best outcome of the brute-force step.
-  \item \cite{GP.07.Kong, GEO.12.Key} optimize a filter for the $J_0$ filter,
-    and then use the obtained shift and spacing value to calculate the $J_1$
-    values. The presented algorithm can optimize either $J_0$, $J_1$, or both
-    in the optimization process. Five transform-pairs for each $J_0$ and $J_1$,
-    and three transform pairs for each sine and cosine of often used functions
-    are included in the routine, but any other transform pair can be provided
-    as input.
-  \item More complex or more specific models can be provided instead of
-    theoretical transform pairs by using the modeler \texttt{empymod}.
+  \item \cite{GP.07.Kong} and \cite{GEO.12.Key} optimize a filter for the $J_0$
+    filter, and then use the obtained shift and spacing value to calculate the
+    $J_1$ values. The presented algorithm can optimize either $J_0$, $J_1$, or
+    both in the optimization process. Five transform-pairs for each $J_0$ and
+    $J_1$, and three transform pairs for each sine and cosine of often used
+    functions are included in the routine, but any other transform pair can be
+    provided as input.
+  \item More complex or more specific numerical transforms  pairs  can be
+    provided instead of theoretical transform pairs by using the modeler
+    \texttt{empymod}.
   \item \cite{GP.07.Kong} defines a \emph{good} filter as one that recovers
-    \emph{weak} diffusive EM fields. In the presented algorithm you can define
-    a relative error level which defines up to what error an obtained result is
-    good or not. The presented algorithm can also minimize the amplitude, but
-    it has additionally another mode, where you maximise the abscissae $r$ of
-    the right-hand-side of the transform pair instead of minimising the
-    amplitude. The two modes yield the same result in the general, simple case
-    of fast decaying transform pairs. However, if you use complex models for
-    the design, and specifically high frequencies, then maximising $r$ will
-    yield much better and more consistent results.
-  \item The algorithm allows to run under-, equal-, and over-determined
+    \emph{weak} diffusive EM fields. In  \texttt{fdesign}, you can define a
+    relative error level that defines up to what error an obtained result is
+    good or not. The optimization seeks to find smallest amplitude field that
+    is accurately modeled (i.e., minimizing the amplitude), but it has
+    additionally another mode, where it will maximize the abscissae $r$ of the
+    right-hand-side of the transform pair instead of minimizing the amplitude.
+    The two modes yield the same result in the general, simple case of fast
+    decaying transform pairs. However, for some complex models and specifically
+    for very high frequencies, maximizing $r$ will yield much better and more
+    consistent results.
+  \item The algorithm can solve under-, equal-, and over-determined linear
     systems.
@@
 The quality of a filter depends heavily on the model chosen for comparison. An
-obtained filter might be very good for one model, but not that good for another
-one, which is no different for this designing tool. There is no way to estimate
-the error of a result obtained with a certain filter if you apply it to any
-other model, unless you calculate this other model with another method for
-comparison. All results presented here should therefore be taken with a certain
-care.
+obtained filter might be very good for one model or data acquisition setup, but
+not that good for another one, which is no different for this designing tool.
+There is no way to estimate the error of a result obtained with a certain
+filter if you apply it to any other model, unless you calculate this other
+model with another method for comparison. All results presented here should
+therefore be taken with a certain care. Further, when applying a new filter to
+a significantly different model and/or data space than what it has previously
+been tested on, we recommend first verifying its accuracy in that space by a
+comparison with results computed with a quadrature method such as QWE, which is
+also available in \texttt{empymod}.
 
@@
 %
-where $l$ and $r$ denote left-hand-side and right-hand-side evaluation values,
-and $K$ is the kernel function. In the specific case of the Hankel transform
-$l$ corresponds to wavenumber, $r$ to offset, and $K$ to Bessel functions; in
-the case of the Fourier transform $l$ corresponds to frequency, $r$ to time,
-and $K$ to sine or cosine functions. In both cases it is an infinite integral
-which numerical integration is very time-consuming because of the slow decay
-of the kernel function and its oscillatory behaviour.
+where $l$ and $r$ denote evaluation values, and $K$ is the kernel function. In
+the specific case of the Hankel transform $l$ corresponds to wavenumber, $r$ to
+offset, and $K$ to Bessel functions; in the case of the Fourier transform $l$
+corresponds to frequency, $r$ to time, and $K$ to sine or cosine functions. In
+both cases it is an infinite integral which numerical integration is very
+time-consuming because of the slow decay of the kernel function and its
+oscillatory behaviour.
 
 By substituting $r = e^x$ and $l = e^{-y}$ we get
+%
 \begin{equation}
@@
 \end{equation}
+%
 This can be re-written as a convolution integral and be approximated for an
 $N$-point filter by
+%
 \begin{equation}
@@
 \end{equation}
+%
 where $h$ is the digital linear filter, and the logarithmically spaced filter
 abscissae is a function of the spacing $\Delta$ and the shift $\delta$,
+%
 \begin{equation}
@@
 \end{equation}
+%
 From equation~\ref{eq:filtapprox} it can be seen that the filter method
-requires $N$ evaluations at each $r$. To calculate for instance the frequency
-domain result for 100 offsets with a 201\,pt filter requires 20'100 evaluations
+requires $N$ evaluations at each $r$. For example, to calculate the frequency
+domain result for 100 offsets with a 201\,pt filter requires 20,100 evaluations
 in the wavenumber domain. This is why the DLF often uses interpolation to
-minimise the required evaluations, either in the right-hand side (often called
-\emph{lagged convolution DLF}) or in the left-hand side (\emph{splined DLF}).
-
-\subsection{Pseudo-code}
+minimize the required evaluations, either for $F(r)$ in what is referred to as
+\emph{lagged convolution DLF}, or for $f(l)$, which we here call \emph{splined
+DLF}.
+
+In the direct matrix inversion method for solving for the digital filter
+coefficients, equation \ref{eq:filtapprox} is cast as a linear system as
+follows. $\Delta$ and $\delta$ are preassigned scalar values and $r$ is a range
+of $M$ preassigned values $r_m$. The filter base coefficients $b_n$ are
+computed using equation \ref{eq:base} and an array of values for $l$ are
+computed using $l_{mn} = b_n/r_m$. The linear system's left hand side (LHS)
+matrix $\bf{A}$ has dimensions $M \times N$ with coefficients $A_{mn} =
+f(l_{mn})/r_n$ and the right hand side (RHS) vector $\bf{v}$ has elements $v_n
+= F(r_n)$. The resulting linear system
+%
+\begin{equation}
+  \bf{A}\bf{h} = \bf{v}
+  \label{eq:,linearSystem}
+\end{equation}
+%
+can be solved by direct matrix inversion to obtain the vector of filter
+coefficients $\bf{h}$. For a given filter length $N$, there are many subjective
+choices that go into the practical implementation of this method, including the
+choice of the transform pairs $F(r)$ and $f(l)$, as well as the value for $M$
+and the particular range and spacing of the values $r_m$. Once values are
+chosen for these variables, an optimal filter can be found by a grid search
+over $\Delta$ and $\delta$ for values that produce a high quality filter. The
+choice of metric for what constitutes a high quality filter is also subjective
+and is further discussed below.
+
+\subsection{Pseudo-code for \texttt{fdesign} }
+
 The main input variables are the filter length ($N$), the spacing ($\Delta$)
-and shift ($\delta$) ranges over which to loop, and the transform pairs for
-the inversion ($f_\mr{I}$) and the check of quality ($f_\mr{C}$). If
-$f_\mr{C}$ is not provided, then $f_\mr{I}$ is used for both. There are
-additional, optional input parameters, for instance to adjust how the rhs
-abscissae $r = f(b, N)$ are calculated, where $b$ is the filter base.
+and shift ($\delta$) ranges over which to loop, and the transform pairs for the
+inversion ($f_\mr{I}$) and the check of quality ($f_\mr{C}$). If $f_\mr{C}$ is
+not provided, then $f_\mr{I}$ is used for both. There are additional, optional
+input parameters, for instance to adjust how the RHS abscissae for the
+inversion $r_\mr{I} = p(b, N)$ are calculated, where $b$ is the filter base.
+Additionally, ranges $r$ for evaluating the check function need to be
+specified.
 
@@
 \begin{enumerate}
-  \item Evaluate rhs of check-function $f_\mr{C}$:\newline
+  \item Evaluate RHS of check-function $f_\mr{C}$:\newline
         $d_\mr{R} = f_\mr{C}.\mr{rhs}(r)$
@@
         $b_n = \exp\left\{\Delta_i(-N/2+n) + \delta_j\right\}$
-      \item Get required rhs ($r$) and lhs ($l$) evaluation points:\newline
-        $r = f(b, N)$\newline
-        $l = b/r$
+      \item Get required RHS and LHS evaluation points:\newline
+        $r_\mr{I} = p(b, N)$\newline
+        $l_\mr{I} = b/r_\mr{I}$
       \item Invert for filter coefficients:\newline
-        $J_{ij} = \mr{solve}(f_\mr{I}.\mr{lhs}(l), f_\mr{I}.\mr{rhs}(r))$
-    \item Calculate numerically rhs of check-function $f_\mr{C}$ with current
+        $J_{ij} = \mr{solve}(f_\mr{I}.\mr{lhs}(l_\mr{I}),
+        f_\mr{I}.\mr{rhs}(r_\mr{I}))$
+    \item Calculate numerically RHS of check-function $f_\mr{C}$ with current
       filter $J_{ij}$:\newline
-      $d_\mr{F} = f_\mr{C}.\mr{lhs}(l)\cdot J_{ij}\ /\ r$
+      $d_\mr{F} = f_\mr{C}.\mr{lhs}(b/r)\cdot J_{ij}\ /\ r$
       \item Store minimum recovered amplitude or maximum $r$ where relative
@@
     (a local minimization can be run to polish the brute-force result):\newline
-    return $J\left[\mr{argmin}(\chi_{ij})\right] $\newline
+    return $J\left[\mr{argmin}(\chi)\right] $\newline
 \end{enumerate}
@@
 %
-\plot*{InvCrit}{width=.9\textwidth}{(a) A regular rhs curve of a transform pair
+\plot*{InvCrit}{width=.9\textwidth}{(a) A regular RHS curve of a transform pair
   with a purely decaying function. Minimum amplitude or maximum $r$ yield the
   same result in this scenario. Using a relative error criteria is more stable
-  than just the absolute minimum amplitude. (b) A 1D model for $f=80\,$MHz.
+  than just the absolute minimum amplitude. (b) A 1D model for $f=200\,$MHz.
   Here, the maximum $r$ provides a better criteria for the inversion.}
@@
 
-It is important to note that independent of the inversion criteria the obtained
-absolute value of it is only good to compare inversion results for the same
-model. If you calculate filters with a different transform pair you might get
-very different values, but this difference says nothing about the quality of
-it. For instance, a certain transform pair might yield a lower minimum
-amplitude than the other transform pair, but the resulting filter is worse if
-both filters are compared to the same transform pair. This is due to the
-characteristic of each transform pair.
+While the $\chi$ array is used to identify the optimal values of $\Delta$ and
+$\delta$, the value of $\mr{argmin}(\chi)$ will depend on the choice of the
+check function $f_C$ and can vary from one transform pair to another; thus it
+should not be used to compare the quality of filters created with different
+transform pairs.
 
@@
 The filter length in this case is $N=201$, and the theoretical transform pairs
-are (as given, e.g., in \cite{USGS.75.Anderson})
+are as given, e.g., in \cite{USGS.75.Anderson}:
 %
@@
 %
-where $a$ was set to 5. In the algorithm you can provide one pair for the inversion, and a different pair to get the minimum amplitude or the maximum $r$. In
-this case the same transform pair was used for both the inversion and the check
-of quality. The acceptable relative error was set to 1\,\%.
-The rhs evaluation parameter for the inversion was set to $r_\mr{def} = (1, 1,
-2)$, which means that an over-determined system was evaluated with $2N$
-equations, where $r$ was logarithmically spaced from $\log_{10}(1/\max(b)) - 1$
-to $\log_{10}(1/\min(b)) + 1$, where $b$ are the filter abscissae as given in
-equation~\ref{eq:base}.
+where $a$ was set to 5. In the algorithm you can provide one pair for the
+inversion, and a different pair to get the minimum amplitude or the maximum
+$r$. In this case the same transform pair was used for both the inversion and
+the check of quality. The acceptable relative error was set to 1\,\%. The RHS
+evaluation parameter for the inversion was set so  that an over-determined
+system was evaluated with $M = 2N$ equations, where $r_\mr{I}$ was
+logarithmically spaced from $\log_{10}(1/\max(b)) - 1$ to $\log_{10}(1/\min(b))
++ 1$, where $b$ are the filter abscissae as given in equation~\ref{eq:base}.
+This is accomplished in \texttt{fdesign} by setting parameter $r_\mr{def} = (1,
+1, 2)$.
 
@@
 Designing filters is to a large extent trial and error. All input variables
-influence the outcome, and often you will come across a good filter by shear
-luck of choosing the right starting parameters. Each of the input parameters
-has an effect to the outcome. It depends a lot on the transform pairs
-$f_\mr{I;C}$, and functions that decay rapidly are generally better, as noted
-by earlier authors \citep[e.g.][]{USGS.75.Anderson}. It also depends on the
-filter length, and obviously heavily on the spacing and shift values, as this
-is what we invert for. Another important point is how you define the right-hand
-evaluation points of the inversion ($r_\mr{def}$). Evaluating corresponding
-transform pairs separately or jointly also leads to different filter
-coefficients ($J_0$, $J_1$, or $J_0$ \& $J_1$, or equally sine, cosine, or sine
-\& cosine); also if the real or the imaginary part is used when complex
-transform pairs or \texttt{empymod} is used; and if it is inverted for the
-minimum amplitude or for the maximum $r$.
+influence the outcome, and often  a good filter is found by shear luck of
+choosing the right starting parameters. Each of the input parameters has an
+effect to the outcome. It also depends heavily on the transform pairs
+$f_\mr{I;C}$; functions that decay rapidly are generally better, as noted by
+earlier authors \citep[e.g.][]{USGS.75.Anderson}. It also depends on the filter
+length, and obviously heavily on the spacing and shift values. Another
+important point is how you define the right-hand evaluation points of the
+inversion ($r_\mr{def}$). Evaluating corresponding transform pairs separately
+or jointly also leads to different filter coefficients ($J_0$, $J_1$, or $J_0$
+\& $J_1$, or equally sine, cosine, or sine \& cosine); also if the real or the
+imaginary part is used when complex transform pairs or \texttt{empymod} is
+used; and if it is inverted for the minimum amplitude or for the maximum $r$.
 
@@
 In this section we compare the 201\,pt filter derived in the previous section
-to CSEM models used in \cite{GP.07.Kong}, \cite{GEO.12.Key}, and a land case.
+to marine CSEM models used in \cite{GP.07.Kong}, \cite{GEO.12.Key}, as well as
+for a land EM case.
 
@@
 half-spaces case used by \cite{GP.07.Kong} in his figure 5. The model consists
-of a water layer with $\rho_\mr{w} = 0.3125\,\Omega\,$m of infinite thickness
-above a subsurface half-space with $\rho = 1\,\Omega\,$m. The signal of an
-$x$-directed electric source 50\,m above the interface is measured at an
-$x$-directed electric receiver at the interface, frequency is $f=1\,$Hz.
+of a $0.3125\,\Omega\,$m sea layer of infinite thickness above a subsurface
+$1\,\Omega\,$m half-space. The signal of a $1\,$Hz frequency $x$-directed
+electric source 50\,m above the interface is measured at an $x$-directed
+electric receiver at the interface.
 %
@@
 
-Figure \ref{fig:keyf5} (a) is the canonical CSEM model as used in
-\cite{GEO.12.Key} in his figure 5: Water depth of 2\,km with
-$\rho_\mr{w}=0.303\,\Omega\,$m, a background resistivity of
-$\rho_\mr{b}=1\,\Omega\,$m, in which a target is embedded of
-$\rho_\mr{t}=100\,\Omega\,$m at 1000\,m below the seafloor, 100\,m thick.
-Source depth is 1990\,m, receivers are on the seafloor.
+Figure \ref{fig:keyf5}(a) considers the canonical CSEM model as used in
+\cite{GEO.12.Key} in his figure 5: a $0.303\,\Omega\,$m sea with depth of
+2\,km, a background resistivity of $1\,\Omega\,$m, in which a $100\,\Omega\,$m
+target layer of 100\,m thickness is embedded at 1000\,m below the seafloor.
+The source is located at 1990\,m and receivers are positioned on the seafloor.
 %
@@
 %
-Figure \ref{fig:keyf5} (b) is a land case, with a background resistivity
-of $\rho_\mr{b}=10\,\Omega\,$m, in which a target is embedded of
-$\rho_\mr{t}=500\,\Omega\,$m at 1000\,m below the surface, 100\,m thick. Source
-depth is 0.5\,m, receiver depth is 0.8\,m. In both cases the new filter
-\emph{Wer201} has generally a relative error which is orders of magnitude
-lower than the other filters. The relative error of the real part is given
-in black, whereas the relative error of the imaginary part is given in grey.
-It is interesting to note that the real and imaginary parts have very similar
-errors in \emph{Wer201}, \emph{Key201}, and \emph{And801}, but that the
-imaginary part of \emph{Kong241} seems to behave considerably better for most
-part than its real counterpart. To compare the real and the imaginary parts is
-insofar interesting as the digital filters are purely real valued.
+
+Figure \ref{fig:keyf5}(b) is a land case, with a background resistivity of
+$10\,\Omega\,$m, in which a $500\,\Omega\,$m target of 100\,m thickness is
+embedded at 1000\,m below the surface. The source depth is 0.5\,m and the
+receiver depth is 0.8\,m. In both cases the new filter \emph{Wer201} has
+generally a relative error which is orders of magnitude lower than the other
+filters. The relative error of the real part is given in black, whereas the
+relative error of the imaginary part is given in grey. It is interesting to
+note that the real and imaginary parts have very similar errors in
+\emph{Wer201}, \emph{Key201}, and \emph{And801}, but that the imaginary part of
+\emph{Kong241} seems to behave considerably better for most part than its real
+counterpart. To compare the real and the imaginary parts is insofar interesting
+as the digital filters are purely real valued.
 
@@
 accurate for these three models, it might not be the best filter in other
-cases.
+cases.  We also comment that while it is tempting to design very short length
+filters (much less than about $N=100$) since shorter filters can greatly reduce
+EM simulation times, in our experience, shorter filters tend to be less general
+purpose than longer filters and we highly recommend thoroughly testing their
+accuracy for the desired range of data and model parameters.
 
@@
 Figure~\ref{fig:GPR} shows the GPR example as calculated in
-\cite{GEO.15.Hunziker, GEO.17.Werthmuller}; the model parameters are given in
-subplot (a). The filter used for this example for the Hankel transform is a
-2001\,pt filter, derived with the fullspace solution with $f=500\,$MHz for the
-inversion and the check of quality. Vertical source and receiver separation is
-1\,m, resistivity $\rho=200\,\Omega\,$m, $\varepsilon_\mr{r}$=10,
-$\mu_\mr{r}=1$. For the Fourier transform a 4096\,pt FFT was used with
-regularly spaced frequencies from 0.5\,MHz to 850\,MHz and then zero-padded up
-to 2048\,MHz for both the calculation with the adaptive quadrature (QUA) and
-with the DLF. The frequency result is multiplied with a Ricker wavelet with a
-center frequency of 250\,MHz, and a gain function ($1 + |t^3|, t$ in ns) is
-applied. The calculation with the digital filter took under 9 minutes and is
-therefore roughly 80 times faster than the adaptive quadrature calculation
-which took roughly 11 hours and 27 minutes. Calculating the same model with QWE
-took 7 hours and 20 minutes. However, QWE uses in roughly 1/3 of the
-calculation internally the adaptive quadrature in this example, see
-\cite{GEO.17.Werthmuller}. (Note that DLF was run in parallel using 4 threads
-at once, taking effectively only 2 minutes and 10 seconds to calculate.
-%EMmod \citep{GEO.15.Hunziker} took for the same model about 18 hours 55
-%minutes, using a 61\,pt Gauss-Kronrod integration routine.
+\cite{GEO.15.Hunziker} and \cite{GEO.17.Werthmuller}; the model parameters are
+given in subplot (a). The filter used for this example for the Hankel transform
+is a 2001\,pt filter, derived with the fullspace solution with $f=500\,$MHz for
+the inversion and the check of quality. The whole space solution used
+resistivity $200\,\Omega\,$m, relative permittivity $\varepsilon_\mr{r}=10$,
+and permeability $\mu_\mr{r}=1$. The source and receiver vertical separation is
+1\,m. For the Fourier transform a 4096\,pt FFT was used with regularly spaced
+frequencies from 0.5\,MHz to 850\,MHz and then zero-padded up to 2048\,MHz. The
+frequency result is multiplied with a Ricker wavelet with a center frequency of
+250\,MHz, and a gain function ($1 + |t^3|, t$ in ns) is applied. Subplot (b)
+shows the result when using the adaptive quadrature (QUAD) option of
+\texttt{empymod}, which uses the QAGSE routine from the Fortran QUADPACK
+library. Subplot (c) shows the result using DLF with the new 2001\,pt digital
+filter. The calculation with DLF took under 9 minutes and is therefore roughly
+80 times faster than the calculation with QUAD, which took roughly 11 hours and
+27 minutes. Calculating the same model with QWE took 7 hours and 20 minutes.
+However, QWE uses in roughly 1/3 of the calculation internally QUAD in this
+example, see \cite{GEO.17.Werthmuller}. (Note that DLF was run in parallel
+using 4 threads at once, taking effectively only 2 minutes and 10 seconds to
+calculate.
+% EMmod \citep{GEO.15.Hunziker} took for the same model about 18 hours 55
+% minutes, using a 61\,pt Gauss-Kronrod integration routine.
 The lagged convolution version of DLF and the splined version of QWE were used
@@
 \plot*{GPR}{width=\textwidth}{GPR example for the model given in (a) using
-  \texttt{empymod} with (b) the adaptive quadrature (QAGSE from the Fortran
-  QUADPACK library) or with (c) a 2001\,pt digital linear filter (b). The
+  \texttt{empymod} with (b) the adaptive quadrature (QUAD) or with (c) a 2001\,pt digital linear filter (b). The
   quadrature took about 11\,h 27\,min to calculate, whereas the filter took
@@
 example. To show that this filter can also be applied to different layer
-parameters we run a test where we swapped layers one and two of the model
-in Figure \ref{fig:GPR} (a). The resulting real and imaginary parts for the
+parameters we run a test where we swapped layers one and two of the model in
+Figure \ref{fig:GPR} (a). The resulting real and imaginary parts for the
 offsets 0.2\,m, 2.0\,m, and 3.0\,m are given in Figure~\ref{fig:ftGPR2nd}.
-
 %
@@
 $J_0$ and $J_1$, with different spacing and shift values. The first idea would
-triple the calculation cost, the second idea double them; however, they both
-would still be very fast compared to standard quadrature methods. One could
-also apply the digital filter method to acoustic and elastic wavefields as
-well.
+roughly triple the calculation cost, the second idea double them; however, they
+both would still be very fast compared to standard quadrature methods. One
+could also apply the digital filter method to acoustic and elastic wavefields
+as well.
 
@@
 available from GitHub in the add-on repository of \texttt{empymod}, called
-\texttt{empyscripts}, from version 0.1.2 onwards.
+\texttt{empyscripts}, from version 0.2.0 onwards.
 
@@
 depends heavily on the model, and this new filter might or might not behave
-that well for other models.
-
-The shown GPR result shows that the digital linear filter method can also be
-used for wave phenomena, not only for the diffusive approximation limit of low
+that well for other models or data parameters.
+
+The GPR example shows that the digital linear filter method can also be used
+for wave phenomena, not only for the diffusive approximation limit of low
 frequency EM modeling.
 
-We except the presented algorithm is useful for at least 3 scenarios:
+We see \texttt{fdesign} as being useful for at least 3 scenarios:
 \begin{enumerate}
-  \item Provide a fast method to design problem-specific filters. For bigger
+  \item Providing a fast method to design problem-specific filters. For bigger
     inversion projects (such as, for instance, massive stochastic CSEM
@@
     as an optional pre-inversion step.
-  \item Extend the filter method to new areas, namely higher frequencies and
+  \item Extending the filter method to new areas, namely higher frequencies and
     acoustic and elastic wavefields.
-  \item Raise a new interest for digital linear filters in geophysics, by
-    making it very easy for anyone to play around, create their own filters,
-    and get a better understanding of it. We are sure there are many great
-    filters to be discovered.
+  \item Raising  interest for digital linear filters in geophysics by making it
+    very easy for anyone to play around, create their own filters, and get a
+    better understanding of the relative simplicity of the method. We are sure
+    there are many great filters that are yet to be discovered.
 \end{enumerate}
-~
+
+\section{Acknowledgment}
+
+D.W. would like to thank the electromagnetic research group of Aleksandr
+Mousatov at the IMP for fruitful discussions, and Jürg Hunziker for valuable
+feedback which improved this manuscript considerably.
 

### 2018-01-17 ManuscriptGeophysics2018.tex
Typos

@@
   filters shows that it performs better for some standard CSEM cases. The
-  derivation of a longer 2001\,pt filter for a GPR example with 250\,MHz centre
+  derivation of a longer 2001\,pt filter for a GPR example with 250\,MHz center
   frequency proves that the filter method works also for wave phenomena, not
@@
 power to calculate the filter coefficients. He further states that credit goes
-to his Ph.D. supervisor Otto \cite{BK.68.Koefoed, GP.70.Koefoed}, who retook
-the task of direct interpretational methods with the introduction of raised
-kernel functions. DLF is, as such, an improvement of that approach, providing a
-faster and simpler method.
+to \cite{BK.68.Koefoed, GP.70.Koefoed}, who retook the task of direct
+interpretational methods with the introduction of raised kernel functions. DLF
+is, as such, an improvement of that approach, providing a faster and simpler
+method.
 
@@
 used to measure the accuracy of the DLF. \cite{GEO.12.Key} presented codes for
-comparing the DLF approach with a speed optimized quadrature method call
+comparing the DLF approach with a speed optimized quadrature method called
 quadrature-with-extrapolation (QWE); despite its optimizations, QWE is still
@@
 Theoretically, it can be used to derive linear filters for any linear
-transform, as long as it is supplied with the input and outputs of a transform
-pair. Previous works have relied on using theoretical transforms pairs commonly
+transform, as long as it is supplied with the inputs and outputs of a transform
+pair. Previous works have relied on using theoretical transform pairs commonly
 found in tables of integral transforms. Here we follow a similar approach, but
@@
     provided as input.
-  \item More complex or more specific numerical transforms  pairs  can be
+  \item More complex or more specific numerical transform  pairs  can be
     provided instead of theoretical transform pairs by using the modeler
@@
     relative error level that defines up to what error an obtained result is
-    good or not. The optimization seeks to find smallest amplitude field that
-    is accurately modeled (i.e., minimizing the amplitude), but it has
-    additionally another mode, where it will maximize the abscissae $r$ of the
-    right-hand-side of the transform pair instead of minimizing the amplitude.
-    The two modes yield the same result in the general, simple case of fast
-    decaying transform pairs. However, for some complex models and specifically
-    for very high frequencies, maximizing $r$ will yield much better and more
-    consistent results.
+    good or not. The optimization seeks to find the smallest amplitude field
+    that is accurately modeled (i.e., minimizing the amplitude), but it has
+    additionally another mode, where it will maximize the output abscissae $r$
+    of the transform pair instead of minimizing the amplitude. The two modes
+    yield the same result in the general, simple case of fast decaying
+    transform pairs. However, for some complex models and specifically for very
+    high frequencies, maximizing $r$ will yield much better and more consistent
+    results.
   \item The algorithm can solve under-, equal-, and over-determined linear
@@
 %
-where $l$ and $r$ denote evaluation values, and $K$ is the kernel function. In
-the specific case of the Hankel transform $l$ corresponds to wavenumber, $r$ to
-offset, and $K$ to Bessel functions; in the case of the Fourier transform $l$
-corresponds to frequency, $r$ to time, and $K$ to sine or cosine functions. In
-both cases it is an infinite integral which numerical integration is very
-time-consuming because of the slow decay of the kernel function and its
-oscillatory behaviour.
+where $l$ and $r$ denote input and output evaluation values, respectively, and
+$K$ is the kernel function. In the specific case of the Hankel transform $l$
+corresponds to wavenumber, $r$ to offset, and $K$ to Bessel functions; in the
+case of the Fourier transform $l$ corresponds to frequency, $r$ to time, and
+$K$ to sine or cosine functions. In both cases it is an infinite integral which
+numerical integration is very time-consuming because of the slow decay of the
+kernel function and its oscillatory behaviour.
 
@@
 \plot*{GPR}{width=\textwidth}{GPR example for the model given in (a) using
-  \texttt{empymod} with (b) the adaptive quadrature (QUAD) or with (c) a 2001\,pt digital linear filter (b). The
-  quadrature took about 11\,h 27\,min to calculate, whereas the filter took
-  2\,min 10\,s in parallel on 4 nodes, hence under 9\,min in total.}
+  \texttt{empymod} with (b) the adaptive quadrature (QUAD) or with (c) a
+  2001\,pt digital linear filter (b). QUAD took about 11\,h 27\,min to
+  calculate, whereas DLF took 2\,min 10\,s in parallel on 4 nodes, hence under
+  9\,min in total.}
 %

### 2018-01-17 notebooks/printinfo.py
nb with empymod 1.5.0 and empyscripts 0.2.0

@@ -1,7 +1,8 @@
 """
 Utility to print date, time, and version information.
 
-This script was heavily inspired by:
+Adjusted from https://github.com/empymod/example-notebooks, which in turn was
+heavily inspired by:
 
     - ipynbtools.py from qutip https://github.com/qutip
     - watermark.py from https://github.com/rasbt/watermark
@@ -11,10 +12,11 @@ This script was heavily inspired by:
 import sys
 import time
 import platform
+import multiprocessing
 import numpy
 import scipy
-import numexpr
 import IPython
+import empyscripts
 import empymod
 import matplotlib
 from IPython.display import HTML
@@ -38,22 +40,22 @@ def info():
     html += "<tr" + style1 + ">"
     html += "<td" + style2 + ">%s</td>" % empymod.__version__
     html += "<td" + style1 + ">empymod</td>"
+    html += "<td" + style2 + ">%s</td>" % numpy.__version__
+    html += "<td" + style1 + ">numpy</td>"
     html += "<td" + style2 + ">%s</td>" % platform.system()
     html += "<td" + style1 + ">OS</td>"
-    html += "<td" + style2 + ">%s</td>" % numexpr.detect_number_of_threads()
-    html += "<td" + style1 + ">CPU(s)</td>"
     html += "<td" + style2 + ">%s</td>" % IPython.__version__
     html += "<td" + style1 + ">IPython</td>"
     html += "</tr>"
 
-    # numpy, scipy, numexpr, matplotlib
+    # numpy, scipy, matplotlib
     html += "<tr" + style1 + ">"
-    html += "<td" + style2 + ">%s</td>" % numpy.__version__
-    html += "<td" + style1 + ">numpy</td>"
+    html += "<td" + style2 + ">%s</td>" % empyscripts.__version__
+    html += "<td" + style1 + ">empyscripts</td>"
     html += "<td" + style2 + ">%s</td>" % scipy.__version__
     html += "<td" + style1 + ">scipy</td>"
-    html += "<td" + style2 + ">%s</td>" % numexpr.__version__
-    html += "<td" + style1 + ">numexpr</td>"
+    html += "<td" + style2 + ">%s</td>" % multiprocessing.cpu_count()
+    html += "<td" + style1 + ">CPU(s)</td>"
     html += "<td" + style2 + ">%s</td>" % matplotlib.__version__
     html += "<td" + style1 + ">matplotlib</td>"
     html += "</tr>"
[…]

### 2018-01-17 ManuscriptGeophysics2018.tex
nb with empymod 1.5.0 and empyscripts 0.2.0
Include error plots

@@
 In the direct matrix inversion method for solving for the digital filter
-coefficients, equation \ref{eq:filtapprox} is cast as a linear system as
+coefficients, equation~\ref{eq:filtapprox} is cast as a linear system as
 follows. $\Delta$ and $\delta$ are preassigned scalar values and $r$ is a range
 of $M$ preassigned values $r_m$. The filter base coefficients $b_n$ are
-computed using equation \ref{eq:base} and an array of values for $l$ are
+computed using equation~\ref{eq:base} and an array of values for $l$ are
 computed using $l_{mn} = b_n/r_m$. The linear system's left hand side (LHS)
@@
 \subsection{Design}
-Figure \ref{fig:201} shows the solution spaces of four consecutive inversion
+Figure~\ref{fig:201} shows the solution spaces of four consecutive inversion
 runs, where each run is a focus on a subsection of the previous run, indicated
@@
 
-Figure \ref{fig:201b} shows in (a) the filter values for $J_0$ and $J_1$ of the
+Figure~\ref{fig:201b} shows in (a) the filter values for $J_0$ and $J_1$ of the
 best 201 point filter obtained this way, and in (b) its right-hand-side
@@
 Figure~\ref{fig:kongf5} compares the derived 201\,pt filter with the two
-half-spaces case used by \cite{GP.07.Kong} in his figure 5. The model consists
+half-spaces case used by \cite{GP.07.Kong} in his figure~5. The model consists
 of a $0.3125\,\Omega\,$m sea layer of infinite thickness above a subsurface
@@
 \plot*{kongf5}{width=.9\textwidth}{(a) Results of different filters and QWE for
-  the model of figure 5 from \cite{GP.07.Kong} with the relative errors shown
+  the model of figure~5 from \cite{GP.07.Kong} with the relative errors shown
   in (b), using the QWE result. The relative error is meaningless from about
@@
 
-Figure \ref{fig:keyf5}(a) considers the canonical CSEM model as used in
-\cite{GEO.12.Key} in his figure 5: a $0.303\,\Omega\,$m sea with depth of
+Figure~\ref{fig:keyf5}(a) considers the canonical CSEM model as used in
+\cite{GEO.12.Key} in his figure~5: a $0.303\,\Omega\,$m sea with depth of
 2\,km, a background resistivity of $1\,\Omega\,$m, in which a $100\,\Omega\,$m
@@
 
-Figure \ref{fig:keyf5}(b) is a land case, with a background resistivity of
+Figure~\ref{fig:keyf5}(b) is a land case, with a background resistivity of
 $10\,\Omega\,$m, in which a $500\,\Omega\,$m target of 100\,m thickness is
@@
 accurate for these three models, it might not be the best filter in other
-cases.  We also comment that while it is tempting to design very short length
+cases. To verify the applicability of the filter we run two different tests
+over a wide range of scenarios. Figure~\ref{fig:error} shows the relative
+errors of DLF using the filter \emph{Wer201} compared to QWE for a deep water
+model, a shallow water model, and a land model. The model consists of an air
+halfspace, a water layer in the first two examples, and a subsurface halfspace.
+Source depth is 1990\,m, 10\,m, and 0.5\,m in the deep, shallow, and land case,
+respectively. Receiver depth is 2000\,m, 100\,m, and 0.8\,m. The subsurface
+halfspace resistivity varies from 1 to 10, 100, and 1000\,$\Omega\,$m from the
+left to the right column. The error plots cover offsets from 50\,m to 20\,km on
+the x-axis, and 0.01\,Hz to 10\,Hz on the y-axis. It can be seen in the result
+that the filter is sufficiently precise for all practical CSEM cases, hence has
+generally an error far below 1\,\% (1e-2) in all cases. The exception are large
+offsets with relatively high frequencies, e.g., 10\,km offset with 8\,Hz.
+%
+\plot*{error}{width=\textwidth}{Relative error plots comparing the DLF
+  (\emph{Wer201}) with QWE for the three cases deep water (top row), shallow
+  water (middle row), and land (bottom row). The subsurface halfspace
+  resistivity varies from 1 to 1000\,$\Omega\,$m from the left column to the
+  right column; offsets vary from 50\,m to 20\,km, and frequencies vary from
+  0.01\,Hz to 10\,Hz. The DLF provides sufficiently accurate results for all
+  practical CSEM applications, only breaking down for large offsets with
+  relatively high frequencies. However, the signal is usually below the noise
+  level anyway in this region.}
+%
+
+Another error test is shown in Figure~\ref{fig:skindepth1}. It shows in (a) the
+amplitude of the analytical fullspace solution for a wide range of
+resistivities and frequencies: $\rho=$ 1e-10$\,\Omega\,$m to 1e10\,$\Omega\,$m,
+and $f =$ 1e-10\,Hz to 1e14\,Hz. In this case, the horizontal offset is set to
+the skin depth, and the vertical source-receiver separation to a tenth of it,
+where the skin depth is given by
+%
+\begin{equation}
+  z_\mr{skin} = \left[\omega \sqrt{\frac{\mu\varepsilon}{2}}
+  \sqrt{\left\{1 + \left(\frac{1}{\omega\varepsilon\rho}\right\}^2 \right) -1}
+  \right]^{-1} \, . \\
+  \label{eq:skin}
+\end{equation}
+%
+The contour lines in the figure show the magnitude of the offset: the line with
+0 is where the offset is 1\,m, the line with 3 where the offset is 1\,km, etc.
+The contour lines of the offset have a knee, after which they become almost
+independent of frequency. This is where the quasi static approximation breaks
+down, and wave phenomena become dominant. Subplot (b) shows the relative error
+of the DLF using the filter \emph{Wer201} compared to the analytical solution.
+It shows nicely that the filter can be used for all cases that are in the quasi
+static region, but not outside.
+%
+\plot*{skindepth1}{width=.8\textwidth}{(a) Amplitude of the analytical
+  fullspace solution for a wide range of resistivities and frequencies. The
+  contour lines show the power of the offset used for the calculation, where
+  the offset is set to the skin depth. (b) Relative error of the DLF compared
+  with the analytical solution. The filter \emph{Wer201} works well in the
+  diffusive limit, but fails when wave phenomena become dominant.}
+%
+
+We also comment that while it is tempting to design very short length
 filters (much less than about $N=100$) since shorter filters can greatly reduce
@@
 parameters we run a test where we swapped layers one and two of the model in
-Figure \ref{fig:GPR} (a). The resulting real and imaginary parts for the
+Figure~\ref{fig:GPR} (a). The resulting real and imaginary parts for the
 offsets 0.2\,m, 2.0\,m, and 3.0\,m are given in Figure~\ref{fig:ftGPR2nd}.

### 2018-01-18 ManuscriptGeophysics2018.tex
Add links for remaining 2 references
Add Ward and Hohmann reference

@@
 the skin depth, and the vertical source-receiver separation to a tenth of it,
-where the skin depth is given by
+where the skin depth is given by \citep[e.g., ][ equation~1.49]{B.SEG.88.Ward}
 %
@@
   integration and digital linear filtering: USGS, {\bf PB221240}.
+\newblock
+  (\href{https://ntrl.ntis.gov/NTRL/dashboard/searchResults/titleDetail/PB221240.xhtml}{https://ntrl.ntis.gov\-/NTRL\-/dashboard\-/searchResults\-/titleDetail\-/PB221240.xhtml}).
 
@@
   0 and 1 by adaptive digital filtering: Geophysics, {\bf 44}, 1287--1305.
-\newblock (\href{http://doi.org/10.1190/1.1441007}{doi: 10.1190/1.1441007}).
+\newblock (\href{https://doi.org/10.1190/1.1441007}{doi: 10.1190/1.1441007}).
 
@@
   ACM Trans. Math. Softw., {\bf 8}, 344--368.
-\newblock (\href{http://doi.acm.org/10.1145/356012.356014}{doi:
+\newblock (\href{https://doi.org/10.1145/356012.356014}{doi:
   10.1145/356012.356014}).
@@
   Geophysics, {\bf 49}, 1811--1812.
-\newblock (\href{http://doi.org/10.1190/1.1441595}{doi: 10.1190/1.1441595}).
+\newblock (\href{https://doi.org/10.1190/1.1441595}{doi: 10.1190/1.1441595}).
 
@@
   modeling: Geophysics, {\bf 54}, 263--266.
-\newblock (\href{http://doi.org/10.1190/1.1442650}{doi: 10.1190/1.1442650}).
+\newblock (\href{https://doi.org/10.1190/1.1442650}{doi: 10.1190/1.1442650}).
 
@@
   {\bf 26}, 841--852.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1978.tb01636.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1978.tb01636.x}{doi:
   10.1111/j.1365-2478.1978.tb01636.x}).
@@
   soundings*: Geophysical Prospecting, {\bf 24}, 354--370.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1976.tb00932.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1976.tb00932.x}{doi:
   10.1111/j.1365-2478.1976.tb00932.x}).
@@
   1671--1686.
-\newblock (\href{http://doi.org/10.1190/1.1441448}{doi: 10.1190/1.1441448}).
+\newblock (\href{https://doi.org/10.1190/1.1441448}{doi: 10.1190/1.1441448}).
 
@@
   Geophysical Prospecting, {\bf 38}, 545--568.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1990.tb01861.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1990.tb01861.x}{doi:
   10.1111/j.1365-2478.1990.tb01861.x}).
@@
   765--780.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1974.tb00117.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1974.tb00117.x}{doi:
   10.1111/j.1365-2478.1974.tb00117.x}).
@@
   filtering: Geophysical Prospecting, {\bf 22}, 476--489.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1974.tb00100.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1974.tb00100.x}{doi:
   10.1111/j.1365-2478.1974.tb00100.x}).
@@
   Prospecting, {\bf 28}, 610--619.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1980.tb01246.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1980.tb01246.x}{doi:
   10.1111/j.1365-2478.1980.tb01246.x}).
@@
   International, {\bf 66}, 733--740.
-\newblock (\href{http://doi.org/10.1111/j.1365-246X.1981.tb04897.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-246X.1981.tb04897.x}{doi:
   10.1111/j.1365-246X.1981.tb04897.x}).
@@
   resistivity and em sounding curves: Geoexploration, {\bf 18}, 297--310.
-\newblock (\href{http://doi.org/10.1016/0016-7142(81)90059-4}{doi:
+\newblock (\href{https://doi.org/10.1016/0016-7142(81)90059-4}{doi:
   10.1016/0016-7142(81)90059-4}).
@@
   Geophysical Journal International, {\bf 69}, 55--66.
-\newblock (\href{http://doi.org/10.1111/j.1365-246X.1982.tb04935.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-246X.1982.tb04935.x}{doi:
   10.1111/j.1365-246X.1982.tb04935.x}).
@@
   Geophysics, {\bf 11}, 54--62.
-\newblock (\href{http://library.seg.org/doi/abs/10.1071/EG980054}{doi:
-  10.1071/EG980054}).
+\newblock (\href{https://doi.org/10.1071/EG980054}{doi: 10.1071/EG980054}).
 
@@
   Geophysical Prospecting, {\bf 19}, 192--217.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1971.tb00593.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1971.tb00593.x}{doi:
   10.1111/j.1365-2478.1971.tb00593.x}).
@@
   Prospecting, {\bf 19}, 769--775.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1971.tb00915.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1971.tb00915.x}{doi:
   10.1111/j.1365-2478.1971.tb00915.x}).
@@
   increased accuracy: Geophysical Prospecting, {\bf 30}, 501--514.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1982.tb01320.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1982.tb01320.x}{doi:
   10.1111/j.1365-2478.1982.tb01320.x}).
@@
   {J}0 and {J}1 transforms: Geophysical Prospecting, {\bf 45}, 745--762.
-\newblock (\href{http://doi.org/10.1046/j.1365-2478.1997.500292.x}{doi:
+\newblock (\href{https://doi.org/10.1046/j.1365-2478.1997.500292.x}{doi:
   10.1046/10.1046/j.1365-2478.1997.500292.x}).
@@
   problem: Geophysics, {\bf 80}, no. 1, F1--F18.
-\newblock (\href{http://doi.org/10.1190/geo2013-0411.1}{doi:
+\newblock (\href{https://doi.org/10.1190/geo2013-0411.1}{doi:
   10.1190/geo2013-0411.1}).
@@
   23}, 449--458.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1975.tb01541.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1975.tb01541.x}{doi:
   10.1111/j.1365-2478.1975.tb01541.x}).
@@
   Prospecting, {\bf 27}, 876--901.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1979.tb01005.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1979.tb01005.x}{doi:
   10.1111/j.1365-2478.1979.tb01005.x}).
@@
   layers: Geophysics, {\bf 74}, no. 2, F9--F20.
-\newblock (\href{http://doi.org/10.1190/1.3058434}{doi: 10.1190/1.3058434}).
+\newblock (\href{https://doi.org/10.1190/1.3058434}{doi: 10.1190/1.3058434}).
 
@@
   Geophysics, {\bf 77}, no. 3, F21--F30.
-\newblock (\href{http://doi.org/10.1190/GEO2011-0237.1}{doi:
+\newblock (\href{https://doi.org/10.1190/GEO2011-0237.1}{doi:
   10.1190/GEO2011-0237.1}).
@@
   geoelectrical resistivity measurements: Gebrüder Borntraeger.
-\newblock ({I}SBN: 9783443130022).
+\newblock (\href{https://isbnsearch.org/isbn/9783443130022}{{I}SBN:
+  9783443130022}).
 
@@
   {\bf 18}, 564--570.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1970.tb02129.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1970.tb02129.x}{doi:
   10.1111/j.1365-2478.1970.tb02129.x}).
@@
   sounding data: Geophysical Prospecting, {\bf 20}, 403--405.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1972.tb00643.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1972.tb00643.x}{doi:
   10.1111/j.1365-2478.1972.tb00643.x}).
@@
   resistivity sounding data: Geophysical Prospecting, {\bf 24}, 31--48.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1976.tb00383.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1976.tb00383.x}{doi:
   10.1111/j.1365-2478.1976.tb00383.x}).
@@
   {\bf 27}, 245--250.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1979.tb00968.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1979.tb00968.x}{doi:
   10.1111/j.1365-2478.1979.tb00968.x}).
@@
   406--420.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1972.tb00644.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1972.tb00644.x}{doi:
   10.1111/j.1365-2478.1972.tb00644.x}).
@@
   conductive medium: Geophysical Prospecting, {\bf 55}, 83--89.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.2006.00585.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.2006.00585.x}{doi:
   10.1111/j.1365-2478.2006.00585.x}).
@@
   {\bf 25}, 780--789.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1977.tb01204.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1977.tb01204.x}{doi:
   10.1111/j.1365-2478.1977.tb01204.x}).
@@
   filters: Geophysical Prospecting, {\bf 26}, 352--358.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1978.tb01598.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1978.tb01598.x}{doi:
   10.1111/j.1365-2478.1978.tb01598.x}).
@@
   Geophysical Prospecting, {\bf 42}, 131--139.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1994.tb00202.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1994.tb00202.x}{doi:
   10.1111/j.1365-2478.1994.tb00202.x}).
@@
   Geophysics, {\bf 51}, 1608--1627.
-\newblock (\href{http://doi.org/10.1190/1.1442212}{doi: 10.1190/1.1442212}).
+\newblock (\href{https://doi.org/10.1190/1.1442212}{doi: 10.1190/1.1442212}).
 
@@
   transform: Geophysical Prospecting, {\bf 34}, 897--903.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1986.tb00500.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1986.tb00500.x}{doi:
   10.1111/j.1365-2478.1986.tb00500.x}).
@@
   apparent resistivity computations: Exploration Geophysics, {\bf 6}, 104--109.
-\newblock (\href{http://library.seg.org/doi/abs/10.1071/EG975104}{doi:
-  10.1071/EG975104}).
+\newblock (\href{https://doi.org/10.1071/EG975104}{doi: 10.1071/EG975104}).
 
@@
   Prospecting, {\bf 32}, 105--123.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1984.tb00720.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1984.tb00720.x}{doi:
   10.1111/j.1365-2478.1984.tb00720.x}).
@@
   the {P223F} software suite: ASEG Technical Program Expanded Abstracts,  1--4.
-\newblock (\href{http://doi.org/10.1071/ASEG2007ab114}{doi:
+\newblock (\href{https://doi.org/10.1071/ASEG2007ab114}{doi:
   10.1071/ASEG2007ab114}).
@@
   method for horizontal structures: Physics, {\bf 4}, 307--322.
-\newblock (\href{http://doi.org/10.1063/1.1745198}{doi : 10.1063/1.1745198}).
+\newblock (\href{https://doi.org/10.1063/1.1745198}{doi : 10.1063/1.1745198}).
 
@@
   70--76.
-\newblock (\href{http://doi.org/10.1111/j.1365-2478.1973.tb00015.x}{doi:
+\newblock (\href{https://doi.org/10.1111/j.1365-2478.1973.tb00015.x}{doi:
   10.1111/j.1365-2478.1973.tb00015.x}).
+
+\bibitem[Ward and Hohmann, 1988]{B.SEG.88.Ward}
+Ward, S.~H., and G.~W. Hohmann,  1988, Electromagnetic theory for geophysical
+applications: SEG, Investigations in Geophysics, No.~3, 4, 130--131.
+\newblock  (\href{https://doi.org/10.1190/1.9781560802631}{doi:
+  10.1190/1.9781560802631}).
 
@@
   WB9--WB19.
-\newblock (\href{http://doi.org/10.1190/geo2016-0626.1}{doi:
+\newblock (\href{https://doi.org/10.1190/geo2016-0626.1}{doi:
   10.1190/geo2016-0626.1}).

### 2018-01-19 ManuscriptGeophysics2018.tex
Changes after call with Key

@@
 
-\title{A tool for designing digital filters for the Hankel and Fourier transforms in
-potential, diffusive, and wavefield modeling}
+\title{A tool for designing digital filters for the Hankel and Fourier
+transforms in potential, diffusive, and wavefield modeling}
 
@@
 errors of DLF using the filter \emph{Wer201} compared to QWE for a deep water
-model, a shallow water model, and a land model. The model consists of an air
-halfspace, a water layer in the first two examples, and a subsurface halfspace.
-Source depth is 1990\,m, 10\,m, and 0.5\,m in the deep, shallow, and land case,
-respectively. Receiver depth is 2000\,m, 100\,m, and 0.8\,m. The subsurface
-halfspace resistivity varies from 1 to 10, 100, and 1000\,$\Omega\,$m from the
-left to the right column. The error plots cover offsets from 50\,m to 20\,km on
-the x-axis, and 0.01\,Hz to 10\,Hz on the y-axis. It can be seen in the result
-that the filter is sufficiently precise for all practical CSEM cases, hence has
-generally an error far below 1\,\% (1e-2) in all cases. The exception are large
-offsets with relatively high frequencies, e.g., 10\,km offset with 8\,Hz.
+model, a shallow water model, and a land model. The models consist of air
+(upper halfspace), a water layer in the first two examples, and a subsurface
+halfspace. Source depth is 1990\,m, 10\,m, and 0.5\,m in the deep, shallow, and
+land case, respectively. Receiver depth is 2000\,m, 100\,m, and 0.8\,m. The
+subsurface halfspace resistivity varies from 1 to 10, 100, and
+1000\,$\Omega\,$m from the left to the right column. The error plots cover
+offsets from 50\,m to 20\,km on the x-axis, and 0.01\,Hz to 10\,Hz on the
+y-axis. It can be seen in the result that the filter is sufficiently precise
+for all practical CSEM cases, hence has generally an error far below 1\,\%
+(1e-2) in all cases. The exception are large offsets with relatively high
+frequencies, e.g., 10\,km offset with 8\,Hz. In the shallow water case this
+zone has a very high error for subsurface resistivities of 1\,$\Omega\,$m and
+10\,$\Omega\,$m (yellow zones). However, this is due to the fact that the
+amplitudes of the signal for these offsets and frequencies are in the order of
+1e-25~V/m.
 %
@@
   0.01\,Hz to 10\,Hz. The DLF provides sufficiently accurate results for all
-  practical CSEM applications, only breaking down for large offsets with
-  relatively high frequencies. However, the signal is usually below the noise
-  level anyway in this region.}
+  practical CSEM applications. The exception is in the shallow water case for
+  $\rho = 1\,\Omega\,$m and $10\,\Omega\,$m for large offsets and high
+  frequencies (yellow zones). However, the amplitudes in these regions are in
+  the order of 1e-25~V/m, and therefore far below the noise level.}
 %
@@
 However, QWE uses in roughly 1/3 of the calculation internally QUAD in this
-example, see \cite{GEO.17.Werthmuller}. (Note that DLF was run in parallel
-using 4 threads at once, taking effectively only 2 minutes and 10 seconds to
+example, see \cite{GEO.17.Werthmuller}. Note that DLF was run in parallel using
+4 threads at once, taking effectively only 2 minutes and 10 seconds to
 calculate.
@@
 The lagged convolution version of DLF and the splined version of QWE were used
-in this comparison.)
+in this comparison.
 %

### 2018-01-22 ManuscriptGeophysics2018.tex
Corrections and additions re z_skin

@@
 the quadrature with extrapolation (QWE), for which we used a 51\,pt quadrature
-with relative and absolute tolerance of 1e-12 and 1e-30, respectively. The
-relative error is shown in (b), where the QWE result was taken as \emph{truth}.
-For offsets greater than roughly 15.5\,km the relative error becomes
-meaningless, as the QWE fails itself; this part is greyed out in the figure.
+with relative and absolute tolerance of $10^{-12}$ and $10^{-30}$,
+respectively. The relative error is shown in (b), where the QWE result was
+taken as \emph{truth}. For offsets greater than roughly 15.5\,km the relative
+error becomes meaningless, as the QWE fails itself; this part is greyed out in
+the figure.
 
@@
 for all practical CSEM cases, hence has generally an error far below 1\,\%
-(1e-2) in all cases. The exception are large offsets with relatively high
+($10^{-2}$) in all cases. The exception are large offsets with relatively high
 frequencies, e.g., 10\,km offset with 8\,Hz. In the shallow water case this
@@
 amplitudes of the signal for these offsets and frequencies are in the order of
-1e-25~V/m.
+$10^{-25}\,$V/m.
 %
@@
   frequencies (yellow zones). However, the amplitudes in these regions are in
-  the order of 1e-25~V/m, and therefore far below the noise level.}
+  the order of $10^{-25}\,$V/m, and therefore far below the noise level.}
 %
@@
 amplitude of the analytical fullspace solution for a wide range of
-resistivities and frequencies: $\rho=$ 1e-10$\,\Omega\,$m to 1e10\,$\Omega\,$m,
-and $f =$ 1e-10\,Hz to 1e14\,Hz. In this case, the horizontal offset is set to
-the skin depth, and the vertical source-receiver separation to a tenth of it,
-where the skin depth is given by \citep[e.g., ][ equation~1.49]{B.SEG.88.Ward}
+resistivities and frequencies: $\rho=$ $10^{-10}\,\Omega\,$m to
+$10^{10}\,\Omega\,$m, and $f =10^{-10}$\,Hz to $10^{14}$\,Hz. In this case, the
+horizontal offset is set to the skin depth, and the vertical source-receiver
+separation to a tenth of it, where the skin depth is given by \citep[e.g., ][
+equation~1.49]{B.SEG.88.Ward}
 %
 \begin{equation}
-  z_\mr{skin} = \left[\omega \sqrt{\frac{\mu\varepsilon}{2}}
-  \sqrt{\left\{1 + \left(\frac{1}{\omega\varepsilon\rho}\right\}^2 \right) -1}
-  \right]^{-1} \, . \\
+  z_\mr{skin} = \left\{\frac{\omega^2 \varepsilon\mu}{2}
+    \left[ \sqrt{1 + \frac{1}{\left(\omega\varepsilon\rho\right)^2}} -1\right]
+  \right\}^{-1/2} \, . \\
   \label{eq:skin}
@@
 %
-The contour lines in the figure show the magnitude of the offset: the line with
-0 is where the offset is 1\,m, the line with 3 where the offset is 1\,km, etc.
-The contour lines of the offset have a knee, after which they become almost
-independent of frequency. This is where the quasi static approximation breaks
-down, and wave phenomena become dominant. Subplot (b) shows the relative error
-of the DLF using the filter \emph{Wer201} compared to the analytical solution.
-It shows nicely that the filter can be used for all cases that are in the quasi
-static region, but not outside.
+The grey contour lines in the figure show the magnitude of the offset: the line
+with 0 is where the offset is 1\,m, the line with 3 where the offset is 1\,km,
+etc. The contour lines of the offset have a knee, after which they become
+almost independent of frequency. This is where the quasi static approximation
+breaks down, and wave phenomena become dominant. The red line indicates where
+$\omega\varepsilon\rho = 1$. Subplot (b) shows the relative error of the DLF
+using the filter \emph{Wer201} compared to the analytical solution. It shows
+nicely that the filter can be used for all cases that are in the quasi static
+region, but not outside.
 %
@@
   with the analytical solution. The filter \emph{Wer201} works well in the
-  diffusive limit, but fails when wave phenomena become dominant.}
+  diffusive limit, but fails when wave phenomena become dominant. The red
+ line shows where $\omega\varepsilon\rho = 1$.}
 %

### 2018-01-22 ManuscriptGeophysics2018.tex
Version of initias submission to Geophysics

@@
   diffusive-, and wavefield modeling. Digital filters can be derived for any
-  electromagnetic method, such as methods in the diffusive limits (DC, CSEM) as
-  well as methods using higher frequency content (GPR, acoustic and elastic
-  wavefields).
+  electromagnetic method, such as methods in the diffusive limits (direct
+  current, controlled-source electromagnetics) as well as methods using higher
+  frequency content (ground-penetrating radar, acoustic and elastic wavefields).
 %
@@
   theoretical transform pairs are used for the inversion. Alternatively, one
-  can provide 1D subsurface models that will be computed with a precise
-  quadrature method using the EM modeler \texttt{empymod} to generate
-  numerical transform pairs.
+  can provide layered subsurface models that will be computed with a precise
+  quadrature method using the electromagnetic modeler \texttt{empymod} to
+  generate numerical transform pairs.
 %
@@
   The comparison of the presented 201\,pt filter with previously presented
-  filters shows that it performs better for some standard CSEM cases. The
-  derivation of a longer 2001\,pt filter for a GPR example with 250\,MHz center
-  frequency proves that the filter method works also for wave phenomena, not
-  only for diffusive EM fields.
+  filters shows that it performs better for some standard controlled-source
+  electromagnetic cases. The derivation of a longer 2001\,pt filter for a
+  ground-penetrating radar example with 250\,MHz center frequency proves that
+  the filter method works also for wave phenomena, not only for diffusive
+  electromagnetic fields.
 %
@@
 pair; and in (b) for a more complex, high frequency layered model. The circles
-show the minimum amplitude used in previous approaches. This criteria is not
-ideal, as it is subject to some random fluctuations and also depends on the
-choice of $r$. The squares show the minimum amplitude given a certain
-acceptable error, and the diamonds show the maximum $r$ given a certain
-acceptable error (the inversion is a minimization process, it therefore
-minimizes $1/r$, not $r$). In simple cases the minimum amplitude given an
-acceptable error and the maximum $r$ given an acceptable error will yield the
-same result. However, in complex cases the maximum $r$ is more consistent and
-therefore a much better criteria.
+show the minimum amplitude used in previous approaches. This criterion is not
+ideal, as it is subject to some random fluctuations and depends on the choice
+of $r$. The squares show the minimum amplitude given a certain acceptable
+error, and the diamonds show the maximum $r$ given a certain acceptable error
+(the inversion is a minimization process, it therefore minimizes $1/r$, not
+$r$). In simple cases the minimum amplitude given an acceptable error and the
+maximum $r$ given an acceptable error will yield the same result. However, in
+complex cases the maximum $r$ is more consistent and therefore a much better
+criterion.
 %
@@
 Designing filters is to a large extent trial and error. All input variables
-influence the outcome, and often  a good filter is found by shear luck of
+influence the outcome, and often  a good filter is found by sheer luck of
 choosing the right starting parameters. Each of the input parameters has an
-effect to the outcome. It also depends heavily on the transform pairs
+effect on the outcome. It also depends heavily on the transform pairs
 $f_\mr{I;C}$; functions that decay rapidly are generally better, as noted by
@@
 Figure~\ref{fig:kongf5} compares the derived 201\,pt filter with the two
-half-spaces case used by \cite{GP.07.Kong} in his figure~5. The model consists
+half-spaces case used by \cite{GP.07.Kong} in his Figure~5. The model consists
 of a $0.3125\,\Omega\,$m sea layer of infinite thickness above a subsurface
@@
 \plot*{kongf5}{width=.9\textwidth}{(a) Results of different filters and QWE for
-  the model of figure~5 from \cite{GP.07.Kong} with the relative errors shown
+  the model of Figure~5 from \cite{GP.07.Kong} with the relative errors shown
   in (b), using the QWE result. The relative error is meaningless from about
@@
 Figure~\ref{fig:keyf5}(a) considers the canonical CSEM model as used in
-\cite{GEO.12.Key} in his figure~5: a $0.303\,\Omega\,$m sea with depth of
+\cite{GEO.12.Key} in his Figure~5: a $0.303\,\Omega\,$m sea with depth of
 2\,km, a background resistivity of $1\,\Omega\,$m, in which a $100\,\Omega\,$m
@@
     \left[ \sqrt{1 + \frac{1}{\left(\omega\varepsilon\rho\right)^2}} -1\right]
-  \right\}^{-1/2} \, . \\
+  \right\}^{-1/2} \, , \\
   \label{eq:skin}
@@
 %
-The grey contour lines in the figure show the magnitude of the offset: the line
-with 0 is where the offset is 1\,m, the line with 3 where the offset is 1\,km,
-etc. The contour lines of the offset have a knee, after which they become
-almost independent of frequency. This is where the quasi static approximation
-breaks down, and wave phenomena become dominant. The red line indicates where
-$\omega\varepsilon\rho = 1$. Subplot (b) shows the relative error of the DLF
-using the filter \emph{Wer201} compared to the analytical solution. It shows
-nicely that the filter can be used for all cases that are in the quasi static
-region, but not outside.
+where we used in the example $10\,\varepsilon_0$ for the electric permittivity
+and $\mu_0$ for the magnetic permeability; $\omega = 2\pi f$. The grey contour
+lines in the figure show the magnitude of the offset: the line with 0 is where
+the offset is 1\,m, the line with 3 where the offset is 1\,km, etc. The contour
+lines of the offset have a knee, after which they become almost independent of
+frequency. This is where the quasi static approximation breaks down, and wave
+phenomena become dominant. The red line indicates where $\omega\varepsilon\rho
+= 1$. Subplot (b) shows the relative error of the DLF using the filter
+\emph{Wer201} compared to the analytical solution. It shows nicely that the
+filter can be used for all cases that are in the quasi static region, but not
+outside.
 %
@@
 resistivity $200\,\Omega\,$m, relative permittivity $\varepsilon_\mr{r}=10$,
-and permeability $\mu_\mr{r}=1$. The source and receiver vertical separation is
-1\,m. For the Fourier transform a 4096\,pt FFT was used with regularly spaced
-frequencies from 0.5\,MHz to 850\,MHz and then zero-padded up to 2048\,MHz. The
-frequency result is multiplied with a Ricker wavelet with a center frequency of
-250\,MHz, and a gain function ($1 + |t^3|, t$ in ns) is applied. Subplot (b)
-shows the result when using the adaptive quadrature (QUAD) option of
-\texttt{empymod}, which uses the QAGSE routine from the Fortran QUADPACK
-library. Subplot (c) shows the result using DLF with the new 2001\,pt digital
-filter. The calculation with DLF took under 9 minutes and is therefore roughly
-80 times faster than the calculation with QUAD, which took roughly 11 hours and
-27 minutes. Calculating the same model with QWE took 7 hours and 20 minutes.
-However, QWE uses in roughly 1/3 of the calculation internally QUAD in this
-example, see \cite{GEO.17.Werthmuller}. Note that DLF was run in parallel using
-4 threads at once, taking effectively only 2 minutes and 10 seconds to
+and permeability $\mu_\mr{r}=1$. Using Equation~\ref{eq:skin} yields a skin
+depth of roughly 3.6\,m for these parameters. The source and receiver vertical
+separation is 1\,m. For the Fourier transform a 4096\,pt FFT was used with
+regularly spaced frequencies from 0.5\,MHz to 850\,MHz and then zero-padded up
+to 2048\,MHz. The frequency result is multiplied with a Ricker wavelet with a
+center frequency of 250\,MHz, and a gain function ($1 + |t^3|, t$ in ns) is
+applied. Subplot (b) shows the result when using the adaptive quadrature (QUAD)
+option of \texttt{empymod}, which uses the QAGSE routine from the Fortran
+QUADPACK library. Subplot (c) shows the result using DLF with the new 2001\,pt
+digital filter. The calculation with DLF took under 9 minutes and is therefore
+roughly 80 times faster than the calculation with QUAD, which took roughly 11
+hours and 27 minutes. Calculating the same model with QWE took 7 hours and 20
+minutes. However, QWE uses in roughly 1/3 of the calculation internally QUAD in
+this example, see \cite{GEO.17.Werthmuller}. Note that DLF was run in parallel
+using 4 threads at once, taking effectively only 2 minutes and 10 seconds to
 calculate.

### 2018-06-08 ManuscriptGeophysics2018.tex
Add number (GEO-2018-0069)
In review; pseudo-code

@@
 % \documentclass[manuscript]{geophysics}
-\documentclass[paper,twocolumn,twoside]{geophysics}
+% \documentclass[paper,twocolumn,twoside]{geophysics}
 % \documentclass[paper]{geophysics}
 % \documentclass[paper, revised]{geophysics}
-% \documentclass[manuscript,revised]{geophysics}
+\documentclass[manuscript,revised]{geophysics}
 
@@
 
-\ms{GEO-2018-????}
+\ms{GEO-2018-0069.R2}
 
 \address{
-\footnotemark[1]Instituto Mexicano del Petróleo,
-Eje Central Lázaro Cárdenas Norte 152,
-Col. San Bartolo Atepehuacan C.P. 07730,
-Ciudad de México, México,
-E-mail: \href{mailto:dieter@werthmuller.org}{Dieter@Werthmuller.org};
-\footnotemark[2]Lamont-Doherty Earth Observatory, Columbia University,
-305C Oceanography,
-61 Route 9W, PO Box 1000,
-Palisades NY 10964-8000 US,
-E-mail: \href{mailto:KKey@ldeo.columbia.edu}{KKey@ldeo.columbia.edu};
-\footnotemark[3]TU Delft,
-Building 23,
-Stevinweg 1 / PO-box 5048,
-2628 CN Delft,
-E-mail: \href{mailto:E.C.Slob@tudelft.nl}{E.C.Slob@tudelft.nl}.}
+% \footnotemark[1]TU Delft,
+% Building 23,
+% Stevinweg 1 / PO-box 5048,
+% 2628 CN Delft,
+% E-mail: \href{mailto:E.C.Slob@tudelft.nl}{E.C.Slob@tudelft.nl};
+% \footnotemark[2]Instituto Mexicano del Petróleo,
+% Eje Central Lázaro Cárdenas Norte 152,
+% Col. San Bartolo Atepehuacan C.P. 07730,
+% Ciudad de México, México,
+% E-mail: \href{mailto:dieter@werthmuller.org}{Dieter@Werthmuller.org};
+% \footnotemark[3]Lamont-Doherty Earth Observatory, Columbia University,
+% 305C Oceanography,
+% 61 Route 9W, PO Box 1000,
+% Palisades NY 10964-8000 US,
+% E-mail: \href{mailto:KKey@ldeo.columbia.edu}{KKey@ldeo.columbia.edu}.
+}
 
 \author{%
-Dieter Werthmüller\footnotemark[1], %
-Kerry Key\footnotemark[2], and %
-Evert C.\ Slob\footnotemark[3]%
+% Dieter Werthmüller\footnotemark[1]\footnotemark[2], %
+% Kerry Key\footnotemark[3], and %
+% Evert C.\ Slob\footnotemark[1]%
 }
@@
 \footer{}
-\lefthead{Werthmüller et al.}
+\lefthead{}%Werthmüller et al.}
 \righthead{Digital filter designing tool}
@@
 \begin{equation}
-  b_n = \exp\left\{\Delta(-N/2+n) + \delta\right\} \ .
+  b_n = \exp\left\{\Delta(-\lfloor{(N+1)/2\rfloor}+n) + \delta\right\} \ .
   \label{eq:base}
@@
 %
-From equation~\ref{eq:filtapprox} it can be seen that the filter method
-requires $N$ evaluations at each $r$. For example, to calculate the frequency
-domain result for 100 offsets with a 201\,pt filter requires 20,100 evaluations
-in the wavenumber domain. This is why the DLF often uses interpolation to
-minimize the required evaluations, either for $F(r)$ in what is referred to as
-\emph{lagged convolution DLF}, or for $f(l)$, which we here call \emph{splined
-DLF}.
+\new{Note to reviewers: In equation~\ref{eq:base} $N/2$ changed to
+$\lfloor{(N+1)/2\rfloor}$.} From equation~\ref{eq:filtapprox} it can be
+seen that the filter method requires $N$ evaluations at each $r$. For example,
+to calculate the frequency domain result for 100 offsets with a 201\,pt filter
+requires 20,100 evaluations in the wavenumber domain. This is why the DLF often
+uses interpolation to minimize the required evaluations, either for $F(r)$ in
+what is referred to as \emph{lagged convolution DLF}, or for $f(l)$, which we
+here call \emph{splined DLF}.
 
 In the direct matrix inversion method for solving for the digital filter
-coefficients, equation~\ref{eq:filtapprox} is cast as a linear system as
-follows. $\Delta$ and $\delta$ are preassigned scalar values and $r$ is a range
-of $M$ preassigned values $r_m$. The filter base coefficients $b_n$ are
-computed using equation~\ref{eq:base} and an array of values for $l$ are
-computed using $l_{mn} = b_n/r_m$. The linear system's left hand side (LHS)
-matrix $\bf{A}$ has dimensions $M \times N$ with coefficients $A_{mn} =
-f(l_{mn})/r_n$ and the right hand side (RHS) vector $\bf{v}$ has elements $v_n
-= F(r_n)$. The resulting linear system
+coefficients, equation~\ref{eq:filtapprox} is cast as a linear system\old{ as
+follows.}\new{, where} $\Delta$ and $\delta$ are preassigned scalar values and
+$r$ is a range of $M$ preassigned values $r_m$. The filter base coefficients
+$b_n$ are computed using equation~\ref{eq:base} and an array of values for $l$
+are computed using $l_{mn} = b_n/r_m$. The linear system's left hand side (LHS)
+matrix $\bf{A}$ has dimensions $M \times N$ with coefficients \old{$A_{mn} =
+f(l_{mn})/r_n$}\new{$A_{mn} = f(l_{mn})/r_m$} and the right hand side (RHS)
+vector $\bf{v}$ has elements \old{$v_n = F(r_n)$}\new{$v_m = F(r_m)$}. The
+resulting linear system
 %
@@
   \item Evaluate RHS of check-function $f_\mr{C}$:\newline
-        $d_\mr{R} = f_\mr{C}.\mr{rhs}(r)$
-  \item Loop over each value $\Delta_i$, $\delta_j$ (brute force):
+    \old{$d_\mr{R} = f_\mr{C}.\mr{rhs}(r)$}\\
+    \new{$d_\mr{R} = f_\mr{C}^\mr{rhs}(r)$}
+  \item Loop over each value $\Delta_i$, $\delta_j$ (brute force\new{;
+      parameters on the left-hand side are overwritten where $i$, $j$
+      are not mentioned}):
     \begin{enumerate}
       \item Calculate filter base:\newline
-        $b_n = \exp\left\{\Delta_i(-N/2+n) + \delta_j\right\}$
+        \old{$b_n = \exp\left\{\Delta_i(-N/2+n) + \delta_j\right\}$}\\
+        \new{$b_n = \exp\left\{\Delta_i(-\lfloor{(N+1)/2\rfloor}+n) +
+      \delta_j\right\}$}
       \item Get required RHS and LHS evaluation points:\newline
@@
       \item Invert for filter coefficients:\newline
-        $J_{ij} = \mr{solve}(f_\mr{I}.\mr{lhs}(l_\mr{I}),
-        f_\mr{I}.\mr{rhs}(r_\mr{I}))$
+        \old{$J_{ij} = \mr{solve}(f_\mr{I}.\mr{lhs}(l_\mr{I}),
+        f_\mr{I}.\mr{rhs}(r_\mr{I}))$}\\
+        \new{$J = \mr{solve}\left[f_\mr{I}^\mr{lhs}(l_\mr{I}),
+        f_\mr{I}^\mr{rhs}(r_\mr{I})\right]$}
     \item Calculate numerically RHS of check-function $f_\mr{C}$ with current
-      filter $J_{ij}$:\newline
-      $d_\mr{F} = f_\mr{C}.\mr{lhs}(b/r)\cdot J_{ij}\ /\ r$
+      filter \old{$J_{ij}$}\new{$J$}:\newline
+      \old{$d_\mr{F} = f_\mr{C}.\mr{lhs}(b/r)\cdot J_{ij}\ /\ r$}\\
+      \new{$d_\mr{F} = f_\mr{C}^\mr{lhs}(b/r)\cdot J\ /\ r$}
       \item Store minimum recovered amplitude or maximum $r$ where relative
@@
     \end{enumerate}
-  \item Return filter coefficients which yield minimum amplitude or maximum $r$
-    (a local minimization can be run to polish the brute-force result):\newline
-    return $J\left[\mr{argmin}(\chi)\right] $\newline
+  \item \old{Return filter}\new{Re-calculate (Steps 2a-2c) filter base and}
+    coefficients which yield minimum amplitude or maximum $r$ (a local
+    minimization can be run to polish the brute-force result):\newline
+    \new{$i_\mr{best}$, $j_\mr{best}$ = $\mr{argmin}(\chi)$}\newline
+    \new{recalculate $b$, $J$ with $i_\mr{best}$, $j_\mr{best}$}\newline
+    \old{return $J\left[\mr{argmin}(\chi)\right] $}\newline
+  \item \new{Return filter base and coefficients and error matrix (for visual
+    QC):}\newline
+    \new{return $b$, $J$, $\chi$}\newline
 \end{enumerate}
@@
 1D subsurface models together with the EM-modeler \texttt{empymod}. The code is
-available from GitHub in the add-on repository of \texttt{empymod}, called
-\texttt{empyscripts}, from version 0.2.0 onwards.
+available from GitHub \old{in the add-on repository}\new{as part} of
+\texttt{empymod}\old{, called \texttt{empyscripts},} from version
+\old{0.2.0}\new{1.7.0} onwards.
 
@@
 
-D.W. would like to thank the electromagnetic research group of Aleksandr
-Mousatov at the IMP for fruitful discussions, and Jürg Hunziker for valuable
-feedback which improved this manuscript considerably.
+% D.W. would like to thank the electromagnetic research group of Aleksandr
+% Mousatov at the IMP for fruitful discussions, and Jürg Hunziker for valuable
+% feedback which improved this manuscript considerably.
+We thank the reviewers for valuable inputs not only regarding the manuscript,
+but also regarding the code and its documentation.
 

### 2018-06-10 ManuscriptGeophysics2018.tex
All answered; needs work on code/figs/manuscripts

@@
 arrivals at short time-offsets and the later arrivals had poor quality. Yet,
-this promising initial test suggested it might be possible to specifically
-design a special filter for GPR frequencies where the EM fields propagate as
-waves rather than by diffusion.
+this promising initial test \old{suggested}\new{suggests} it might be possible
+to specifically design a special filter for GPR frequencies where the EM fields
+propagate as waves rather than by diffusion.
 
@@
 therefore into all three mentioned categories by the new application of DLF to
-wave phenomena and by providing new filters with the accompanying code. The
-algorithm and many more examples of its usage can be found on
+wave phenomena, \new{which means that the DLF can be used to model seismic
+wavefields as well,} and by providing new filters with the accompanying code.
+The algorithm and many more examples of its usage can be found on
 \href{https://github.com/empymod/article-fdesign}{github.com/empymod/article-fdesign}.
@@
 %
-can be solved by direct matrix inversion to obtain the vector of filter
-coefficients $\bf{h}$. For a given filter length $N$, there are many subjective
-choices that go into the practical implementation of this method, including the
-choice of the transform pairs $F(r)$ and $f(l)$, as well as the value for $M$
-and the particular range and spacing of the values $r_m$. Once values are
-chosen for these variables, an optimal filter can be found by a grid search
-over $\Delta$ and $\delta$ for values that produce a high quality filter. The
-choice of metric for what constitutes a high quality filter is also subjective
-and is further discussed below.
+can be solved by direct matrix inversion\new{, or any other matrix inversion
+routine,} to obtain the vector of filter coefficients $\bf{h}$. For a given
+filter length $N$, there are many subjective choices that go into the practical
+implementation of this method, including the choice of the transform pairs
+$F(r)$ and $f(l)$, as well as the value for $M$ and the particular range and
+spacing of the values $r_m$. Once values are chosen for these variables, an
+optimal filter can be found by a grid search over $\Delta$ and $\delta$ for
+values that produce a high quality filter. The choice of metric for what
+constitutes a high quality filter is also subjective and is further discussed
+below.
 

### 2018-06-18 ManuscriptGeophysics2018.tex
In revision; move from shelve to plain txt files

@@
 therefore into all three mentioned categories by the new application of DLF to
-wave phenomena, \new{which means that the DLF can be used to model seismic
-wavefields as well,} and by providing new filters with the accompanying code.
-The algorithm and many more examples of its usage can be found on
+wave phenomena, \new{which means that the DLF can probably be used to model
+seismic wavefields as well,} and by providing new filters with the accompanying
+code. The algorithm and many more examples of its usage can be found on
 \href{https://github.com/empymod/article-fdesign}{github.com/empymod/article-fdesign}.
@@
     results.
-  \item The algorithm can solve under-, equal-, and over-determined linear
-    systems.
+  \item The algorithm can solve \old{under-,} equal-\old{,} and over-determined
+    linear systems.
   \item The real or the imaginary part can be used for the inversion of complex
@@
 
+\new{To solve step 2(c) we implemented the QR-decomposition method in
+  \texttt{fdesign}.}
+
 \subsection{Minimization criteria}
@@
 minimization approaches; in (a) for a conventional, fast decaying transform
-pair; and in (b) for a more complex, high frequency layered model. The circles
-show the minimum amplitude used in previous approaches. This criterion is not
-ideal, as it is subject to some random fluctuations and depends on the choice
-of $r$. The squares show the minimum amplitude given a certain acceptable
-error, and the diamonds show the maximum $r$ given a certain acceptable error
-(the inversion is a minimization process, it therefore minimizes $1/r$, not
-$r$). In simple cases the minimum amplitude given an acceptable error and the
-maximum $r$ given an acceptable error will yield the same result. However, in
-complex cases the maximum $r$ is more consistent and therefore a much better
-criterion.
+pair \new{(using the $J_0$ kernel of a fullspace with vertical distance between
+source and receiver of 50\,m, frequency $f=1\,$Hz, and resistivity
+$\rho=1\,\Omega\,$m)}; and in (b) for a more complex, high frequency layered
+model \new{(using the $J_0$ kernel for $f=200\,$MHz for the GPR model described
+in the section \emph{Ground-penetrating radar}, with receiver located at 1\,m
+depth)}. The circles show the minimum amplitude used in previous approaches.
+This criterion is not ideal, as it is subject to some random fluctuations and
+depends on the choice of $r$. The squares show the minimum amplitude given a
+certain acceptable error, and the diamonds show the maximum $r$ given a certain
+acceptable error (the inversion is a minimization process, it therefore
+minimizes $1/r$, not $r$). In simple cases the minimum amplitude given an
+acceptable error and the maximum $r$ given an acceptable error will yield the
+same result. However, in complex cases the maximum $r$ is more consistent and
+therefore a much better criterion.
 %
@@
 \subsection{Design}
-Figure~\ref{fig:201} shows the solution spaces of four consecutive inversion
-runs, where each run is a focus on a subsection of the previous run, indicated
-by the red square.
-%
-\plot*{201}{width=\textwidth}{%
-  Solution spaces of four consecutive inversion runs for a 201\,pt filter. Each
-  consecutive run zooms into a portion of the previous solution space,
-  indicated by the red square. The more in detail we obtain the solution space,
-  the more random appears to be the distribution.}%
-%
-The filter length in this case is $N=201$, and the theoretical transform pairs
-are as given, e.g., in \cite{USGS.75.Anderson}:
+Figure~\ref{fig:201} shows the solution spaces of \old{four}\new{eight}
+consecutive inversion runs\old{, where each run is a focus on a subsection of
+  the previous run, indicated by the red square. The filter length in this case
+is $N=201$, and}\new{where} the theoretical transform pairs
+\old{are} as given, e.g., in \cite{USGS.75.Anderson}\old{:}\new{,}
 %
@@
 %
-where $a$ was set to 5. In the algorithm you can provide one pair for the
-inversion, and a different pair to get the minimum amplitude or the maximum
-$r$. In this case the same transform pair was used for both the inversion and
-the check of quality. The acceptable relative error was set to 1\,\%. The RHS
-evaluation parameter for the inversion was set so  that an over-determined
-system was evaluated with $M = 2N$ equations, where $r_\mr{I}$ was
-logarithmically spaced from $\log_{10}(1/\max(b)) - 1$ to $\log_{10}(1/\min(b))
-+ 1$, where $b$ are the filter abscissae as given in equation~\ref{eq:base}.
-This is accomplished in \texttt{fdesign} by setting parameter $r_\mr{def} = (1,
-1, 2)$.
-
-From the low resolution overview runs (a) and (b) it looks like this is a
-standard, straight-forward minimization problem. However, from the more
-detailed results in (c) and (d) it becomes obvious that it is a minimization
+%
+\plot*{201}{width=\textwidth}{%
+  Solution spaces of four consecutive inversion runs for a 201\,pt filter. Each
+  consecutive run zooms into a portion of the previous solution space,
+  indicated by the red square. The more in detail we obtain the solution space,
+  the more random appears to be the distribution.}%
+%
+where \new{used; } $a$ was set to 5. In the algorithm you can provide one pair
+for the inversion, and a different pair to get the minimum amplitude or the
+maximum $r$. In this case the same transform pair was used for both the
+inversion and the check of quality. The acceptable relative error was set to
+1\,\%. The RHS evaluation parameter for the inversion was set so  that an
+over-determined system was evaluated with $M = 2N$ equations, where $r_\mr{I}$
+was logarithmically spaced from $\log_{10}(1/\max(b)) - 1$ to
+$\log_{10}(1/\min(b)) + 1$, where $b$ are the filter abscissae as given in
+equation~\ref{eq:base}. This is accomplished in \texttt{fdesign} by setting
+parameter $r_\mr{def} = (1, 1, 2)$.
+
+\new{The top row shows in subplots (a) to (d) four different inversion runs
+for $N=51$, 101, 201, and 401 for a wide range of spacing and shift values.
+From this one has to decide for a filter length. Longer filters tend to give
+more precise results, but their transform is computationally more expensive as
+more values have to be evaluated. The length of the filter does not has to be
+an odd number, but many published filters do have an odd number of
+coefficients. The integral runs from $-\infty$ to $+\infty$ and hence an odd
+point filter comes naturally by truncating the integral to a finite sum from
+$-(N-1)/2$ to $(N-1)/2$. However, this is only true if the shift value is
+$\delta = 0$. If your shift value is not zero then there is no advantage in
+having an odd point filter.}
+
+\new{The bottom row shows four consecutive inversion runs for $N=201$, where
+each run is a focus on a subsection of the previous run, indicated by the red
+square.} From the low resolution overview runs (e) and (f) it looks like this
+is a standard, straight-forward minimization problem. However, from the more
+detailed results in (g) and (h) it becomes obvious that it is a minimization
 problem that has to be solved stochastically, as there are solutions with 2
@@
 influence the outcome, and often a good filter is found by sheer luck of
-choosing the right starting parameters. Each of the input parameters has an
-effect on the outcome. It also depends heavily on the transform pairs
-$f_\mr{I;C}$; functions that decay rapidly are generally better, as noted by
-earlier authors \citep[e.g.][]{USGS.75.Anderson}. It also depends on the filter
-length, and obviously heavily on the spacing and shift values. Another
-important point is how you define the right-hand evaluation points of the
-inversion ($r_\mr{def}$). Evaluating corresponding transform pairs separately
-or jointly also leads to different filter coefficients ($J_0$, $J_1$, or $J_0$
-\& $J_1$, or equally sine, cosine, or sine \& cosine); also if the real or the
-imaginary part is used when complex transform pairs or \texttt{empymod} is
-used; and if it is inverted for the minimum amplitude or for the maximum $r$.
+choosing the right starting parameters\new{, as can be seen in
+Figure~\ref{fig:201b} (h). However, once you are in this zone every filter is a
+very good filter}. Each of the input parameters has an effect on the outcome.
+It also depends heavily on the transform pairs $f_\mr{I;C}$; functions that
+decay rapidly are generally better, as noted by earlier authors
+\citep[e.g.][]{USGS.75.Anderson}. It also depends on the filter length, and
+obviously heavily on the spacing and shift values. Another important point is
+how you define the right-hand evaluation points of the inversion
+($r_\mr{def}$). Evaluating corresponding transform pairs separately or jointly
+also leads to different filter coefficients ($J_0$, $J_1$, or $J_0$ \& $J_1$,
+or equally sine, cosine, or sine \& cosine); also if the real or the imaginary
+part is used when complex transform pairs or \texttt{empymod} is used; and if
+it is inverted for the minimum amplitude or for the maximum $r$.
 
@@
   frequencies (yellow zones). However, the amplitudes in these regions are in
-  the order of $10^{-25}\,$V/m, and therefore far below the noise level.}
+  the order of $10^{-25}\,$V/m, and therefore \old{far}\new{several orders of
+  magnitude} below the noise level\new{s of current EM instrument systems}.}
 %

### 2018-06-18 notebooks/create_binaries.py

@@ -1,67 +1,0 @@
-from numpy import savetxt, loadtxt
-from empymod.filters import DigitalFilter
-from empyscripts.fdesign import load_filter, save_filter
-
-
-def shelved2ascii(name):
-    """Store filter and inversion output in plain ASCII files.
-
-    Geophysics doesn't permit binary files to be distributed. So we load
-    the information stored with `empyscripts.fdesign.save_filter` and store
-    it in plain ASCII files.
-
-    """
-    # Base-name for files
-    base = './filters/'+name
-
-    # Get the filter
-    filt, out = load_filter(name, True)
-
-    # Store arrays
-    savetxt(base+'-filt-base.txt', filt.base)
-    savetxt(base+'-filt-j0.txt', filt.j0)
-    savetxt(base+'-filt-j1.txt', filt.j1)
-    savetxt(base+'-out-2a.txt', out[2][0])
-    savetxt(base+'-out-2b.txt', out[2][1])
-    savetxt(base+'-out-3.txt', out[3])
-
-    # Store scalars in one file
-    savetxt(base+'-scalars.txt',
-            [filt.factor, out[0][0], out[0][1], out[1]],
-            header=filt.name + '; filt.factor; out[0][1]; out[0][1]; out[1]')
-
-
-def ascii2shelved(name):
-    """Convert ASCII-files to empyscripts.fdesign.save_filter-object.
-
-    Geophysics doesn't permit binary files to be distributed. So we load
-    the information stored in plain ASCII files, and create the required
-    binaries with `empyscripts.fdesign.save_filter`.
-
-    """
-    # Base-name of files
-    base = './filters/'+name
-
-    # Get scalars
-    numbers = loadtxt(base+'-scalars.txt')
-
-    # Collect inversion-result tuple
-    out = ([numbers[1], numbers[2]],
-           numbers[3],
-           [loadtxt(base+'-out-2a.txt'), loadtxt(base+'-out-2b.txt')],
-           loadtxt(base+'-out-3.txt'))
-
-    # Collect digital filter
-    dlf = DigitalFilter(name)
-    dlf.base = loadtxt(base+'-filt-base.txt')
-    dlf.factor = numbers[0]
-    dlf.j0 = loadtxt(base+'-filt-j0.txt')
-    dlf.j1 = loadtxt(base+'-filt-j1.txt')
-
-    # Save the filter and the inversion result
-    save_filter(name, dlf, out)
-
-
[…]

### 2018-06-19 ManuscriptGeophysics2018.tex
Notebooks ready for re-submission

@@
 % \documentclass[paper,twocolumn,twoside]{geophysics}
-% \documentclass[paper]{geophysics}
+\documentclass[paper]{geophysics}
 % \documentclass[paper, revised]{geophysics}
-\documentclass[manuscript,revised]{geophysics}
+% \documentclass[manuscript,revised]{geophysics}
 
@@
 \subsection{Design}
-Figure~\ref{fig:201} shows the solution spaces of \old{four}\new{eight}
-consecutive inversion runs\old{, where each run is a focus on a subsection of
-  the previous run, indicated by the red square. The filter length in this case
-is $N=201$, and}\new{where} the theoretical transform pairs
-\old{are} as given, e.g., in \cite{USGS.75.Anderson}\old{:}\new{,}
+\old{Figure~\ref{fig:201b} shows the solution spaces of four consecutive
+inversion runs, where each run is a focus on a subsection of the previous run,
+indicated by the red square. The filter length in this case is $N=201$, and the
+theoretical transform pairs are as given, e.g., in
+\cite{USGS.75.Anderson}:}
+\new{Figures~\ref{fig:201a} and~\ref{fig:201b} show the solution spaces of
+seven consecutive inversion runs using the following theoretical transform
+pairs \citep[e.g.,][]{USGS.75.Anderson},}
 %
@@
 %
-%
-\plot*{201}{width=\textwidth}{%
-  Solution spaces of four consecutive inversion runs for a 201\,pt filter. Each
-  consecutive run zooms into a portion of the previous solution space,
-  indicated by the red square. The more in detail we obtain the solution space,
-  the more random appears to be the distribution.}%
-%
-where \new{used; } $a$ was set to 5. In the algorithm you can provide one pair
+where $a$ was set to 5. In the algorithm you can provide one pair
 for the inversion, and a different pair to get the minimum amplitude or the
@@
 over-determined system was evaluated with $M = 2N$ equations, where $r_\mr{I}$
-was logarithmically spaced from $\log_{10}(1/\max(b)) - 1$ to
-$\log_{10}(1/\min(b)) + 1$, where $b$ are the filter abscissae as given in
+was logarithmically spaced from $\log_{10}[1/\max(b)] - 1$ to
+$\log_{10}[1/\min(b)] + 1$, where $b$ are the filter abscissae as given in
 equation~\ref{eq:base}. This is accomplished in \texttt{fdesign} by setting
@@
 
-\new{The top row shows in subplots (a) to (d) four different inversion runs
-for $N=51$, 101, 201, and 401 for a wide range of spacing and shift values.
+\new{Figure~\ref{fig:201a} shows three different inversion runs for $N=101$,
+201, and 401 for a wide range of spacing and shift values.
 From this one has to decide for a filter length. Longer filters tend to give
@@
 point filter comes naturally by truncating the integral to a finite sum from
-$-(N-1)/2$ to $(N-1)/2$. However, this is only true if the shift value is
-$\delta = 0$. If your shift value is not zero then there is no advantage in
-having an odd point filter.}
-
-\new{The bottom row shows four consecutive inversion runs for $N=201$, where
-each run is a focus on a subsection of the previous run, indicated by the red
-square.} From the low resolution overview runs (e) and (f) it looks like this
-is a standard, straight-forward minimization problem. However, from the more
-detailed results in (g) and (h) it becomes obvious that it is a minimization
-problem that has to be solved stochastically, as there are solutions with 2
-orders of magnitude difference apparently randomly next to each other.
-
-Figure~\ref{fig:201b} shows in (a) the filter values for $J_0$ and $J_1$ of the
+$-(N-1)/2$ to $(N-1)/2$. However, this is only true if the shift value
+$\delta = 0$. If it differs from zero then there is no advantage in having an
+odd point filter.}
+%
+\plot*{201a}{width=\textwidth}{%
+  \new{Test of different filter lengths. Filter length is a trade-off between
+    precision and speed, as shorter filters are faster but less precise, and
+  longer filters are more precise but computationally more expensive. The red
+  square in subplot (b) is the starting point of Figure~\ref{fig:201b}.} }%
+%
+
+\new{Figure~\ref{fig:201b} shows four consecutive inversion runs for $N=201$,
+where each run is a focus on a subsection of the previous run, indicated by the
+red square.} From the low resolution overview runs (a) and (b) it looks like
+this is a standard, straight-forward minimization problem. However, from the
+more detailed results in (c) and (d) it becomes obvious that it is a
+minimization problem that has to be solved stochastically, as there are
+solutions with 2 orders of magnitude difference apparently randomly next to
+each other.
+%
+\plot*{201b}{width=\textwidth}{%
+  Solution spaces of four consecutive inversion runs for a 201\,pt filter. Each
+  consecutive run zooms into a portion of the previous solution space,
+  indicated by the red square. The more in detail we obtain the solution space,
+  the more random appears to be the distribution.}%
+%
+
+Figure~\ref{fig:201c} shows in (a) the filter values for $J_0$ and $J_1$ of the
 best 201 point filter obtained this way, and in (b) its right-hand-side
@@
 %
-\plot*{201b}{width=.9\textwidth}{(a) Filter values of the best obtained 201\,pt
+\plot*{201c}{width=.9\textwidth}{(a) Filter values of the best obtained 201\,pt
   filter with the corresponding check of quality in (b). Black points indicate
@@
 Designing filters is to a large extent trial and error. All input variables
-influence the outcome, and often a good filter is found by sheer luck of
-choosing the right starting parameters\new{, as can be seen in
-Figure~\ref{fig:201b} (h). However, once you are in this zone every filter is a
+influence the outcome, and often a \old{good}\new{even better} filter is found
+by sheer luck of choosing the right starting parameters\new{, as can be seen in
+Figure~\ref{fig:201c} (h). However, once you are in this zone every filter is a
 very good filter}. Each of the input parameters has an effect on the outcome.
@@
 %
-\plot*{kongf5}{width=.9\textwidth}{(a) Results of different filters and QWE for
+\plot*{kongf5}{width=\textwidth}{(a) Results of different filters and QWE for
   the model of Figure~5 from \cite{GP.07.Kong} with the relative errors shown
@@
 %
-\plot*{keyf5}{width=.9\textwidth}{(a) The canonical marine CSEM model of
+\plot*{keyf5}{width=\textwidth}{(a) The canonical marine CSEM model of
   \cite{GEO.12.Key} and (b) a land model with source and receiver at the
@@
 for all practical CSEM cases, hence has generally an error far below 1\,\%
-($10^{-2}$) in all cases. The exception are large offsets with relatively high
-frequencies, e.g., 10\,km offset with 8\,Hz. In the shallow water case this
-zone has a very high error for subsurface resistivities of 1\,$\Omega\,$m and
+($10^{-2}$) in all cases. \new{A relative noise of less than 1\,\% is generally
+considered really good for real data.} The exception\new{s} are \new{in the
+shallow water case at} large offsets with relatively high frequencies, e.g.,
+10\,km offset with 8\,Hz\old{. In the shallow water case}\new{;} this zone has
+a very high error for subsurface resistivities of 1\,$\Omega\,$m and
 10\,$\Omega\,$m (yellow zones). However, this is due to the fact that the
 amplitudes of the signal for these offsets and frequencies are in the order of
-$10^{-25}\,$V/m.
+$10^{-25}\,$V/m. \new{The grey contour lines in the plot indicate the power of
+  the amplitude of the signal.}
 %
@@
   the order of $10^{-25}\,$V/m, and therefore \old{far}\new{several orders of
-  magnitude} below the noise level\new{s of current EM instrument systems}.}
+  magnitude} below the noise level\new{s of current EM instrument systems}.
+  \new{Contour lines show the power of the amplitude of the signal.}}
 %

### 2018-06-19 ManuscriptGeophysics2018.tex
Manuscript re-submission

@@
 % \documentclass[manuscript]{geophysics}
-% \documentclass[paper,twocolumn,twoside]{geophysics}
-\documentclass[paper]{geophysics}
+\documentclass[paper,twocolumn,twoside]{geophysics}
+% \documentclass[paper]{geophysics}
 % \documentclass[paper, revised]{geophysics}
@@
 \address{
-% \footnotemark[1]TU Delft,
-% Building 23,
-% Stevinweg 1 / PO-box 5048,
-% 2628 CN Delft,
-% E-mail: \href{mailto:E.C.Slob@tudelft.nl}{E.C.Slob@tudelft.nl};
-% \footnotemark[2]Instituto Mexicano del Petróleo,
-% Eje Central Lázaro Cárdenas Norte 152,
-% Col. San Bartolo Atepehuacan C.P. 07730,
-% Ciudad de México, México,
-% E-mail: \href{mailto:dieter@werthmuller.org}{Dieter@Werthmuller.org};
-% \footnotemark[3]Lamont-Doherty Earth Observatory, Columbia University,
-% 305C Oceanography,
-% 61 Route 9W, PO Box 1000,
-% Palisades NY 10964-8000 US,
-% E-mail: \href{mailto:KKey@ldeo.columbia.edu}{KKey@ldeo.columbia.edu}.
+\footnotemark[1]TU Delft,
+Building 23,
+Stevinweg 1 / PO-box 5048,
+2628 CN Delft,
+E-mail: \href{mailto:dieter@werthmuller.org}{Dieter@Werthmuller.org};
+\href{mailto:E.C.Slob@tudelft.nl}{E.C.Slob@tudelft.nl};
+\footnotemark[2]Instituto Mexicano del Petróleo,
+Eje Central Lázaro Cárdenas Norte 152,
+Col. San Bartolo Atepehuacan C.P. 07730,
+Ciudad de México, México,
+\footnotemark[3]Lamont-Doherty Earth Observatory, Columbia University,
+305C Oceanography,
+61 Route 9W, PO Box 1000,
+Palisades NY 10964-8000 US,
+E-mail: \href{mailto:KKey@ldeo.columbia.edu}{KKey@ldeo.columbia.edu}.
 }
@@
 \author{%
-% Dieter Werthmüller\footnotemark[1]\footnotemark[2], %
-% Kerry Key\footnotemark[3], and %
-% Evert C.\ Slob\footnotemark[1]%
+Dieter Werthmüller\footnotemark[1]\footnotemark[2], %
+Kerry Key\footnotemark[3], and %
+Evert C.\ Slob\footnotemark[1]%
 }
@@
 \footer{}
-\lefthead{}%Werthmüller et al.}
+\lefthead{Werthmüller et al.}
 \righthead{Digital filter designing tool}
@@
 %
-\new{Note to reviewers: In equation~\ref{eq:base} $N/2$ changed to
-$\lfloor{(N+1)/2\rfloor}$.} From equation~\ref{eq:filtapprox} it can be
-seen that the filter method requires $N$ evaluations at each $r$. For example,
-to calculate the frequency domain result for 100 offsets with a 201\,pt filter
-requires 20,100 evaluations in the wavenumber domain. This is why the DLF often
-uses interpolation to minimize the required evaluations, either for $F(r)$ in
-what is referred to as \emph{lagged convolution DLF}, or for $f(l)$, which we
-here call \emph{splined DLF}.
+From equation~\ref{eq:filtapprox} it can be seen that the filter method
+requires $N$ evaluations at each $r$. For example, to calculate the frequency
+domain result for 100 offsets with a 201\,pt filter requires 20,100 evaluations
+in the wavenumber domain. This is why the DLF often uses interpolation to
+minimize the required evaluations, either for $F(r)$ in what is referred to as
+\emph{lagged convolution DLF}, or for $f(l)$, which we here call \emph{splined
+DLF}.
 
@@
     \end{enumerate}
-  \item \old{Return filter}\new{Re-calculate (Steps 2a-2c) filter base and}
+  \item \old{Return filter}\new{Re-calculate (steps 2a-2c) filter base and}
     coefficients which yield minimum amplitude or maximum $r$ (a local
@@
 theoretical transform pairs are as given, e.g., in
-\cite{USGS.75.Anderson}:}
+}\cite{USGS.75.Anderson}\old{:}
 \new{Figures~\ref{fig:201a} and~\ref{fig:201b} show the solution spaces of
 seven consecutive inversion runs using the following theoretical transform
-pairs \citep[e.g.,][]{USGS.75.Anderson},}
+pairs} \citep[e.g.,][]{USGS.75.Anderson}\new{,}
 %
@@
 Designing filters is to a large extent trial and error. All input variables
-influence the outcome, and often a \old{good}\new{even better} filter is found
-by sheer luck of choosing the right starting parameters\new{, as can be seen in
-Figure~\ref{fig:201c} (h). However, once you are in this zone every filter is a
-very good filter}. Each of the input parameters has an effect on the outcome.
-It also depends heavily on the transform pairs $f_\mr{I;C}$; functions that
-decay rapidly are generally better, as noted by earlier authors
+influence the outcome, and often \old{a good}\new{an even better} filter is
+found by sheer luck of choosing the right starting parameters\new{, as can be
+seen in Figure~\ref{fig:201c} (h). However, once you are in this zone every
+filter is a very good filter}. Each of the input parameters has an effect on
+the outcome. It also depends heavily on the transform pairs $f_\mr{I;C}$;
+functions that decay rapidly are generally better, as noted by earlier authors
 \citep[e.g.][]{USGS.75.Anderson}. It also depends on the filter length, and
@@
 available from GitHub \old{in the add-on repository}\new{as part} of
-\texttt{empymod}\old{, called \texttt{empyscripts},} from version
-\old{0.2.0}\new{1.7.0} onwards.
+\texttt{empymod}\old{, called \texttt{empyscripts},}\new{; the version
+  corresponding to this publication is \texttt{v1.7.1.}}\old{ from version
+  0.2.0 onwards}.
 
@@
 
-% D.W. would like to thank the electromagnetic research group of Aleksandr
-% Mousatov at the IMP for fruitful discussions, and Jürg Hunziker for valuable
-% feedback which improved this manuscript considerably.
-We thank the reviewers for valuable inputs not only regarding the manuscript,
-but also regarding the code and its documentation.
+D.W. would like to thank the electromagnetic research group of Aleksandr
+Mousatov at the IMP for fruitful discussions, and Jürg Hunziker for valuable
+feedback which improved this manuscript considerably. We thank the reviewers
+for valuable inputs not only regarding the manuscript, but also regarding the
+code and its documentation.
 

### 2018-11-21 ManuscriptGeophysics2018.tex
Move // inside \old{}
Within preparation for final manuscript

@@
 
-%%fakesection ===    PACKAGES & DEFINITIONS    ===
-% \usepackage{showframe}
-
-%% ~ ADDITIONAL PACKAGES TO GEOPHYSICS.CLS
+% Additional packages to geophysics.cls
 \usepackage[USenglish]{babel}
@@
 \usepackage[strings]{underscore}
-\usepackage[pdftex, final]{hyperref}
-\hypersetup{allcolors=blue, allbordercolors={0 0 .5}, colorlinks=true}
+% \usepackage[pdftex, final]{hyperref}
+% \hypersetup{allcolors=blue, allbordercolors={0 0 .5}, colorlinks=true}
+\usepackage[pdftex, hidelinks]{hyperref}
 
@@
 % Own commands
-\usepackage{xspace}
-\newcommand{\empymod}{\texttt{empymod}\xspace}
 \newcommand{\mr}[1]{\mathrm{#1}}
@@
 2628 CN Delft,
-E-mail: \href{mailto:dieter@werthmuller.org}{Dieter@Werthmuller.org};
+E-mail: \href{mailto:D.Werthmuller@tudelft.nl}{D.Werthmuller@tudelft.nl};
 \href{mailto:E.C.Slob@tudelft.nl}{E.C.Slob@tudelft.nl};
@@
 Fourier sine/cosine transform), as all higher Bessel functions can be
-rewritten, via recurrence relations, using only use these two.
+rewritten, via recurrence relations, using only\old{use} these two.
 \cite{GP.94.Mohsen} is one of the rare cases which provides $J_2$ filter
@@
 Nevertheless, \cite{GEO.17.Werthmuller} tested DLF for modeling 250 MHz
-center-frequency GPR data using a 401 point filter that was designed for
-diffusive EM, and found that it was orders of magnitude faster than quadrature
-approaches. However, good quality results were only obtained for the first
-arrivals at short time-offsets and the later arrivals had poor quality. Yet,
-this promising initial test \old{suggested}\new{suggests} it might be possible
-to specifically design a special filter for GPR frequencies where the EM fields
+center-frequency GPR data using a 401 \old{point}\new{pt} filter that was
+designed for diffusive EM, and found that it was orders of magnitude faster
+than quadrature approaches. However, good quality results were only obtained
+for the first arrivals at short time-offsets and the later arrivals had poor
+quality. Yet, this promising initial test suggests it might be possible to
+specifically design a special filter for GPR frequencies where the EM fields
 propagate as waves rather than by diffusion.
@@
 therefore into all three mentioned categories by the new application of DLF to
-wave phenomena, \new{which means that the DLF can probably be used to model
-seismic wavefields as well,} and by providing new filters with the accompanying
+wave phenomena, which means that the DLF can probably be used to model
+seismic wavefields as well, and by providing new filters with the accompanying
 code. The algorithm and many more examples of its usage can be found on
@@
     results.
-  \item The algorithm can solve \old{under-,} equal-\old{,} and over-determined
+  \item The algorithm can solve equal- and over-determined
     linear systems.
@@
 In the direct matrix inversion method for solving for the digital filter
-coefficients, equation~\ref{eq:filtapprox} is cast as a linear system\old{ as
-follows.}\new{, where} $\Delta$ and $\delta$ are preassigned scalar values and
+coefficients, equation~\ref{eq:filtapprox} is cast as a linear system,
+where $\Delta$ and $\delta$ are preassigned scalar values and
 $r$ is a range of $M$ preassigned values $r_m$. The filter base coefficients
@@
 are computed using $l_{mn} = b_n/r_m$. The linear system's left hand side (LHS)
-matrix $\bf{A}$ has dimensions $M \times N$ with coefficients \old{$A_{mn} =
-f(l_{mn})/r_n$}\new{$A_{mn} = f(l_{mn})/r_m$} and the right hand side (RHS)
-vector $\bf{v}$ has elements \old{$v_n = F(r_n)$}\new{$v_m = F(r_m)$}. The
+matrix $\bf{A}$ has dimensions $M \times N$ with coefficients
+$A_{mn} = f(l_{mn})/r_m$ and the right hand side (RHS)
+vector $\bf{v}$ has elements $v_m = F(r_m)$. The
 resulting linear system
@@
 %
-can be solved by direct matrix inversion\new{, or any other matrix inversion
-routine,} to obtain the vector of filter coefficients $\bf{h}$. For a given
+can be solved by direct matrix inversion, or any other matrix inversion
+routine, to obtain the vector of filter coefficients $\bf{h}$. For a given
 filter length $N$, there are many subjective choices that go into the practical
@@
   \item Evaluate RHS of check-function $f_\mr{C}$:\newline
-    \old{$d_\mr{R} = f_\mr{C}.\mr{rhs}(r)$}\\
-    \new{$d_\mr{R} = f_\mr{C}^\mr{rhs}(r)$}
-  \item Loop over each value $\Delta_i$, $\delta_j$ (brute force\new{;
+    $d_\mr{R} = f_\mr{C}^\mr{rhs}(r)$
+  \item Loop over each value $\Delta_i$, $\delta_j$ (brute force;
       parameters on the left-hand side are overwritten where $i$, $j$
-      are not mentioned}):
+      are not mentioned):
     \begin{enumerate}
       \item Calculate filter base:\newline
-        \old{$b_n = \exp\left\{\Delta_i(-N/2+n) + \delta_j\right\}$}\\
-        \new{$b_n = \exp\left\{\Delta_i(-\lfloor{(N+1)/2\rfloor}+n) +
-      \delta_j\right\}$}
+        $b_n = \exp\left\{\Delta_i(-\lfloor{(N+1)/2\rfloor}+n) +
+      \delta_j\right\}$
       \item Get required RHS and LHS evaluation points:\newline
@@
       \item Invert for filter coefficients:\newline
-        \old{$J_{ij} = \mr{solve}(f_\mr{I}.\mr{lhs}(l_\mr{I}),
-        f_\mr{I}.\mr{rhs}(r_\mr{I}))$}\\
-        \new{$J = \mr{solve}\left[f_\mr{I}^\mr{lhs}(l_\mr{I}),
-        f_\mr{I}^\mr{rhs}(r_\mr{I})\right]$}
+        $J = \mr{solve}\left[f_\mr{I}^\mr{lhs}(l_\mr{I}),
+        f_\mr{I}^\mr{rhs}(r_\mr{I})\right]$
     \item Calculate numerically RHS of check-function $f_\mr{C}$ with current
-      filter \old{$J_{ij}$}\new{$J$}:\newline
-      \old{$d_\mr{F} = f_\mr{C}.\mr{lhs}(b/r)\cdot J_{ij}\ /\ r$}\\
-      \new{$d_\mr{F} = f_\mr{C}^\mr{lhs}(b/r)\cdot J\ /\ r$}
+      filter $J$:\newline
+      $d_\mr{F} = f_\mr{C}^\mr{lhs}(b/r)\cdot J\ /\ r$
       \item Store minimum recovered amplitude or maximum $r$ where relative
@@
     \end{enumerate}
-  \item \old{Return filter}\new{Re-calculate (steps 2a-2c) filter base and}
+  \item Re-calculate (steps 2a-2c) filter base and
     coefficients which yield minimum amplitude or maximum $r$ (a local
     minimization can be run to polish the brute-force result):\newline
-    \new{$i_\mr{best}$, $j_\mr{best}$ = $\mr{argmin}(\chi)$}\newline
-    \new{recalculate $b$, $J$ with $i_\mr{best}$, $j_\mr{best}$}\newline
-    \old{return $J\left[\mr{argmin}(\chi)\right] $}\newline
-  \item \new{Return filter base and coefficients and error matrix (for visual
-    QC):}\newline
-    \new{return $b$, $J$, $\chi$}\newline
+    $i_\mr{best}$, $j_\mr{best}$ = $\mr{argmin}(\chi)$\newline
+    recalculate $b$, $J$ with $i_\mr{best}$, $j_\mr{best}$\newline
+    \newline
+  \item Return filter base and coefficients and error matrix (for visual
+    QC):\newline
+    return $b$, $J$, $\chi$\newline
 \end{enumerate}
 
-\new{To solve step 2(c) we implemented the QR-decomposition method in
-  \texttt{fdesign}.}
+To solve step 2(c) we implemented the QR-decomposition method in
+  \texttt{fdesign}.
 
 \subsection{Minimization criteria}
-Figure~\ref{fig:InvCrit} shows the differences between the different
+Figure~\ref{fig:Figure-1-bw} shows the differences between the different
 minimization approaches; in (a) for a conventional, fast decaying transform
-pair \new{(using the $J_0$ kernel of a fullspace with vertical distance between
+pair (using the $J_0$ kernel of a fullspace with vertical distance between
 source and receiver of 50\,m, frequency $f=1\,$Hz, and resistivity
-$\rho=1\,\Omega\,$m)}; and in (b) for a more complex, high frequency layered
-model \new{(using the $J_0$ kernel for $f=200\,$MHz for the GPR model described
+$\rho=1\,\Omega\,$m); and in (b) for a more complex, high frequency layered
+model (using the $J_0$ kernel for $f=200\,$MHz for the GPR model described
 in the section \emph{Ground-penetrating radar}, with receiver located at 1\,m
-depth)}. The circles show the minimum amplitude used in previous approaches.
+depth). The circles show the minimum amplitude used in previous approaches.
 This criterion is not ideal, as it is subject to some random fluctuations and
@@
 %
-\plot*{InvCrit}{width=.9\textwidth}{(a) A regular RHS curve of a transform pair
-  with a purely decaying function. Minimum amplitude or maximum $r$ yield the
-  same result in this scenario. Using a relative error criteria is more stable
-  than just the absolute minimum amplitude. (b) A 1D model for $f=200\,$MHz.
-  Here, the maximum $r$ provides a better criteria for the inversion.}
+\plot*{Figure-1-bw}{width=.9\textwidth}{(a) A regular \new{right-hand-side
+  (}RHS\new{)} curve of a transform pair with a purely decaying function.
+  Minimum amplitude or maximum $r$ yield the same result in this scenario.
+  Using a relative error criteria is more stable than just the absolute minimum
+  amplitude. (b) A 1D model for $f=200\,$MHz. Here, the maximum $r$ provides a
+  better criteria for the inversion.}
 %
@@
 \subsection{Design}
-\old{Figure~\ref{fig:201b} shows the solution spaces of four consecutive
-inversion runs, where each run is a focus on a subsection of the previous run,
-indicated by the red square. The filter length in this case is $N=201$, and the
-theoretical transform pairs are as given, e.g., in
-}\cite{USGS.75.Anderson}\old{:}
-\new{Figures~\ref{fig:201a} and~\ref{fig:201b} show the solution spaces of
+Figures~\ref{fig:Figure-2} and~\ref{fig:Figure-3} show the solution spaces of
 seven consecutive inversion runs using the following theoretical transform
-pairs} \citep[e.g.,][]{USGS.75.Anderson}\new{,}
+pairs \citep[e.g.,][]{USGS.75.Anderson},
 %
@@
   \int^\infty_0\,l \exp\left(-al^2\right) J_0(rl)\,dl &=
-  \frac{\exp\left(\frac{-r^2}{4a}\right)}{2a}\ , \\
+  \frac{1}{2a} \exp\left(\frac{-r^2}{4a}\right)\ , \\
   \int^\infty_0\,l^2 \exp\left(-al^2\right) J_1(rl)\,dl &=
@@
 
-\new{Figure~\ref{fig:201a} shows three different inversion runs for $N=101$,
+Figure~\ref{fig:Figure-2} shows three different inversion runs for $N=101$,
 201, and 401 for a wide range of spacing and shift values.
@@
 $\delta = 0$. If it differs from zero then there is no advantage in having an
-odd point filter.}
-%
-\plot*{201a}{width=\textwidth}{%
-  \new{Test of different filter lengths. Filter length is a trade-off between
+odd point filter.
+%
+\plot*{Figure-2}{width=\textwidth}{%
+  Test of different filter lengths. Filter length is a trade-off between
     precision and speed, as shorter filters are faster but less precise, and
   longer filters are more precise but computationally more expensive. The red
-  square in subplot (b) is the starting point of Figure~\ref{fig:201b}.} }%
-%
-
-\new{Figure~\ref{fig:201b} shows four consecutive inversion runs for $N=201$,
+  square in subplot (b) is the starting point of Figure~\ref{fig:Figure-3}.}%
+%
+
+Figure~\ref{fig:Figure-3} shows four consecutive inversion runs for $N=201$,
 where each run is a focus on a subsection of the previous run, indicated by the
-red square.} From the low resolution overview runs (a) and (b) it looks like
+red square. From the low resolution overview runs (a) and (b) it looks like
 this is a standard, straight-forward minimization problem. However, from the
@@
 %
-\plot*{201b}{width=\textwidth}{%
+\plot*{Figure-3}{width=\textwidth}{%
   Solution spaces of four consecutive inversion runs for a 201\,pt filter. Each
@@
 
-Figure~\ref{fig:201c} shows in (a) the filter values for $J_0$ and $J_1$ of the
-best 201 point filter obtained this way, and in (b) its right-hand-side
-solution. The black dots indicate negative values, from which it can seen that
-adjacent values are often alternating between positive and negative
-contributions.
-%
-\plot*{201c}{width=.9\textwidth}{(a) Filter values of the best obtained 201\,pt
-  filter with the corresponding check of quality in (b). Black points indicate
-  negative values, which shows that adjacent values have often opposite signs.}
+Figure~\ref{fig:Figure-4} shows in (a) the filter values for $J_0$ and $J_1$ of
+the best 201 \old{point}\new{pt} filter obtained this way, and in (b) its
+right-hand-side solution. The black dots indicate negative values, from which
+it can \new{be} seen that adjacent values are often alternating between
+positive and negative contributions.
+%
+\plot*{Figure-4}{width=.9\textwidth}{(a) Filter values of the best obtained
+  201\,pt filter with the corresponding check of quality in (b). Black points
+  indicate negative values, which shows that adjacent values have often
+  opposite signs.}
 %
@@
 Designing filters is to a large extent trial and error. All input variables
-influence the outcome, and often \old{a good}\new{an even better} filter is
-found by sheer luck of choosing the right starting parameters\new{, as can be
-seen in Figure~\ref{fig:201c} (h). However, once you are in this zone every
-filter is a very good filter}. Each of the input parameters has an effect on
-the outcome. It also depends heavily on the transform pairs $f_\mr{I;C}$;
-functions that decay rapidly are generally better, as noted by earlier authors
+influence the outcome, and often an even better filter is found by sheer luck
+of choosing the right starting parameters, as can be seen in
+Figure~\ref{fig:Figure-4} (h). However, once you are in this zone every filter
+is a very good filter. Each of the input parameters has an effect on the
+outcome. It also depends heavily on the transform pairs $f_\mr{I;C}$; functions
+that decay rapidly are generally better, as noted by earlier authors
 \citep[e.g.][]{USGS.75.Anderson}. It also depends on the filter length, and
@@
 In this section we compare the 201\,pt filter derived in the previous section
-to marine CSEM models used in \cite{GP.07.Kong}, \cite{GEO.12.Key}, as well as
-for a land EM case.
-
-Figure~\ref{fig:kongf5} compares the derived 201\,pt filter with the two
+to marine CSEM models \new{as} used in \cite{GP.07.Kong}, \new{in}
+\cite{GEO.12.Key}, \old{as well as}\new{and} for a land EM case.
+
+Figure~\ref{fig:Figure-5-bw} compares the derived 201\,pt filter with the two
 half-spaces case used by \cite{GP.07.Kong} in his Figure~5. The model consists
@@
 %
-\plot*{kongf5}{width=\textwidth}{(a) Results of different filters and QWE for
-  the model of Figure~5 from \cite{GP.07.Kong} with the relative errors shown
-  in (b), using the QWE result. The relative error is meaningless from about
-  15.5\,km onwards, as QWE failed as well for these very low amplitudes.}
+\plot*{Figure-5-bw}{width=\textwidth}{(a) Results of different filters and QWE
+  for the model of Figure~5 from \cite{GP.07.Kong} with the relative errors
+  shown in (b), using the QWE result. The relative error is meaningless from
+  about 15.5\,km onwards, as QWE failed as well for these very low amplitudes.}
 %
@@
 
-Figure~\ref{fig:keyf5}(a) considers the canonical CSEM model as used in
+Figure~\ref{fig:Figure-6-bw}(a) considers the canonical CSEM model as used in
 \cite{GEO.12.Key} in his Figure~5: a $0.303\,\Omega\,$m sea with depth of
@@
 %
-\plot*{keyf5}{width=\textwidth}{(a) The canonical marine CSEM model of
+\plot*{Figure-6-bw}{width=\textwidth}{(a) The canonical marine CSEM model of
   \cite{GEO.12.Key} and (b) a land model with source and receiver at the
-  surface. The new filter has generally relative errors which are several
-  orders of magnitude lower than the other filters.}
-%
-
-Figure~\ref{fig:keyf5}(b) is a land case, with a background resistivity of
-$10\,\Omega\,$m, in which a $500\,\Omega\,$m target of 100\,m thickness is
+  surface. \old{The new filter has generally relative errors which are several
+  orders of magnitude lower than the other filters.} \new{The new filter shows
+  relative errors that are generally several orders of magnitude lower than
+  for the other filters.}
+}
+%
+
+Figure~\ref{fig:Figure-6-bw}(b) is a land case, with a background resistivity
+of $10\,\Omega\,$m, in which a $500\,\Omega\,$m target of 100\,m thickness is
 embedded at 1000\,m below the surface. The source depth is 0.5\,m and the
@@
 errors of DLF using the filter \emph{Wer201} compared to QWE for a deep water
-model, a shallow water model, and a land model. The models consist of air
-(upper halfspace), a water layer in the first two examples, and a subsurface
-halfspace. Source depth is 1990\,m, 10\,m, and 0.5\,m in the deep, shallow, and
-land case, respectively. Receiver depth is 2000\,m, 100\,m, and 0.8\,m. The
-subsurface halfspace resistivity varies from 1 to 10, 100, and
-1000\,$\Omega\,$m from the left to the right column. The error plots cover
-offsets from 50\,m to 20\,km on the x-axis, and 0.01\,Hz to 10\,Hz on the
-y-axis. It can be seen in the result that the filter is sufficiently precise
-for all practical CSEM cases, hence has generally an error far below 1\,\%
-($10^{-2}$) in all cases. \new{A relative noise of less than 1\,\% is generally
-considered really good for real data.} The exception\new{s} are \new{in the
-shallow water case at} large offsets with relatively high frequencies, e.g.,
-10\,km offset with 8\,Hz\old{. In the shallow water case}\new{;} this zone has
-a very high error for subsurface resistivities of 1\,$\Omega\,$m and
-10\,$\Omega\,$m (yellow zones). However, this is due to the fact that the
-amplitudes of the signal for these offsets and frequencies are in the order of
-$10^{-25}\,$V/m. \new{The grey contour lines in the plot indicate the power of
-  the amplitude of the signal.}
+model \new{(2\,km water depth)}, a shallow water model \new{(400\,m water
+depth)}, and a land model. The models consist of air (upper halfspace), a water
+layer in the first two examples, and a subsurface halfspace. Source depth is
+1990\,m, 10\,m, and 0.5\,m in the deep, shallow, and land case, respectively.
+Receiver depth is 2000\,m, 100\,m, and 0.8\,m. The subsurface halfspace
+resistivity varies from 1 to 10, 100, and 1000\,$\Omega\,$m from the left to
+the right column. The error plots cover offsets from 50\,m to 20\,km on the
+x-axis, and 0.01\,Hz to 10\,Hz on the y-axis. It can be seen in the result that
+the filter is sufficiently precise for all practical CSEM cases, hence has
+generally an error far below 1\,\% ($10^{-2}$) in all cases. A relative noise
+of less than 1\,\% is generally considered really good for real data. The
+exceptions are in the shallow water case at large offsets with relatively high
+frequencies, e.g., 10\,km offset with 8\,Hz; this zone has a very high error
+for subsurface resistivities of 1\,$\Omega\,$m and 10\,$\Omega\,$m (yellow
+zones). However, this is due to the fact that the amplitudes of the signal for
+these offsets and frequencies are in the order of $10^{-25}\,$V/m. The grey
+contour lines in the plot indicate the power of the amplitude of the signal.
 %
@@
   frequencies (yellow zones). However, the amplitudes in these regions are in
-  the order of $10^{-25}\,$V/m, and therefore \old{far}\new{several orders of
-  magnitude} below the noise level\new{s of current EM instrument systems}.
-  \new{Contour lines show the power of the amplitude of the signal.}}
+  the order of $10^{-25}\,$V/m, and therefore several orders of
+  magnitude below the noise levels of current EM instrument systems.
+  Contour lines show the power of the amplitude of the signal.}
 %
@@
 \plot*{GPR}{width=\textwidth}{GPR example for the model given in (a) using
-  \texttt{empymod} with (b) the adaptive quadrature (QUAD) or with (c) a
-  2001\,pt digital linear filter (b). QUAD took about 11\,h 27\,min to
+  \texttt{empymod} with (b) \old{the} adaptive quadrature (QUAD) or with (c) a
+  2001\,pt digital linear filter\old{ (b)}. QUAD took about 11\,h 27\,min to
   calculate, whereas DLF took 2\,min 10\,s in parallel on 4 nodes, hence under
@@
 
-Figure~\ref{fig:ftGPR} shows in (a)-(c) the real part of the frequency domain
-results and in (d)-(f) the time domain results for offsets of 0.2\,m, 2.0\,m,
-and 3.0\,m.
-%
-\plot*{ftGPR}{width=\textwidth}{The real-part frequency-domain and the
-  time-domain responses for offsets of (a/d) 0.2\,m, (b/e) 2.0\,m, and (c/f)
-  3.0\,m.}
+Figure~\ref{fig:ftGPR} shows in (a)-(c) the real \new{and imaginary}
+part\new{s} of the frequency domain results and in (d)-(f) the time domain
+results for offsets of 0.2\,m, 2.0\,m, and 3.0\,m.
+%
+\plot*{ftGPR}{width=\textwidth}{The real\old{-part} \new{and imaginary parts of
+the} frequency-domain \new{(a-c)} and the time-domain \new{(d-f)} responses for
+offsets of (a/d) 0.2\,m, (b/e) 2.0\,m, and (c/f) 3.0\,m.}
 %
@@
 1D subsurface models together with the EM-modeler \texttt{empymod}. The code is
-available from GitHub \old{in the add-on repository}\new{as part} of
-\texttt{empymod}\old{, called \texttt{empyscripts},}\new{; the version
-  corresponding to this publication is \texttt{v1.7.1.}}\old{ from version
-  0.2.0 onwards}.
+available from GitHub as part of \texttt{empymod}; the version corresponding to
+this publication is \texttt{\new{v1.8.1}\old{v1.7.1.}}.
 
@@
 Bichara, M., and J. Lakshmanan,  1976, Fast automatic processing of resistivity
-  soundings*: Geophysical Prospecting, {\bf 24}, 354--370.
+soundings\old{*}: Geophysical Prospecting, {\bf 24}, 354--370.
 \newblock (\href{https://doi.org/10.1111/j.1365-2478.1976.tb00932.x}{doi:
@@
 --------, 1981b, The versatility of digital linear filters used in computing
-  resistivity and em sounding curves: Geoexploration, {\bf 18}, 297--310.
-\newblock (\href{https://doi.org/10.1016/0016-7142(81)90059-4}{doi:
+resistivity and \old{em}\new{EM} sounding curves: Geoexploration, {\bf 18},
+297--310. \newblock (\href{https://doi.org/10.1016/0016-7142(81)90059-4}{doi:
 10.1016/0016-7142(81)90059-4}).

### 2018-11-21 ManuscriptGeophysics2018.tex
Final figures

@@
 \subsection{Design}
-Figures~\ref{fig:Figure-2} and~\ref{fig:Figure-3} show the solution spaces of
-seven consecutive inversion runs using the following theoretical transform
+Figures~\ref{fig:Figure-2-bw} and~\ref{fig:Figure-3-bw} show the solution spaces
+of seven consecutive inversion runs using the following theoretical transform
 pairs \citep[e.g.,][]{USGS.75.Anderson},
@@
 
-Figure~\ref{fig:Figure-2} shows three different inversion runs for $N=101$,
-201, and 401 for a wide range of spacing and shift values.
-From this one has to decide for a filter length. Longer filters tend to give
+Figure~\ref{fig:Figure-2-bw} shows three different inversion runs for $N=101$,
+201, and 401 for a wide range of spacing and shift values. From this one has to
+decide for a filter length. Longer filters tend to give
 more precise results, but their transform is computationally more expensive as
@@
 %
-\plot*{Figure-2}{width=\textwidth}{%
+\plot*{Figure-2-bw}{width=\textwidth}{%
   Test of different filter lengths. Filter length is a trade-off between
@@
   longer filters are more precise but computationally more expensive. The red
-  square in subplot (b) is the starting point of Figure~\ref{fig:Figure-3}.}%
-%
-
-Figure~\ref{fig:Figure-3} shows four consecutive inversion runs for $N=201$,
+  square in subplot (b) is the starting point of
+  Figure~\ref{fig:Figure-3-bw}.}%
+%
+
+Figure~\ref{fig:Figure-3-bw} shows four consecutive inversion runs for $N=201$,
 where each run is a focus on a subsection of the previous run, indicated by the
-red square. From the low resolution overview runs (a) and (b) it looks like
-this is a standard, straight-forward minimization problem. However, from the
-more detailed results in (c) and (d) it becomes obvious that it is a
+\old{red} square. From the low resolution overview runs (a) and (b) it looks
+like this is a standard, straight-forward minimization problem. However, from
+the more detailed results in (c) and (d) it becomes obvious that it is a
 minimization problem that has to be solved stochastically, as there are
@@
 %
-\plot*{Figure-3}{width=\textwidth}{%
+\plot*{Figure-3-bw}{width=\textwidth}{%
   Solution spaces of four consecutive inversion runs for a 201\,pt filter. Each
   consecutive run zooms into a portion of the previous solution space,
-  indicated by the red square. The more in detail we obtain the solution space,
-  the more random appears to be the distribution.}%
-%
-
-Figure~\ref{fig:Figure-4} shows in (a) the filter values for $J_0$ and $J_1$ of
-the best 201 \old{point}\new{pt} filter obtained this way, and in (b) its
+  indicated by the \old{red} square. The more in detail we obtain the solution
+  space, the more random appears to be the distribution.}%
+%
+
+Figure~\ref{fig:Figure-4-bw} shows in (a) the filter values for $J_0$ and $J_1$
+of the best 201 \old{point}\new{pt} filter obtained this way, and in (b) its
 right-hand-side solution. The black dots indicate negative values, from which
@@
 %
-\plot*{Figure-4}{width=.9\textwidth}{(a) Filter values of the best obtained
+\plot*{Figure-4-bw}{width=.9\textwidth}{(a) Filter values of the best obtained
   201\,pt filter with the corresponding check of quality in (b). Black points
@@
 of choosing the right starting parameters, as can be seen in
-Figure~\ref{fig:Figure-4} (h). However, once you are in this zone every filter
-is a very good filter. Each of the input parameters has an effect on the
+Figure~\ref{fig:Figure-4-bw} (h). However, once you are in this zone every
+filter is a very good filter. Each of the input parameters has an effect on the
 outcome. It also depends heavily on the transform pairs $f_\mr{I;C}$; functions
@@
 cases. To verify the applicability of the filter we run two different tests
-over a wide range of scenarios. Figure~\ref{fig:error} shows the relative
+over a wide range of scenarios. Figure~\ref{fig:Figure-7} shows the relative
 errors of DLF using the filter \emph{Wer201} compared to QWE for a deep water
@@
 %
-\plot*{error}{width=\textwidth}{Relative error plots comparing the DLF
+\plot*{Figure-7}{width=\textwidth}{Relative error plots comparing the DLF
   (\emph{Wer201}) with QWE for the three cases deep water (top row), shallow
@@
 
-Another error test is shown in Figure~\ref{fig:skindepth1}. It shows in (a) the
+Another error test is shown in Figure~\ref{fig:Figure-8}. It shows in (a) the
 amplitude of the analytical fullspace solution for a wide range of
@@
 %
-\plot*{skindepth1}{width=.8\textwidth}{(a) Amplitude of the analytical
+\plot*{Figure-8}{width=.8\textwidth}{(a) Amplitude of the analytical
   fullspace solution for a wide range of resistivities and frequencies. The
@@
 
-Figure~\ref{fig:GPR} shows the GPR example as calculated in
+Figure~\ref{fig:Figure-9} shows the GPR example as calculated in
 \cite{GEO.15.Hunziker} and \cite{GEO.17.Werthmuller}; the model parameters are
@@
 %
-\plot*{GPR}{width=\textwidth}{GPR example for the model given in (a) using
+\plot*{Figure-9}{width=\textwidth}{GPR example for the model given in (a) using
   \texttt{empymod} with (b) \old{the} adaptive quadrature (QUAD) or with (c) a
@@
 
-Figure~\ref{fig:ftGPR} shows in (a)-(c) the real \new{and imaginary}
+Figure~\ref{fig:Figure-10-bw} shows in (a)-(c) the real \new{and imaginary}
 part\new{s} of the frequency domain results and in (d)-(f) the time domain
@@
 %
-\plot*{ftGPR}{width=\textwidth}{The real\old{-part} \new{and imaginary parts of
-the} frequency-domain \new{(a-c)} and the time-domain \new{(d-f)} responses for
-offsets of (a/d) 0.2\,m, (b/e) 2.0\,m, and (c/f) 3.0\,m.}
+\plot*{Figure-10-bw}{width=\textwidth}{The real\old{-part} \new{and imaginary
+  parts of the} frequency-domain \new{(a-c)} and the time-domain \new{(d-f)}
+  responses for \new{horizontal} offsets of (a/d) 0.2\,m, (b/e) 2.0\,m, and
+  (c/f) 3.0\,m.}
 %
@@
 parameters we run a test where we swapped layers one and two of the model in
-Figure~\ref{fig:GPR} (a). The resulting real and imaginary parts for the
-offsets 0.2\,m, 2.0\,m, and 3.0\,m are given in Figure~\ref{fig:ftGPR2nd}.
-%
-\plot*{ftGPR2nd}{width=\textwidth}{The real and imaginary parts of the
-  frequency-domain response for the model given in Figure~\ref{fig:GPR} (a)
-  with swapped first and second layers for offsets of (a) 0.2\,m, (b) 2.0\,m,
-  and (c) 3.0\,m.}
+Figure~\ref{fig:Figure-9} (a). The resulting real and imaginary parts for the
+offsets 0.2\,m, 2.0\,m, and 3.0\,m are given in Figure~\ref{fig:Figure-11-bw}.
+%
+\plot*{Figure-11-bw}{width=\textwidth}{The real and imaginary parts of the
+  frequency-domain response for the model given in Figure~\ref{fig:Figure-9}
+  (a) with swapped first and second layers for \new{horizontal} offsets of (a)
+  0.2\,m, (b) 2.0\,m, and (c) 3.0\,m.}
 %

### 2018-11-22 ManuscriptGeophysics2018.tex
Final Submission

@@
 % \documentclass[manuscript]{geophysics}
-% \documentclass[paper,twocolumn,twoside]{geophysics}
+\documentclass[paper,twocolumn,twoside]{geophysics}
 % \documentclass[paper]{geophysics}
 % \documentclass[paper, revised]{geophysics}
-\documentclass[manuscript,revised]{geophysics}
+% \documentclass[manuscript,revised]{geophysics}
 
@@
 \usepackage[strings]{underscore}
-% \usepackage[pdftex, final]{hyperref}
-% \hypersetup{allcolors=blue, allbordercolors={0 0 .5}, colorlinks=true}
-\usepackage[pdftex, hidelinks]{hyperref}
+\usepackage[pdftex, final]{hyperref}
+\hypersetup{allcolors=blue, allbordercolors={0 0 .5}, colorlinks=true}
+% \usepackage[pdftex, hidelinks]{hyperref}
 
@@
 \href{mailto:E.C.Slob@tudelft.nl}{E.C.Slob@tudelft.nl};
-\footnotemark[2]Instituto Mexicano del Petróleo,
+\footnotemark[2]\emph{formerly at the} Instituto Mexicano del Petróleo,
 Eje Central Lázaro Cárdenas Norte 152,
@@
 Fourier sine/cosine transform), as all higher Bessel functions can be
-rewritten, via recurrence relations, using only\old{use} these two.
-\cite{GP.94.Mohsen} is one of the rare cases which provides $J_2$ filter
-weights.
+rewritten, via recurrence relations, using only these two. \cite{GP.94.Mohsen}
+is one of the rare cases which provides $J_2$ filter weights.
 
@@
 Nevertheless, \cite{GEO.17.Werthmuller} tested DLF for modeling 250 MHz
-center-frequency GPR data using a 401 \old{point}\new{pt} filter that was
-designed for diffusive EM, and found that it was orders of magnitude faster
-than quadrature approaches. However, good quality results were only obtained
-for the first arrivals at short time-offsets and the later arrivals had poor
-quality. Yet, this promising initial test suggests it might be possible to
-specifically design a special filter for GPR frequencies where the EM fields
-propagate as waves rather than by diffusion.
+center-frequency GPR data using a 401 pt filter that was designed for diffusive
+EM, and found that it was orders of magnitude faster than quadrature
+approaches. However, good quality results were only obtained for the first
+arrivals at short time-offsets and the later arrivals had poor quality. Yet,
+this promising initial test suggests it might be possible to specifically
+design a special filter for GPR frequencies where the EM fields propagate as
+waves rather than by diffusion.
 
@@
 %
-\plot*{Figure-1-bw}{width=.9\textwidth}{(a) A regular \new{right-hand-side
-  (}RHS\new{)} curve of a transform pair with a purely decaying function.
-  Minimum amplitude or maximum $r$ yield the same result in this scenario.
-  Using a relative error criteria is more stable than just the absolute minimum
-  amplitude. (b) A 1D model for $f=200\,$MHz. Here, the maximum $r$ provides a
-  better criteria for the inversion.}
+\plot*{Figure-1-bw}{width=.9\textwidth}{(a) A regular right-hand-side (RHS)
+  curve of a transform pair with a purely decaying function. Minimum amplitude
+  or maximum $r$ yield the same result in this scenario. Using a relative error
+  criteria is more stable than just the absolute minimum amplitude. (b) A 1D
+  model for $f=200\,$MHz. Here, the maximum $r$ provides a better criteria for
+  the inversion.}
 %
@@
 where each run is a focus on a subsection of the previous run, indicated by the
-\old{red} square. From the low resolution overview runs (a) and (b) it looks
-like this is a standard, straight-forward minimization problem. However, from
-the more detailed results in (c) and (d) it becomes obvious that it is a
-minimization problem that has to be solved stochastically, as there are
-solutions with 2 orders of magnitude difference apparently randomly next to
-each other.
+square. From the low resolution overview runs (a) and (b) it looks like this is
+a standard, straight-forward minimization problem. However, from the more
+detailed results in (c) and (d) it becomes obvious that it is a minimization
+problem that has to be solved stochastically, as there are solutions with 2
+orders of magnitude difference apparently randomly next to each other.
 %
@@
   consecutive run zooms into a portion of the previous solution space,
-  indicated by the \old{red} square. The more in detail we obtain the solution
-  space, the more random appears to be the distribution.}%
+  indicated by the square. The more in detail we obtain the solution space, the
+  more random appears to be the distribution.}%
 %
@@
 Figure~\ref{fig:Figure-4-bw} shows in (a) the filter values for $J_0$ and $J_1$
-of the best 201 \old{point}\new{pt} filter obtained this way, and in (b) its
-right-hand-side solution. The black dots indicate negative values, from which
-it can \new{be} seen that adjacent values are often alternating between
-positive and negative contributions.
+of the best 201 pt filter obtained this way, and in (b) its right-hand-side
+solution. The black dots indicate negative values, from which it can be seen
+that adjacent values are often alternating between positive and negative
+contributions.
 %
@@
 In this section we compare the 201\,pt filter derived in the previous section
-to marine CSEM models \new{as} used in \cite{GP.07.Kong}, \new{in}
-\cite{GEO.12.Key}, \old{as well as}\new{and} for a land EM case.
+to marine CSEM models as used in \cite{GP.07.Kong}, in \cite{GEO.12.Key},
+and for a land EM case.
 
@@
   \cite{GEO.12.Key} and (b) a land model with source and receiver at the
-  surface. \old{The new filter has generally relative errors which are several
-  orders of magnitude lower than the other filters.} \new{The new filter shows
-  relative errors that are generally several orders of magnitude lower than
-  for the other filters.}
+  surface. The new filter shows relative errors that are generally several
+  orders of magnitude lower than for the other filters.
 }
@@
 errors of DLF using the filter \emph{Wer201} compared to QWE for a deep water
-model \new{(2\,km water depth)}, a shallow water model \new{(400\,m water
-depth)}, and a land model. The models consist of air (upper halfspace), a water
-layer in the first two examples, and a subsurface halfspace. Source depth is
-1990\,m, 10\,m, and 0.5\,m in the deep, shallow, and land case, respectively.
-Receiver depth is 2000\,m, 100\,m, and 0.8\,m. The subsurface halfspace
-resistivity varies from 1 to 10, 100, and 1000\,$\Omega\,$m from the left to
-the right column. The error plots cover offsets from 50\,m to 20\,km on the
-x-axis, and 0.01\,Hz to 10\,Hz on the y-axis. It can be seen in the result that
-the filter is sufficiently precise for all practical CSEM cases, hence has
-generally an error far below 1\,\% ($10^{-2}$) in all cases. A relative noise
-of less than 1\,\% is generally considered really good for real data. The
-exceptions are in the shallow water case at large offsets with relatively high
-frequencies, e.g., 10\,km offset with 8\,Hz; this zone has a very high error
-for subsurface resistivities of 1\,$\Omega\,$m and 10\,$\Omega\,$m (yellow
-zones). However, this is due to the fact that the amplitudes of the signal for
-these offsets and frequencies are in the order of $10^{-25}\,$V/m. The grey
-contour lines in the plot indicate the power of the amplitude of the signal.
+model (2\,km water depth), a shallow water model (400\,m water depth), and a
+land model. The models consist of air (upper halfspace), a water layer in the
+first two examples, and a subsurface halfspace. Source depth is 1990\,m, 10\,m,
+and 0.5\,m in the deep, shallow, and land case, respectively. Receiver depth is
+2000\,m, 100\,m, and 0.8\,m. The subsurface halfspace resistivity varies from 1
+to 10, 100, and 1000\,$\Omega\,$m from the left to the right column. The error
+plots cover offsets from 50\,m to 20\,km on the x-axis, and 0.01\,Hz to 10\,Hz
+on the y-axis. It can be seen in the result that the filter is sufficiently
+precise for all practical CSEM cases, hence has generally an error far below
+1\,\% ($10^{-2}$) in all cases. A relative noise of less than 1\,\% is
+generally considered really good for real data. The exceptions are in the
+shallow water case at large offsets with relatively high frequencies, e.g.,
+10\,km offset with 8\,Hz; this zone has a very high error for subsurface
+resistivities of 1\,$\Omega\,$m and 10\,$\Omega\,$m (yellow zones). However,
+this is due to the fact that the amplitudes of the signal for these offsets and
+frequencies are in the order of $10^{-25}\,$V/m. The grey contour lines in the
+plot indicate the power of the amplitude of the signal.
 %
@@
 \plot*{Figure-9}{width=\textwidth}{GPR example for the model given in (a) using
-  \texttt{empymod} with (b) \old{the} adaptive quadrature (QUAD) or with (c) a
-  2001\,pt digital linear filter\old{ (b)}. QUAD took about 11\,h 27\,min to
-  calculate, whereas DLF took 2\,min 10\,s in parallel on 4 nodes, hence under
-  9\,min in total.}
-%
-
-Figure~\ref{fig:Figure-10-bw} shows in (a)-(c) the real \new{and imaginary}
-part\new{s} of the frequency domain results and in (d)-(f) the time domain
-results for offsets of 0.2\,m, 2.0\,m, and 3.0\,m.
-%
-\plot*{Figure-10-bw}{width=\textwidth}{The real\old{-part} \new{and imaginary
-  parts of the} frequency-domain \new{(a-c)} and the time-domain \new{(d-f)}
-  responses for \new{horizontal} offsets of (a/d) 0.2\,m, (b/e) 2.0\,m, and
-  (c/f) 3.0\,m.}
+  \texttt{empymod} with (b) adaptive quadrature (QUAD) or with (c) a
+  2001\,pt digital linear filter. QUAD took about 11\,h 27\,min to calculate,
+  whereas DLF took 2\,min 10\,s in parallel on 4 nodes, hence under 9\,min in
+  total.}
+%
+
+Figure~\ref{fig:Figure-10-bw} shows in (a)-(c) the real and imaginary
+parts of the frequency domain results and in (d)-(f) the time domain results
+for offsets of 0.2\,m, 2.0\,m, and 3.0\,m.
+%
+\plot*{Figure-10-bw}{width=\textwidth}{The real and imaginary parts of the
+  frequency-domain (a-c) and the time-domain (d-f) responses for
+  horizontal offsets of (a/d) 0.2\,m, (b/e) 2.0\,m, and (c/f) 3.0\,m.}
 %
@@
   frequency-domain response for the model given in Figure~\ref{fig:Figure-9}
-  (a) with swapped first and second layers for \new{horizontal} offsets of (a)
+  (a) with swapped first and second layers for horizontal offsets of (a)
   0.2\,m, (b) 2.0\,m, and (c) 3.0\,m.}
@@
 available from GitHub as part of \texttt{empymod}; the version corresponding to
-this publication is \texttt{\new{v1.8.1}\old{v1.7.1.}}.
+this publication is \texttt{v1.8.1}.
 
@@
 
-D.W. would like to thank the electromagnetic research group of Aleksandr
-Mousatov at the IMP for fruitful discussions, and Jürg Hunziker for valuable
-feedback which improved this manuscript considerably. We thank the reviewers
+D.W. would like to thank the electromagnetic research group of A. Mousatov at
+the IMP for fruitful discussions, and J. Hunziker for valuable feedback which
+improved this manuscript considerably. We thank the assistant editor J. Etgen,
+the associate editor J. Dellinger, and the reviewers C. Weiss and R. Mittet
 for valuable inputs not only regarding the manuscript, but also regarding the
@@
 Bichara, M., and J. Lakshmanan,  1976, Fast automatic processing of resistivity
-soundings\old{*}: Geophysical Prospecting, {\bf 24}, 354--370.
+  soundings Geophysical Prospecting, {\bf 24}, 354--370.
 \newblock (\href{https://doi.org/10.1111/j.1365-2478.1976.tb00932.x}{doi:
@@
 --------, 1981b, The versatility of digital linear filters used in computing
-resistivity and \old{em}\new{EM} sounding curves: Geoexploration, {\bf 18},
-297--310. \newblock (\href{https://doi.org/10.1016/0016-7142(81)90059-4}{doi:
+resistivity and EM sounding curves: Geoexploration, {\bf 18}, 297--310.
+\newblock (\href{https://doi.org/10.1016/0016-7142(81)90059-4}{doi:
 10.1016/0016-7142(81)90059-4}).

### 2018-12-04 DLF-in-Geophysics.tex
Correct DOI; add summary

--- 2018-11-22 ManuscriptGeophysics2018.tex
@@
-% \documentclass[manuscript]{geophysics}
-\documentclass[paper,twocolumn,twoside]{geophysics}
-% \documentclass[paper]{geophysics}
-% \documentclass[paper, revised]{geophysics}
-% \documentclass[manuscript,revised]{geophysics}
+\documentclass[a4paper, twoside, parskip, 10pt]{scrartcl}
 
-% Additional packages to geophysics.cls
+% Packages
 \usepackage[USenglish]{babel}
@@
 \usepackage[T1]{fontenc}
+
+\usepackage[hmargin=1.5cm, top=1.5cm, bottom=0cm, includeheadfoot]{geometry}
+\usepackage{multicol}
+\usepackage{fancyhdr}
+\usepackage{xspace}
 \usepackage{amssymb, amsmath, amsfonts}
-\usepackage{upquote}
-\usepackage[strings]{underscore}
+\usepackage{natbib}
+
 \usepackage[pdftex, final]{hyperref}
+
+% Settings
 \hypersetup{allcolors=blue, allbordercolors={0 0 .5}, colorlinks=true}
-% \usepackage[pdftex, hidelinks]{hyperref}
 
-% Figures
-\DeclareGraphicsExtensions{.pdf,.png,.jpg}
-\renewcommand{\figdir}{./figures}
+\frenchspacing
 
-% Own commands
+\setkomafont{subsection}{\normalfont\itshape}
+
+\pagestyle{fancy}
+\fancyhf{}
+\renewcommand{\headrulewidth}{0.0pt}
+\renewcommand{\footrulewidth}{0.0pt}
+\addtolength{\headheight}{0.0pt}
+\fancyhead[LO, RE]{Digital Linear Filters in Geophysics}
+\fancyhead[RO, LE]{\thepage}
+\fancypagestyle{empty}{
+  \fancyhead[LE, RE, LO]{}
+  \fancyhead[RO]{\today}}
+
+\setcounter{secnumdepth}{-1}
+
+\setlength\columnsep{15pt}
+% \columnseprule1pt
+
+% Commands
 \newcommand{\mr}[1]{\mathrm{#1}}
@@
 \begin{document}
+\thispagestyle{empty}
 
-\title{A tool for designing digital filters for the Hankel and Fourier
-transforms in potential, diffusive, and wavefield modeling}
+{\huge \bf \sffamily Digital Linear Filters in Geophysics\\[.5cm]}
+{\small\emph{Excerpt from \cite{GEO.18.Werthmuller}; for the full article see
+\href{https://github.com/empymod/article-fdesign}{github.com/empymod/article-fdesign}.}}
 
-\renewcommand{\thefootnote}{\fnsymbol{footnote}}
+\vspace{.5cm}
 
-\ms{GEO-2018-0069.R2}
+\begin{multicols}{2}
 
-\address{
-\footnotemark[1]TU Delft,
-Building 23,
-Stevinweg 1 / PO-box 5048,
-2628 CN Delft,
-E-mail: \href{mailto:D.Werthmuller@tudelft.nl}{D.Werthmuller@tudelft.nl};
-\href{mailto:E.C.Slob@tudelft.nl}{E.C.Slob@tudelft.nl};
-\footnotemark[2]\emph{formerly at the} Instituto Mexicano del Petróleo,
-Eje Central Lázaro Cárdenas Norte 152,
-Col. San Bartolo Atepehuacan C.P. 07730,
-Ciudad de México, México,
-\footnotemark[3]Lamont-Doherty Earth Observatory, Columbia University,
-305C Oceanography,
-61 Route 9W, PO Box 1000,
-Palisades NY 10964-8000 US,
-E-mail: \href{mailto:KKey@ldeo.columbia.edu}{KKey@ldeo.columbia.edu}.
-}
-
-\author{%
-Dieter Werthmüller\footnotemark[1]\footnotemark[2], %
-Kerry Key\footnotemark[3], and %
-Evert C.\ Slob\footnotemark[1]%
-}
-
-\footer{}
-\lefthead{Werthmüller et al.}
-\righthead{Digital filter designing tool}
-
-\maketitle
-
-%%fakesection ===    ABSTRACT    ===
-\begin{abstract} % 1-2 sentence(s) each
-%
-% (1) PRINCIPAL OBJECTIVES AND SCOPE OF THE WORK
-  The open-source code \texttt{fdesign} makes it possible to design digital
-  linear filters for the Hankel and Fourier transforms used in potential-,
-  diffusive-, and wavefield modeling. Digital filters can be derived for any
-  electromagnetic method, such as methods in the diffusive limits (direct
-  current, controlled-source electromagnetics) as well as methods using higher
-  frequency content (ground-penetrating radar, acoustic and elastic wavefields).
-%
-% (2) METHODOLOGY
-  The direct matrix inversion method is used for the derivation of the filter
-  values, and a brute-force inversion is carried out over the defined spacing
-  and shifting values of the filter basis. Included or user-provided
-  theoretical transform pairs are used for the inversion. Alternatively, one
-  can provide layered subsurface models that will be computed with a precise
-  quadrature method using the electromagnetic modeler \texttt{empymod} to
-  generate numerical transform pairs.
-%
-% (3) RESULTS
-  The comparison of the presented 201\,pt filter with previously presented
-  filters shows that it performs better for some standard controlled-source
-  electromagnetic cases. The derivation of a longer 2001\,pt filter for a
-  ground-penetrating radar example with 250\,MHz center frequency proves that
-  the filter method works also for wave phenomena, not only for diffusive
-  electromagnetic fields.
-%
-% (4) CONCLUSIONS
-  The presented algorithm provides a tool to create problem specific digital
-  filters. Such purpose-built filters can be made shorter and can speed up
-  consecutive  potential-, diffusive-, and wavefield inversions.
-%
-\end{abstract}
-
-\section{Introduction}
+\section{Review}
 
@@
 
-\subsubsection{(1) New applications}
+\subsection{(1) New applications}
 
@@
 
-\subsubsection{(2) Filter improvements}
+\subsection{(2) Filter improvements}
 
@@
 
-\subsubsection{(3) Computational tools}
+\subsection{(3) Computational tools}
 
@@
 
-While there are several versions of digital filters that are freely available,
-there are no open-source filter designing tools that readily available; they
-are either only described methods in articles or the codes are only available
-upon request by the author. Here, we address this gap by presenting the code
-\texttt{fdesign}, which is an open-source tool for designing general or
-purpose-built filters using the direct matrix inversion method. The code is
-completely open-source and can be run using freely available programming
-environments, thus there are few barriers to obtaining and running the code.
-
-In the following sections we give a brief overview of the methodology, the
-theory, and the code, and then show examples of its usage, as well as
-presenting new filters for both CSEM and GPR data. The current article falls
-therefore into all three mentioned categories by the new application of DLF to
-wave phenomena, which means that the DLF can probably be used to model
-seismic wavefields as well, and by providing new filters with the accompanying
-code. The algorithm and many more examples of its usage can be found on
-\href{https://github.com/empymod/article-fdesign}{github.com/empymod/article-fdesign}.
-The examples can be used as templates to design new filters.
-
-\section{Methodology}
-
-The algorithm \texttt{fdesign} is a filter designing tool using the direct
-matrix inversion method as described by \cite{GP.07.Kong} and based on scripts
-developed by \cite{GEO.12.Key}. The tool is an add-on to the EM modeler
-\texttt{empymod} \citep{GEO.17.Werthmuller}, written in Python, and hosted on
-GitHub, which should foster interaction and enable anyone to toy around,
-improve, and extend it. It can be used to derive digital linear filters for the
-Hankel and the Fourier transforms for potential, diffusive, and wavefield
-modeling (hence from DC to GPR).
-
-Theoretically, it can be used to derive linear filters for any linear
-transform, as long as it is supplied with the inputs and outputs of a transform
-pair. Previous works have relied on using theoretical transform pairs commonly
-found in tables of integral transforms. Here we follow a similar approach, but
-we also propose a new method where filters could be derived from accurately
-computed numerical transform pairs, such as those computed using quadrature
-approaches. For example, the EM modeler \texttt{empymod} could be used to
-compute transform pairs for a particular class of model and data parameters
-(for example, a specific EM method and model of interest) using a slow but
-highly accurate quadrature method. A good filter derived from the resulting
-transform pairs could then be used for faster modeling of similar model and
-data parameters.
-
-The main features of \texttt{fdesign} and some differences between it and the
-methods it is based upon are:
-\begin{itemize}
-  \item The algorithm computes many different filters for various spacing and
-    shift values (brute-force), with an optional minimization routine for
-    the best outcome of the brute-force step.
-  \item \cite{GP.07.Kong} and \cite{GEO.12.Key} optimize a filter for the $J_0$
-    filter, and then use the obtained shift and spacing value to calculate the
-    $J_1$ values. The presented algorithm can optimize either $J_0$, $J_1$, or
-    both in the optimization process. Five transform-pairs for each $J_0$ and
-    $J_1$, and three transform pairs for each sine and cosine of often used
-    functions are included in the routine, but any other transform pair can be
-    provided as input.
-  \item More complex or more specific numerical transform  pairs  can be
-    provided instead of theoretical transform pairs by using the modeler
-    \texttt{empymod}.
-  \item \cite{GP.07.Kong} defines a \emph{good} filter as one that recovers
-    \emph{weak} diffusive EM fields. In  \texttt{fdesign}, you can define a
-    relative error level that defines up to what error an obtained result is
-    good or not. The optimization seeks to find the smallest amplitude field
-    that is accurately modeled (i.e., minimizing the amplitude), but it has
-    additionally another mode, where it will maximize the output abscissae $r$
-    of the transform pair instead of minimizing the amplitude. The two modes
-    yield the same result in the general, simple case of fast decaying
-    transform pairs. However, for some complex models and specifically for very
-    high frequencies, maximizing $r$ will yield much better and more consistent
-    results.
-  \item The algorithm can solve equal- and over-determined
-    linear systems.
-  \item The real or the imaginary part can be used for the inversion of complex
-    transform pairs.
-\end{itemize}
-
-The quality of a filter depends heavily on the model chosen for comparison. An
-obtained filter might be very good for one model or data acquisition setup, but
-not that good for another one, which is no different for this designing tool.
-There is no way to estimate the error of a result obtained with a certain
-filter if you apply it to any other model, unless you calculate this other
-model with another method for comparison. All results presented here should
-therefore be taken with a certain care. Further, when applying a new filter to
-a significantly different model and/or data space than what it has previously
-been tested on, we recommend first verifying its accuracy in that space by a
-comparison with results computed with a quadrature method such as QWE, which is
-also available in \texttt{empymod}.
-
-\subsection{Theory}
-
+\section{Theory}
 Most of the articles mentioned in the review have detailed derivations of the
@@
 
-\subsection{Pseudo-code for \texttt{fdesign} }
-
-The main input variables are the filter length ($N$), the spacing ($\Delta$)
-and shift ($\delta$) ranges over which to loop, and the transform pairs for the
-inversion ($f_\mr{I}$) and the check of quality ($f_\mr{C}$). If $f_\mr{C}$ is
-not provided, then $f_\mr{I}$ is used for both. There are additional, optional
-input parameters, for instance to adjust how the RHS abscissae for the
-inversion $r_\mr{I} = p(b, N)$ are calculated, where $b$ is the filter base.
-Additionally, ranges $r$ for evaluating the check function need to be
-specified.
-
-The basic steps are as follows:
-\begin{enumerate}
-  \item Evaluate RHS of check-function $f_\mr{C}$:\newline
-    $d_\mr{R} = f_\mr{C}^\mr{rhs}(r)$
-  \item Loop over each value $\Delta_i$, $\delta_j$ (brute force;
-      parameters on the left-hand side are overwritten where $i$, $j$
-      are not mentioned):
-    \begin{enumerate}
-      \item Calculate filter base:\newline
-        $b_n = \exp\left\{\Delta_i(-\lfloor{(N+1)/2\rfloor}+n) +
-      \delta_j\right\}$
-      \item Get required RHS and LHS evaluation points:\newline
-        $r_\mr{I} = p(b, N)$\newline
-        $l_\mr{I} = b/r_\mr{I}$
-      \item Invert for filter coefficients:\newline
-        $J = \mr{solve}\left[f_\mr{I}^\mr{lhs}(l_\mr{I}),
-        f_\mr{I}^\mr{rhs}(r_\mr{I})\right]$
-    \item Calculate numerically RHS of check-function $f_\mr{C}$ with current
-      filter $J$:\newline
-      $d_\mr{F} = f_\mr{C}^\mr{lhs}(b/r)\cdot J\ /\ r$
-      \item Store minimum recovered amplitude or maximum $r$ where relative
-        error is less than the provided error:\newline
-        $\chi_{ij} = g\left[
-        \mr{argmin}\left(|(d_\mr{F}-d_\mr{R})/d_\mr{R}| >
-        \mr{error}\right) -1\right]\ $, where $g$ is either $d_\mr{R}$
-        or $1/r$.
-    \end{enumerate}
-  \item Re-calculate (steps 2a-2c) filter base and
-    coefficients which yield minimum amplitude or maximum $r$ (a local
-    minimization can be run to polish the brute-force result):\newline
-    $i_\mr{best}$, $j_\mr{best}$ = $\mr{argmin}(\chi)$\newline
-    recalculate $b$, $J$ with $i_\mr{best}$, $j_\mr{best}$\newline
-    \newline
-  \item Return filter base and coefficients and error matrix (for visual
-    QC):\newline
-    return $b$, $J$, $\chi$\newline
-\end{enumerate}
-
-To solve step 2(c) we implemented the QR-decomposition method in
-  \texttt{fdesign}.
-
-\subsection{Minimization criteria}
-Figure~\ref{fig:Figure-1-bw} shows the differences between the different
-minimization approaches; in (a) for a conventional, fast decaying transform
-pair (using the $J_0$ kernel of a fullspace with vertical distance between
-source and receiver of 50\,m, frequency $f=1\,$Hz, and resistivity
-$\rho=1\,\Omega\,$m); and in (b) for a more complex, high frequency layered
-model (using the $J_0$ kernel for $f=200\,$MHz for the GPR model described
-in the section \emph{Ground-penetrating radar}, with receiver located at 1\,m
-depth). The circles show the minimum amplitude used in previous approaches.
-This criterion is not ideal, as it is subject to some random fluctuations and
-depends on the choice of $r$. The squares show the minimum amplitude given a
-certain acceptable error, and the diamonds show the maximum $r$ given a certain
-acceptable error (the inversion is a minimization process, it therefore
-minimizes $1/r$, not $r$). In simple cases the minimum amplitude given an
-acceptable error and the maximum $r$ given an acceptable error will yield the
-same result. However, in complex cases the maximum $r$ is more consistent and
-therefore a much better criterion.
-%
-\plot*{Figure-1-bw}{width=.9\textwidth}{(a) A regular right-hand-side (RHS)
-  curve of a transform pair with a purely decaying function. Minimum amplitude
-  or maximum $r$ yield the same result in this scenario. Using a relative error
-  criteria is more stable than just the absolute minimum amplitude. (b) A 1D
-  model for $f=200\,$MHz. Here, the maximum $r$ provides a better criteria for
-  the inversion.}
-%
-
-While the $\chi$ array is used to identify the optimal values of $\Delta$ and
-$\delta$, the value of $\mr{argmin}(\chi)$ will depend on the choice of the
-check function $f_C$ and can vary from one transform pair to another; thus it
-should not be used to compare the quality of filters created with different
-transform pairs.
-
-\section{Numerical examples}
-
-The numerical examples are focused on the Hankel transform, although
-\texttt{fdesign} can be used to design digital linear filters for both Hankel
-and Fourier transforms. As there is no difference in the procedure of the two,
-this should be sufficient to demonstrate the algorithm.
-
-\subsection{Design}
-Figures~\ref{fig:Figure-2-bw} and~\ref{fig:Figure-3-bw} show the solution spaces
-of seven consecutive inversion runs using the following theoretical transform
-pairs \citep[e.g.,][]{USGS.75.Anderson},
-%
-\begin{align}
-  \int^\infty_0\,l \exp\left(-al^2\right) J_0(rl)\,dl &=
-  \frac{1}{2a} \exp\left(\frac{-r^2}{4a}\right)\ , \\
-  \int^\infty_0\,l^2 \exp\left(-al^2\right) J_1(rl)\,dl &=
-  \frac{r}{4a^2} \exp\left(-\frac{r^2}{4a}\right)\ ,
-  \label{eq:j01}
-\end{align}
-%
-where $a$ was set to 5. In the algorithm you can provide one pair
-for the inversion, and a different pair to get the minimum amplitude or the
-maximum $r$. In this case the same transform pair was used for both the
-inversion and the check of quality. The acceptable relative error was set to
-1\,\%. The RHS evaluation parameter for the inversion was set so  that an
-over-determined system was evaluated with $M = 2N$ equations, where $r_\mr{I}$
-was logarithmically spaced from $\log_{10}[1/\max(b)] - 1$ to
-$\log_{10}[1/\min(b)] + 1$, where $b$ are the filter abscissae as given in
-equation~\ref{eq:base}. This is accomplished in \texttt{fdesign} by setting
-parameter $r_\mr{def} = (1, 1, 2)$.
-
-Figure~\ref{fig:Figure-2-bw} shows three different inversion runs for $N=101$,
-201, and 401 for a wide range of spacing and shift values. From this one has to
-decide for a filter length. Longer filters tend to give
-more precise results, but their transform is computationally more expensive as
-more values have to be evaluated. The length of the filter does not has to be
-an odd number, but many published filters do have an odd number of
-coefficients. The integral runs from $-\infty$ to $+\infty$ and hence an odd
-point filter comes naturally by truncating the integral to a finite sum from
-$-(N-1)/2$ to $(N-1)/2$. However, this is only true if the shift value
-$\delta = 0$. If it differs from zero then there is no advantage in having an
-odd point filter.
-%
-\plot*{Figure-2-bw}{width=\textwidth}{%
-  Test of different filter lengths. Filter length is a trade-off between
-    precision and speed, as shorter filters are faster but less precise, and
-  longer filters are more precise but computationally more expensive. The red
-  square in subplot (b) is the starting point of
-  Figure~\ref{fig:Figure-3-bw}.}%
-%
-
-Figure~\ref{fig:Figure-3-bw} shows four consecutive inversion runs for $N=201$,
-where each run is a focus on a subsection of the previous run, indicated by the
-square. From the low resolution overview runs (a) and (b) it looks like this is
-a standard, straight-forward minimization problem. However, from the more
-detailed results in (c) and (d) it becomes obvious that it is a minimization
-problem that has to be solved stochastically, as there are solutions with 2
-orders of magnitude difference apparently randomly next to each other.
-%
-\plot*{Figure-3-bw}{width=\textwidth}{%
-  Solution spaces of four consecutive inversion runs for a 201\,pt filter. Each
-  consecutive run zooms into a portion of the previous solution space,
-  indicated by the square. The more in detail we obtain the solution space, the
-  more random appears to be the distribution.}%
-%
-
-Figure~\ref{fig:Figure-4-bw} shows in (a) the filter values for $J_0$ and $J_1$
-of the best 201 pt filter obtained this way, and in (b) its right-hand-side
-solution. The black dots indicate negative values, from which it can be seen
-that adjacent values are often alternating between positive and negative
-contributions.
-%
-\plot*{Figure-4-bw}{width=.9\textwidth}{(a) Filter values of the best obtained
-  201\,pt filter with the corresponding check of quality in (b). Black points
-  indicate negative values, which shows that adjacent values have often
-  opposite signs.}
-%
-
-Designing filters is to a large extent trial and error. All input variables
-influence the outcome, and often an even better filter is found by sheer luck
-of choosing the right starting parameters, as can be seen in
-Figure~\ref{fig:Figure-4-bw} (h). However, once you are in this zone every
-filter is a very good filter. Each of the input parameters has an effect on the
-outcome. It also depends heavily on the transform pairs $f_\mr{I;C}$; functions
-that decay rapidly are generally better, as noted by earlier authors
-\citep[e.g.][]{USGS.75.Anderson}. It also depends on the filter length, and
-obviously heavily on the spacing and shift values. Another important point is
-how you define the right-hand evaluation points of the inversion
-($r_\mr{def}$). Evaluating corresponding transform pairs separately or jointly
-also leads to different filter coefficients ($J_0$, $J_1$, or $J_0$ \& $J_1$,
-or equally sine, cosine, or sine \& cosine); also if the real or the imaginary
-part is used when complex transform pairs or \texttt{empymod} is used; and if
-it is inverted for the minimum amplitude or for the maximum $r$.
-
-\subsection{CSEM}
-In this section we compare the 201\,pt filter derived in the previous section
-to marine CSEM models as used in \cite{GP.07.Kong}, in \cite{GEO.12.Key},
-and for a land EM case.
-
-Figure~\ref{fig:Figure-5-bw} compares the derived 201\,pt filter with the two
-half-spaces case used by \cite{GP.07.Kong} in his Figure~5. The model consists
-of a $0.3125\,\Omega\,$m sea layer of infinite thickness above a subsurface
-$1\,\Omega\,$m half-space. The signal of a $1\,$Hz frequency $x$-directed
-electric source 50\,m above the interface is measured at an $x$-directed
-electric receiver at the interface.
-%
-\plot*{Figure-5-bw}{width=\textwidth}{(a) Results of different filters and QWE
-  for the model of Figure~5 from \cite{GP.07.Kong} with the relative errors
-  shown in (b), using the QWE result. The relative error is meaningless from
-  about 15.5\,km onwards, as QWE failed as well for these very low amplitudes.}
-%
-In (a) it can be seen that the new 201\,pt filter is able to recover smaller
-amplitudes than the 241\,pt filter from \cite{GP.07.Kong}, the 201\,pt filter
-from \cite{GEO.12.Key}, and the 801\,pt filter from \cite{TMS.82.Anderson}
-(Wer201, Kong241, Key201, and And801, respectively). It behaves equally well as
-the quadrature with extrapolation (QWE), for which we used a 51\,pt quadrature
-with relative and absolute tolerance of $10^{-12}$ and $10^{-30}$,
-respectively. The relative error is shown in (b), where the QWE result was
-taken as \emph{truth}. For offsets greater than roughly 15.5\,km the relative
-error becomes meaningless, as the QWE fails itself; this part is greyed out in
-the figure.
-
-Figure~\ref{fig:Figure-6-bw}(a) considers the canonical CSEM model as used in
-\cite{GEO.12.Key} in his Figure~5: a $0.303\,\Omega\,$m sea with depth of
-2\,km, a background resistivity of $1\,\Omega\,$m, in which a $100\,\Omega\,$m
-target layer of 100\,m thickness is embedded at 1000\,m below the seafloor.
-The source is located at 1990\,m and receivers are positioned on the seafloor.
-%
-\plot*{Figure-6-bw}{width=\textwidth}{(a) The canonical marine CSEM model of
-  \cite{GEO.12.Key} and (b) a land model with source and receiver at the
-  surface. The new filter shows relative errors that are generally several
-  orders of magnitude lower than for the other filters.
-}
-%
-
-Figure~\ref{fig:Figure-6-bw}(b) is a land case, with a background resistivity
-of $10\,\Omega\,$m, in which a $500\,\Omega\,$m target of 100\,m thickness is
-embedded at 1000\,m below the surface. The source depth is 0.5\,m and the
-receiver depth is 0.8\,m. In both cases the new filter \emph{Wer201} has
-generally a relative error which is orders of magnitude lower than the other
-filters. The relative error of the real part is given in black, whereas the
-relative error of the imaginary part is given in grey. It is interesting to
-note that the real and imaginary parts have very similar errors in
-\emph{Wer201}, \emph{Key201}, and \emph{And801}, but that the imaginary part of
-\emph{Kong241} seems to behave considerably better for most part than its real
-counterpart. To compare the real and the imaginary parts is insofar interesting
-as the digital filters are purely real valued.
-
-It is very important to note again that other scenarios might yield very
-different error plots. Although the new 201\,pt filter proves to be very
-accurate for these three models, it might not be the best filter in other
-cases. To verify the applicability of the filter we run two different tests
-over a wide range of scenarios. Figure~\ref{fig:Figure-7} shows the relative
-errors of DLF using the filter \emph{Wer201} compared to QWE for a deep water
-model (2\,km water depth), a shallow water model (400\,m water depth), and a
-land model. The models consist of air (upper halfspace), a water layer in the
-first two examples, and a subsurface halfspace. Source depth is 1990\,m, 10\,m,
-and 0.5\,m in the deep, shallow, and land case, respectively. Receiver depth is
-2000\,m, 100\,m, and 0.8\,m. The subsurface halfspace resistivity varies from 1
-to 10, 100, and 1000\,$\Omega\,$m from the left to the right column. The error
-plots cover offsets from 50\,m to 20\,km on the x-axis, and 0.01\,Hz to 10\,Hz
-on the y-axis. It can be seen in the result that the filter is sufficiently
-precise for all practical CSEM cases, hence has generally an error far below
-1\,\% ($10^{-2}$) in all cases. A relative noise of less than 1\,\% is
-generally considered really good for real data. The exceptions are in the
-shallow water case at large offsets with relatively high frequencies, e.g.,
-10\,km offset with 8\,Hz; this zone has a very high error for subsurface
-resistivities of 1\,$\Omega\,$m and 10\,$\Omega\,$m (yellow zones). However,
-this is due to the fact that the amplitudes of the signal for these offsets and
-frequencies are in the order of $10^{-25}\,$V/m. The grey contour lines in the
-plot indicate the power of the amplitude of the signal.
-%
-\plot*{Figure-7}{width=\textwidth}{Relative error plots comparing the DLF
-  (\emph{Wer201}) with QWE for the three cases deep water (top row), shallow
-  water (middle row), and land (bottom row). The subsurface halfspace
-  resistivity varies from 1 to 1000\,$\Omega\,$m from the left column to the
-  right column; offsets vary from 50\,m to 20\,km, and frequencies vary from
-  0.01\,Hz to 10\,Hz. The DLF provides sufficiently accurate results for all
-  practical CSEM applications. The exception is in the shallow water case for
-  $\rho = 1\,\Omega\,$m and $10\,\Omega\,$m for large offsets and high
-  frequencies (yellow zones). However, the amplitudes in these regions are in
-  the order of $10^{-25}\,$V/m, and therefore several orders of
-  magnitude below the noise levels of current EM instrument systems.
-  Contour lines show the power of the amplitude of the signal.}
-%
-
-Another error test is shown in Figure~\ref{fig:Figure-8}. It shows in (a) the
-amplitude of the analytical fullspace solution for a wide range of
-resistivities and frequencies: $\rho=$ $10^{-10}\,\Omega\,$m to
-$10^{10}\,\Omega\,$m, and $f =10^{-10}$\,Hz to $10^{14}$\,Hz. In this case, the
-horizontal offset is set to the skin depth, and the vertical source-receiver
-separation to a tenth of it, where the skin depth is given by \citep[e.g., ][
-equation~1.49]{B.SEG.88.Ward}
-%
-\begin{equation}
-  z_\mr{skin} = \left\{\frac{\omega^2 \varepsilon\mu}{2}
-    \left[ \sqrt{1 + \frac{1}{\left(\omega\varepsilon\rho\right)^2}} -1\right]
-  \right\}^{-1/2} \, , \\
-  \label{eq:skin}
-\end{equation}
-%
-where we used in the example $10\,\varepsilon_0$ for the electric permittivity
-and $\mu_0$ for the magnetic permeability; $\omega = 2\pi f$. The grey contour
-lines in the figure show the magnitude of the offset: the line with 0 is where
-the offset is 1\,m, the line with 3 where the offset is 1\,km, etc. The contour
-lines of the offset have a knee, after which they become almost independent of
-frequency. This is where the quasi static approximation breaks down, and wave
-phenomena become dominant. The red line indicates where $\omega\varepsilon\rho
-= 1$. Subplot (b) shows the relative error of the DLF using the filter
-\emph{Wer201} compared to the analytical solution. It shows nicely that the
-filter can be used for all cases that are in the quasi static region, but not
-outside.
-%
-\plot*{Figure-8}{width=.8\textwidth}{(a) Amplitude of the analytical
-  fullspace solution for a wide range of resistivities and frequencies. The
-  contour lines show the power of the offset used for the calculation, where
-  the offset is set to the skin depth. (b) Relative error of the DLF compared
-  with the analytical solution. The filter \emph{Wer201} works well in the
-  diffusive limit, but fails when wave phenomena become dominant. The red
- line shows where $\omega\varepsilon\rho = 1$.}
-%
-
-We also comment that while it is tempting to design very short length
-filters (much less than about $N=100$) since shorter filters can greatly reduce
-EM simulation times, in our experience, shorter filters tend to be less general
-purpose than longer filters and we highly recommend thoroughly testing their
-accuracy for the desired range of data and model parameters.
-
-\subsection{Ground-penetrating radar}
-
-Figure~\ref{fig:Figure-9} shows the GPR example as calculated in
-\cite{GEO.15.Hunziker} and \cite{GEO.17.Werthmuller}; the model parameters are
-given in subplot (a). The filter used for this example for the Hankel transform
-is a 2001\,pt filter, derived with the fullspace solution with $f=500\,$MHz for
-the inversion and the check of quality. The whole space solution used
-resistivity $200\,\Omega\,$m, relative permittivity $\varepsilon_\mr{r}=10$,
-and permeability $\mu_\mr{r}=1$. Using Equation~\ref{eq:skin} yields a skin
-depth of roughly 3.6\,m for these parameters. The source and receiver vertical
-separation is 1\,m. For the Fourier transform a 4096\,pt FFT was used with
-regularly spaced frequencies from 0.5\,MHz to 850\,MHz and then zero-padded up
-to 2048\,MHz. The frequency result is multiplied with a Ricker wavelet with a
-center frequency of 250\,MHz, and a gain function ($1 + |t^3|, t$ in ns) is
-applied. Subplot (b) shows the result when using the adaptive quadrature (QUAD)
-option of \texttt{empymod}, which uses the QAGSE routine from the Fortran
-QUADPACK library. Subplot (c) shows the result using DLF with the new 2001\,pt
-digital filter. The calculation with DLF took under 9 minutes and is therefore
-roughly 80 times faster than the calculation with QUAD, which took roughly 11
-hours and 27 minutes. Calculating the same model with QWE took 7 hours and 20
-minutes. However, QWE uses in roughly 1/3 of the calculation internally QUAD in
-this example, see \cite{GEO.17.Werthmuller}. Note that DLF was run in parallel
-using 4 threads at once, taking effectively only 2 minutes and 10 seconds to
-calculate.
-% EMmod \citep{GEO.15.Hunziker} took for the same model about 18 hours 55
-% minutes, using a 61\,pt Gauss-Kronrod integration routine.
-The lagged convolution version of DLF and the splined version of QWE were used
-in this comparison.
-%
-\plot*{Figure-9}{width=\textwidth}{GPR example for the model given in (a) using
-  \texttt{empymod} with (b) adaptive quadrature (QUAD) or with (c) a
-  2001\,pt digital linear filter. QUAD took about 11\,h 27\,min to calculate,
-  whereas DLF took 2\,min 10\,s in parallel on 4 nodes, hence under 9\,min in
-  total.}
-%
-
-Figure~\ref{fig:Figure-10-bw} shows in (a)-(c) the real and imaginary
-parts of the frequency domain results and in (d)-(f) the time domain results
-for offsets of 0.2\,m, 2.0\,m, and 3.0\,m.
-%
-\plot*{Figure-10-bw}{width=\textwidth}{The real and imaginary parts of the
-  frequency-domain (a-c) and the time-domain (d-f) responses for
-  horizontal offsets of (a/d) 0.2\,m, (b/e) 2.0\,m, and (c/f) 3.0\,m.}
-%
-
-The 2001\,pt filter was derived with the fullspace solution of a medium which
-is very similar to the layer in which source and receiver reside in our GPR
-example. To show that this filter can also be applied to different layer
-parameters we run a test where we swapped layers one and two of the model in
-Figure~\ref{fig:Figure-9} (a). The resulting real and imaginary parts for the
-offsets 0.2\,m, 2.0\,m, and 3.0\,m are given in Figure~\ref{fig:Figure-11-bw}.
-%
-\plot*{Figure-11-bw}{width=\textwidth}{The real and imaginary parts of the
-  frequency-domain response for the model given in Figure~\ref{fig:Figure-9}
-  (a) with swapped first and second layers for horizontal offsets of (a)
-  0.2\,m, (b) 2.0\,m, and (c) 3.0\,m.}
-%
-
-These examples show clearly that the filter method can indeed be applied to
-high frequency EM modeling and therefore wave phenomena. We are convinced that
-with further tests and analysis much better filters could be achieved, and
-various concepts could be checked. One approach is to derive a filter for each
-frequency band, say one for 1\,MHz--10\,MHz, one for 10\,MHz--100\,MHz, and one
-for 100\,MHz--1\,GHz. Another approach could be to derive distinct filters for
-$J_0$ and $J_1$, with different spacing and shift values. The first idea would
-roughly triple the calculation cost, the second idea double them; however, they
-both would still be very fast compared to standard quadrature methods. One
-could also apply the digital filter method to acoustic and elastic wavefields
-as well.
-
-\section{Conclusions}
-
-The presented, free and open-source code \texttt{fdesign} can be used to design
-digital linear filters for the Hankel and Fourier transforms (and more
-generally for any linear transform) using either analytical transform pairs or
-1D subsurface models together with the EM-modeler \texttt{empymod}. The code is
-available from GitHub as part of \texttt{empymod}; the version corresponding to
-this publication is \texttt{v1.8.1}.
-
-The presented 201\,pt filter achieves more precise results in the three
-presented CSEM cases than other filters, and is included in \texttt{empymod}
-from version 1.4.5 onwards. However, as with any digital filter, the quality
-depends heavily on the model, and this new filter might or might not behave
-that well for other models or data parameters.
-
-The GPR example shows that the digital linear filter method can also be used
-for wave phenomena, not only for the diffusive approximation limit of low
-frequency EM modeling.
-
-We see \texttt{fdesign} as being useful for at least 3 scenarios:
-\begin{enumerate}
-  \item Providing a fast method to design problem-specific filters. For bigger
-    inversion projects (such as, for instance, massive stochastic CSEM
-    inversions) a purpose designed, short filter might save much more time than
-    it costs to design it. This could even be integrated into inversion codes,
-    as an optional pre-inversion step.
-  \item Extending the filter method to new areas, namely higher frequencies and
-    acoustic and elastic wavefields.
-  \item Raising  interest for digital linear filters in geophysics by making it
-    very easy for anyone to play around, create their own filters, and get a
-    better understanding of the relative simplicity of the method. We are sure
-    there are many great filters that are yet to be discovered.
-\end{enumerate}
-
-\section{Acknowledgment}
-
-D.W. would like to thank the electromagnetic research group of A. Mousatov at
-the IMP for fruitful discussions, and J. Hunziker for valuable feedback which
-improved this manuscript considerably. We thank the assistant editor J. Etgen,
-the associate editor J. Dellinger, and the reviewers C. Weiss and R. Mittet
-for valuable inputs not only regarding the manuscript, but also regarding the
-code and its documentation.
-
-% REFERENCES
 \begin{thebibliography}{}
@@
 
+\bibitem[Werthmüller et~al., 2018]{GEO.18.Werthmuller} Werthmüller, D., K. Key,
+  and E. Slob, 2018, A tool for designing digital filters for the Hankel and
+  Fourier transforms in potential, diffusive, and wavefield modeling:
+  Geophysics, {\bf 83}, no. ?, ????--????.
+\newblock (\href{https://doi.org/10.1190/geo2018-0069.1}{doi:
+  10.1190/geo2018-0069.1}).
+
 \end{thebibliography}
 
+\end{multicols}
 \end{document}

### 2019-03-08 DLF-in-Geophysics.tex
Update with final page numbers.

@@
 {\huge \bf \sffamily Digital Linear Filters in Geophysics\\[.5cm]}
-{\small\emph{Excerpt from \cite{GEO.18.Werthmuller}; for the full article see
+{\small\emph{Excerpt from \cite{GEO.19.Werthmuller}; for the full article see
 \href{https://github.com/empymod/article-fdesign}{github.com/empymod/article-fdesign}.}}
@@
 \newblock (\href{https://doi.org/10.1046/j.1365-2478.1997.500292.x}{doi:
-  10.1046/10.1046/j.1365-2478.1997.500292.x}).
+  10.1046/j.1365-2478.1997.500292.x}).
 
@@
 
-\bibitem[Werthmüller et~al., 2018]{GEO.18.Werthmuller} Werthmüller, D., K. Key,
-  and E. Slob, 2018, A tool for designing digital filters for the Hankel and
+\bibitem[Werthmüller et~al., 2019]{GEO.19.Werthmuller} Werthmüller, D., K. Key,
+  and E. Slob, 2019, A tool for designing digital filters for the Hankel and
   Fourier transforms in potential, diffusive, and wavefield modeling:
-  Geophysics, {\bf 83}, no. ?, ????--????.
+  Geophysics, {\bf 84}, no. 2, F47--F56.
 \newblock (\href{https://doi.org/10.1190/geo2018-0069.1}{doi:
